---
title: "Descriptive Statistics"
date: "2026-06-18"
output: html_document
---

# Set-Up
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(dplyr)
library(readr)
library(knitr)
library(kableExtra)
library(ggplot2)
library(tidyr)
library(forcats)
library(stringr)
library(purrr)
library(flextable)
library(officer)

theme_set(theme_minimal(base_size = 12))

raw_path <- "../../data/raw/final_results_capstone_data_ver2.csv"
export_dir <- "descriptive_statistics_exports"

dir.create(export_dir, showWarnings = FALSE, recursive = TRUE)
unlink(
  list.files(export_dir, all.files = TRUE, no.. = TRUE, full.names = TRUE),
  recursive = TRUE
)

responses <- read_csv(raw_path, show_col_types = FALSE) %>%
  select(-starts_with("..."))

cases <- responses %>%
  arrange(case_id, model) %>%
  distinct(case_id, .keep_all = TRUE)

format_pct <- function(x, denom) {
  round(100 * x / denom, 1)
}

normalize_yes <- function(x) {
  str_to_lower(str_squish(as.character(x))) %in%
    c("yes", "y", "true", "1", "present")
}

summarize_categorical <- function(data, variable, label) {
  data %>%
    mutate(value = coalesce(as.character(.data[[variable]]), "Missing")) %>%
    count(value, name = "n") %>%
    mutate(
      Variable = label,
      `%` = format_pct(n, sum(n))
    ) %>%
    rename(Category = value) %>%
    select(Variable, Category, n, `%`) %>%
    arrange(Variable, desc(n))
}

summarize_binary <- function(data, variables) {
  map_dfr(names(variables), function(var) {
    tibble(
      Variable = variables[[var]],
      n = sum(normalize_yes(data[[var]]), na.rm = TRUE),
      `%` = format_pct(n, nrow(data))
    )
  }) %>%
    arrange(desc(n))
}

fmt_n_pct <- function(n, denom) {
  paste0(n, " (", format_pct(n, denom), "%)")
}

continuous_table1_row <- function(data, label, variable) {
  x <- data[[variable]]
  tibble(
    Characteristic = label,
    `Summary, n = 100 cases` = paste0(
      round(mean(x, na.rm = TRUE), 1), " (",
      round(sd(x, na.rm = TRUE), 1), "); ",
      round(median(x, na.rm = TRUE), 1), " [",
      round(quantile(x, 0.25, na.rm = TRUE), 1), ", ",
      round(quantile(x, 0.75, na.rm = TRUE), 1), "]"
    )
  )
}

binary_table1_row <- function(data, label, variable) {
  n_yes <- sum(normalize_yes(data[[variable]]), na.rm = TRUE)
  tibble(
    Characteristic = label,
    `Summary, n = 100 cases` = fmt_n_pct(n_yes, nrow(data))
  )
}

category_table1_rows <- function(data, variable, label) {
  data %>%
    mutate(value = coalesce(as.character(.data[[variable]]), "Missing")) %>%
    count(value, name = "n") %>%
    mutate(
      Characteristic = paste0("&nbsp;&nbsp;", value),
      `Summary, n = 100 cases` = fmt_n_pct(n, sum(n))
    ) %>%
    select(Characteristic, `Summary, n = 100 cases`) %>%
    bind_rows(
      tibble(Characteristic = paste0("<strong>", label, "</strong>"),
             `Summary, n = 100 cases` = ""),
      .
    )
}

drug_patterns <- c(
  "No diabetes medication" = "^$|^none$|newly diagnosed",
  "Metformin" = "metformin",
  "SGLT2 inhibitor" = "gliflozin|empagliflozin|dapagliflozin|canagliflozin",
  "GLP-1 RA / GIP-GLP-1" =
    "semaglutide|dulaglutide|liraglutide|tirzepatide|exenatide|lixisenatide",
  "DPP-4 inhibitor" = "gliptin|sitagliptin|linagliptin|saxagliptin",
  "Sulfonylurea" = "glipizide|glyburide|glimepiride|gliclazide",
  "Basal insulin" = "glargine|detemir|degludec|nph|isophane",
  "Prandial insulin" = "aspart|lispro|glulisine"
)

plot_fill <- "#2F6F73"
plot_fill_light <- "#86B7B3"
plot_accent <- "#B75D3A"
```

```{r derived-case-fields, include = FALSE}
cancer_cols <- intersect(
  c("cancer_breast", "cancer_endometrial", "cancer_colon",
    "cancer_lung", "cancer_prostate"),
  names(cases)
)

cases <- cases %>%
  mutate(
    any_cancer_history = if_any(all_of(cancer_cols), normalize_yes),
    egfr_band = cut(
      egfr_ml_min_1_73m2,
      breaks = c(-Inf, 30, 45, 60, Inf),
      labels = c("<30", "30-44", "45-59", ">=60"),
      right = FALSE
    ),
    hba1c_band = cut(
      hba1c_percent,
      breaks = c(-Inf, 7, 8, 9, 10, Inf),
      labels = c("<7", "7-7.9", "8-8.9", "9-9.9", ">=10"),
      right = FALSE
    ),
    albuminuria_band = cut(
      uacr_mg_g,
      breaks = c(-Inf, 30, 300, Inf),
      labels = c("<30", "30-299", ">=300"),
      right = FALSE
    ),
    ckd_stage_report = coalesce(as.character(stage_ckd), as.character(egfr_band))
  )

cases <- cases %>%
  mutate(
    ckd_stage_report = if_else(
      is.na(ckd_stage_report) | ckd_stage_report == "",
      "Missing",
      ckd_stage_report
    )
  )

model_counts <- responses %>%
  count(model, name = "Response rows") %>%
  mutate(`%` = format_pct(`Response rows`, sum(`Response rows`))) %>%
  arrange(desc(`Response rows`))

case_model_completeness <- responses %>%
  count(case_id, name = "n_model_responses") %>%
  mutate(complete_3_models = n_model_responses == 3)

duplicate_case_model <- responses %>%
  count(case_id, model, name = "n_rows") %>%
  filter(n_rows > 1)

drugs_lc <- str_to_lower(str_squish(coalesce(cases$current_dm_drugs, "")))

baseline_med_summary <- map_dfr(names(drug_patterns), function(label) {
  tibble(
    `Drug class` = label,
    n = sum(str_detect(drugs_lc, drug_patterns[[label]]), na.rm = TRUE)
  )
}) %>%
  mutate(`%` = format_pct(n, nrow(cases))) %>%
  arrange(desc(n))
```

# Dataset Overview

This report describes the current raw vignette dataset at the **case level**.
The raw file has one row per model response, so patient characteristics are first
deduplicated to one row per `case_id`.

```{r dataset-overview}
tibble(
  Metric = c(
    "Raw model-response rows",
    "Unique cases",
    "Unique models",
    "Cases with exactly 3 model responses",
    "Duplicate case-model combinations"
  ),
  Value = c(
    nrow(responses),
    nrow(cases),
    n_distinct(responses$model),
    sum(case_model_completeness$complete_3_models),
    nrow(duplicate_case_model)
  )
) %>%
  kable(caption = "Dataset Overview") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

```{r model-counts}
model_counts %>%
  rename(Model = model) %>%
  kable(caption = "Model Response Rows") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

**Key Takeaway:** The current raw file contains 300 responses from 100 unique
cases and 3 models. Each case has the expected 3 model responses.


# Table 1. Case Characteristics

This table summarizes the variables most likely to be used in the manuscript
description of the synthetic vignette population. Continuous variables are shown
as mean (SD); median [Q1, Q3]. Categorical variables are shown as n (%).

```{r table1}
baseline_med_rows <- baseline_med_summary %>%
  transmute(
    Characteristic = paste0("&nbsp;&nbsp;", `Drug class`),
    `Summary, n = 100 cases` = fmt_n_pct(n, nrow(cases))
  ) %>%
  bind_rows(
    tibble(Characteristic = "<strong>Baseline diabetes medications</strong>",
           `Summary, n = 100 cases` = ""),
    .
  )

table1 <- bind_rows(
  tibble(Characteristic = "<strong>Demographics</strong>",
         `Summary, n = 100 cases` = ""),
  continuous_table1_row(cases, "Age, years", "age"),
  tibble(
    Characteristic = "Female sex",
    `Summary, n = 100 cases` = fmt_n_pct(
      sum(str_to_lower(cases$sex) == "female", na.rm = TRUE),
      nrow(cases)
    )
  ),
  category_table1_rows(cases, "race_ethnicity", "Race / ethnicity"),
  tibble(Characteristic = "<strong>Diabetes and metabolic profile</strong>",
         `Summary, n = 100 cases` = ""),
  continuous_table1_row(cases, "Diabetes duration, years",
                        "duration_dm_years"),
  continuous_table1_row(cases, "HbA1c, %", "hba1c_percent"),
  continuous_table1_row(cases, "BMI, kg/m2", "bmi_kg_m2"),
  continuous_table1_row(cases, "eGFR, mL/min/1.73m2",
                        "egfr_ml_min_1_73m2"),
  continuous_table1_row(cases, "UACR, mg/g", "uacr_mg_g"),
  category_table1_rows(cases, "egfr_band", "eGFR band"),
  category_table1_rows(cases, "albuminuria_band", "Albuminuria band"),
  baseline_med_rows,
  tibble(Characteristic = "<strong>Comorbidities and preferences</strong>",
         `Summary, n = 100 cases` = ""),
  binary_table1_row(cases, "Chronic kidney disease",
                    "chronic_kidney_disease"),
  binary_table1_row(cases, "Heart failure", "heart_failure_present"),
  binary_table1_row(cases, "Hypertension", "hypertension"),
  binary_table1_row(cases, "Hyperlipidemia", "hyperlipidemia"),
  binary_table1_row(cases, "Coronary artery disease",
                    "coronary_artery_disease"),
  binary_table1_row(cases, "MI history", "mi_history"),
  binary_table1_row(cases, "Stroke/TIA history", "stroke_tia_history"),
  tibble(
    Characteristic = "Any cancer history",
    `Summary, n = 100 cases` = fmt_n_pct(
      sum(cases$any_cancer_history, na.rm = TRUE),
      nrow(cases)
    )
  ),
  binary_table1_row(cases, "Financial limitations",
                    "financial_limitations"),
  binary_table1_row(cases, "Reluctant injections", "reluctant_injections")
)

section_rows <- str_detect(table1$Characteristic, "<strong>")

table1_export <- table1 %>%
  mutate(
    Characteristic = str_replace_all(Characteristic, "&nbsp;", " "),
    Characteristic = str_replace_all(Characteristic, "<[^>]+>", "")
  )

table1_ft <- flextable(table1_export) %>%
  bold(i = section_rows, bold = TRUE, part = "body") %>%
  bold(part = "header") %>%
  width(j = 1, width = 3.6) %>%
  width(j = 2, width = 2.2) %>%
  align(j = 2, align = "right", part = "all") %>%
  autofit()

table1_doc <- read_docx() %>%
  body_add_par("Table 1. Case Characteristics", style = "heading 1") %>%
  body_add_flextable(table1_ft)

print(
  table1_doc,
  target = file.path(export_dir, "table1_case_characteristics.docx")
)

table1 %>%
  kable(
    caption = "Table 1. Case Characteristics",
    escape = FALSE,
    align = c("l", "r")
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

**Manuscript Emphasis:** This table should support the background statement that
the vignette set intentionally focuses on older adults with type 2 diabetes and
substantial prescribing complexity, especially kidney disease, albuminuria,
baseline medication use, cardiovascular comorbidity, cancer history, and patient
preferences that can affect medication choice.


# Continuous Case Characteristics

```{r continuous-table}
continuous_vars <- c(
  "Age" = "age",
  "Diabetes duration, years" = "duration_dm_years",
  "HbA1c, %" = "hba1c_percent",
  "eGFR, mL/min/1.73m2" = "egfr_ml_min_1_73m2",
  "UACR, mg/g" = "uacr_mg_g",
  "Creatinine, mg/dL" = "creatinine_mg_dl",
  "BMI, kg/m2" = "bmi_kg_m2",
  "SBP, mmHg" = "sbp",
  "DBP, mmHg" = "dbp",
  "LDL cholesterol, mg/dL" = "ldl_cholesterol_mg_dl",
  "Triglycerides, mg/dL" = "triglycerides_mg_dl"
)

map_dfr(names(continuous_vars), function(label) {
  x <- cases[[continuous_vars[[label]]]]
  tibble(
    Variable = label,
    N = sum(!is.na(x)),
    Mean = mean(x, na.rm = TRUE),
    SD = sd(x, na.rm = TRUE),
    Median = median(x, na.rm = TRUE),
    Q1 = quantile(x, 0.25, na.rm = TRUE),
    Q3 = quantile(x, 0.75, na.rm = TRUE),
    Min = min(x, na.rm = TRUE),
    Max = max(x, na.rm = TRUE)
  )
}) %>%
  mutate(across(where(is.numeric), ~ round(., 1))) %>%
  kable(caption = "Continuous Characteristics (n = 100 cases)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r continuous-histograms, fig.height=7, fig.width=10}
hist_vars <- c(
  "Age" = "age",
  "HbA1c (%)" = "hba1c_percent",
  "eGFR" = "egfr_ml_min_1_73m2",
  "BMI" = "bmi_kg_m2",
  "UACR" = "uacr_mg_g"
)

continuous_histogram_plot <- cases %>%
  select(all_of(hist_vars)) %>%
  pivot_longer(everything(), names_to = "Variable", values_to = "Value") %>%
  mutate(Variable = recode(Variable, !!!setNames(names(hist_vars), hist_vars))) %>%
  ggplot(aes(x = Value)) +
  geom_histogram(fill = plot_fill, color = "white", bins = 15) +
  facet_wrap(~ Variable, scales = "free", ncol = 2) +
  labs(x = NULL, y = "Number of cases",
       title = "Distribution of Key Continuous Characteristics") +
  theme(panel.grid.minor = element_blank())

ggsave(
  file.path(export_dir, "figure2_continuous_characteristics_histograms.png"),
  plot = continuous_histogram_plot,
  width = 10,
  height = 7,
  dpi = 300
)

continuous_histogram_plot
```


# Demographic Characteristics

```{r demographics-table}
bind_rows(
  summarize_categorical(cases, "sex", "Sex"),
  summarize_categorical(cases, "race_ethnicity", "Race / ethnicity")
) %>%
  kable(caption = "Demographic Characteristics (n = 100 cases)") %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  collapse_rows(columns = 1, valign = "top")
```

```{r demographics-plots, fig.height=4.5, fig.width=10}
demo_plot_data <- bind_rows(
  summarize_categorical(cases, "sex", "Sex"),
  summarize_categorical(cases, "race_ethnicity", "Race / ethnicity")
)

demographics_plot <- ggplot(
  demo_plot_data,
  aes(x = fct_reorder(Category, n), y = n)
) +
  geom_col(fill = plot_fill_light) +
  geom_text(aes(label = paste0(n, " (", `%`, "%)")),
            hjust = -0.05, size = 3.4) +
  facet_wrap(~ Variable, scales = "free_y") +
  coord_flip() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  labs(x = NULL, y = "Number of cases",
       title = "Sex and Race/Ethnicity Distribution") +
  theme(panel.grid.minor = element_blank())

ggsave(
  file.path(export_dir, "figure3_demographic_characteristics.png"),
  plot = demographics_plot,
  width = 10,
  height = 4.5,
  dpi = 300
)

demographics_plot
```


# Clinical Subgroup Distributions

```{r clinical-subgroup-table}
clinical_subgroups <- bind_rows(
  summarize_categorical(cases, "hba1c_band", "HbA1c band"),
  summarize_categorical(cases, "egfr_band", "eGFR band"),
  summarize_categorical(cases, "albuminuria_band", "Albuminuria band"),
  summarize_categorical(cases, "ckd_stage_report", "CKD stage")
) %>%
  mutate(
    Variable = factor(
      Variable,
      levels = c("HbA1c band", "eGFR band", "Albuminuria band", "CKD stage")
    )
  )

clinical_subgroups %>%
  kable(caption = "Clinical Subgroup Distributions (n = 100 cases)") %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  collapse_rows(columns = 1, valign = "top")
```

```{r clinical-subgroup-plots, fig.height=5.5, fig.width=10}
clinical_subgroup_plot_data <- clinical_subgroups %>%
  filter(Variable %in% c("eGFR band", "HbA1c band", "Albuminuria band")) %>%
  mutate(
    Category = case_when(
      Variable == "HbA1c band" ~ factor(
        as.character(Category),
        levels = c("<7", "7-7.9", "8-8.9", "9-9.9", ">=10")
      ),
      Variable == "eGFR band" ~ factor(
        as.character(Category),
        levels = c(">=60", "45-59", "30-44", "<30")
      ),
      Variable == "Albuminuria band" ~ factor(
        as.character(Category),
        levels = c("<30", "30-299", ">=300")
      ),
      TRUE ~ factor(as.character(Category))
    )
  )

clinical_subgroup_plot <- clinical_subgroup_plot_data %>%
  ggplot(aes(x = Category, y = n)) +
  geom_col(fill = plot_fill) +
  geom_text(aes(label = paste0(n, " (", `%`, "%)")),
            vjust = -0.3, size = 3.4) +
  facet_wrap(~ Variable, scales = "free_x") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18))) +
  labs(x = NULL, y = "Number of cases",
       title = "Distribution of HbA1c, eGFR, and Albuminuria Categories") +
  theme(panel.grid.minor = element_blank())

ggsave(
  file.path(export_dir, "figure1_clinical_subgroup_distributions.png"),
  plot = clinical_subgroup_plot,
  width = 10,
  height = 5.5,
  dpi = 300
)

clinical_subgroup_plot
```

**Key Takeaway:** The vignette set includes many patients with clinically
important kidney disease and albuminuria, which are central to diabetes
medication selection.


# Comorbidities and Preferences

```{r comorbidity-table}
binary_vars <- c(
  chronic_kidney_disease = "Chronic kidney disease",
  heart_failure_present = "Heart failure",
  hypertension = "Hypertension",
  hyperlipidemia = "Hyperlipidemia",
  coronary_artery_disease = "Coronary artery disease",
  mi_history = "MI history",
  stroke_tia_history = "Stroke/TIA history",
  peripheral_neuropathy = "Peripheral neuropathy",
  diabetic_retinopathy = "Diabetic retinopathy",
  masld_present = "MASLD",
  mash_present = "MASH",
  depression = "Depression",
  copd = "COPD",
  asthma = "Asthma",
  anemia = "Anemia",
  osteoporosis = "Osteoporosis",
  gi_distress_history = "GI distress history",
  gu_infection_history = "GU infection history",
  financial_limitations = "Financial limitations",
  preference_tighter_control = "Preference: tighter control",
  preference_simple_regimen = "Preference: simple regimen",
  reluctant_injections = "Reluctant injections"
)

binary_summary <- summarize_binary(cases, binary_vars) %>%
  bind_rows(tibble(
    Variable = "Any cancer history",
    n = sum(cases$any_cancer_history, na.rm = TRUE),
    `%` = format_pct(n, nrow(cases))
  )) %>%
  arrange(desc(n))

binary_summary %>%
  kable(caption = "Comorbidities and Preferences (n = 100 cases)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r comorbidity-plot, fig.height=8, fig.width=10}
comorbidity_plot <- binary_summary %>%
  mutate(Variable = fct_reorder(Variable, n)) %>%
  ggplot(aes(x = Variable, y = n)) +
  geom_col(fill = plot_accent) +
  geom_text(aes(label = paste0(n, " (", `%`, "%)")),
            hjust = -0.05, size = 3.2) +
  coord_flip() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  labs(x = NULL, y = "Number of cases",
       title = "Prevalence of Comorbidities and Patient Preference Features") +
  theme(panel.grid.minor = element_blank())

ggsave(
  file.path(export_dir, "figure4_comorbidities_and_preferences.png"),
  plot = comorbidity_plot,
  width = 10,
  height = 8,
  dpi = 300
)

comorbidity_plot
```


# Baseline Diabetes Medications

```{r baseline-medications}
baseline_med_summary %>%
  kable(caption = paste0(
    "Baseline Diabetes Medications (n = ", nrow(cases),
    " cases; classes overlap)"
  )) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

```{r baseline-medication-plot, fig.height=5, fig.width=9}
baseline_medication_plot <- baseline_med_summary %>%
  mutate(`Drug class` = fct_reorder(`Drug class`, n)) %>%
  ggplot(aes(x = `Drug class`, y = n)) +
  geom_col(fill = plot_fill) +
  geom_text(aes(label = paste0(n, " (", `%`, "%)")),
            hjust = -0.05, size = 3.4) +
  coord_flip() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  labs(x = NULL, y = "Number of cases",
       title = "Baseline Diabetes Medication Classes") +
  theme(panel.grid.minor = element_blank())

ggsave(
  file.path(export_dir, "figure5_baseline_diabetes_medications.png"),
  plot = baseline_medication_plot,
  width = 9,
  height = 5,
  dpi = 300
)

baseline_medication_plot
```

**Key Takeaway:** Baseline medication percentages can sum to more than 100%
because patients may be on combination regimens. The no-medication group
represents patients with `current_dm_drugs` recorded as `none` or blank.


# Missingness and Data Quality

```{r missingness}
key_missing_vars <- c(
  "case_id", "vignette_text", "age", "sex", "race_ethnicity",
  "duration_dm_years", "current_dm_drugs", "hba1c_percent",
  "egfr_ml_min_1_73m2", "uacr_mg_g", "bmi_kg_m2",
  "heart_failure_present", "chronic_kidney_disease", "stage_ckd",
  "financial_limitations", "reluctant_injections"
)

missingness <- map_dfr(key_missing_vars, function(var) {
  x <- cases[[var]]
  missing_n <- sum(is.na(x) | str_squish(as.character(x)) == "")
  tibble(
    Variable = var,
    `Missing (n)` = missing_n,
    `Missing (%)` = format_pct(missing_n, nrow(cases))
  )
}) %>%
  arrange(desc(`Missing (%)`), Variable)

missingness %>%
  kable(caption = "Missingness in Key Case-Level Variables") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

```{r case-completeness}
case_model_completeness %>%
  filter(!complete_3_models) %>%
  rename(
    `Case ID` = case_id,
    `N model responses` = n_model_responses,
    `Complete 3-model case` = complete_3_models
  ) %>%
  kable(caption = "Incomplete Cases (expected = 3 model responses)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

```{r duplicates}
duplicate_case_model %>%
  rename(`Case ID` = case_id, Model = model, `N rows` = n_rows) %>%
  kable(caption = "Duplicated Case-Model Combinations (expected = 1 row each)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

```{r dm026-check}
cases %>%
  filter(str_detect(case_id, "^DM026")) %>%
  select(`Case ID` = case_id, `Vignette text` = vignette_text) %>%
  kable(caption = "DM026 Case Labels in Current Dataset") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

**Key Takeaway:** The previous DM026 labeling problem is resolved in this raw
dataset because the three variants are represented as `DM026(1)`, `DM026(2)`,
and `DM026(3)`. All 100 cases have the expected three model responses.
