---
title: "ADA Guideline Knowledge Base and Executable Decision Model Demo"
date: "2026-07-07"
output:
  html_document:
    toc: true
    toc_depth: 3
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)

library(dplyr)
library(readr)
library(stringr)
library(tidyr)
library(purrr)
library(tibble)
library(knitr)
library(kableExtra)
library(jsonlite)

raw_path <- "../../data/raw/final_results_capstone_data_ver2.csv"
kb_dir <- "knowledge_base"
demo_dir <- "demo_outputs"

dir.create(kb_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(demo_dir, showWarnings = FALSE, recursive = TRUE)

clean_yes <- function(x) {
  str_to_lower(str_squish(as.character(x))) %in%
    c("yes", "y", "true", "1", "present")
}

clean_text <- function(x) {
  str_to_lower(str_squish(coalesce(as.character(x), "")))
}

format_model <- function(x) {
  case_when(
    str_detect(x, "gpt-5\\.4") ~ "GPT-5.4",
    str_detect(x, "claude-opus") ~ "Claude Opus 4.6",
    str_detect(x, "claude-sonnet") ~ "Claude Sonnet 4.6",
    TRUE ~ x
  )
}

show_table <- function(data, caption = NULL) {
  data %>%
    kable(caption = caption) %>%
    kable_styling(
      full_width = FALSE,
      bootstrap_options = c("striped", "hover", "condensed")
    )
}
```

# Purpose

This workflow creates a small ADA 2026 pharmacotherapy knowledge base and
demonstrates an executable guideline decision model on the first six rows
of `final_results_capstone_data_ver2.csv`.

The key design choice is that the decision tree is **not only a CSV**.
The CSV rule registry is an audit table, while the actual decision model is
implemented as deterministic R functions in this R Markdown file.

# Load Demo Data

```{r load-data}
raw_data <- read_csv(raw_path, show_col_types = FALSE) %>%
  select(-starts_with("..."))

demo_responses <- raw_data %>%
  slice(1:6) %>%
  mutate(model_clean = format_model(model))

demo_cases <- demo_responses %>%
  arrange(case_id, model_clean) %>%
  distinct(case_id, .keep_all = TRUE)

demo_overview <- tibble(
  metric = c("demo response rows", "unique demo cases", "unique models"),
  value = c(
    nrow(demo_responses),
    n_distinct(demo_responses$case_id),
    n_distinct(demo_responses$model_clean)
  )
)
```

```{r show-demo-overview}
show_table(demo_overview, "Demo data overview")
```

```{r show-demo-cases}
demo_cases %>%
  select(
    case_id, age, sex, race_ethnicity, hba1c_percent,
    egfr_ml_min_1_73m2, uacr_mg_g, bmi_kg_m2,
    current_dm_drugs, heart_failure_present, hf_type,
    chronic_kidney_disease, stage_ckd, coronary_artery_disease,
    mi_history, masld_present, mash_present, financial_limitations,
    preference_simple_regimen, reluctant_injections
  ) %>%
  show_table("Case-level patient features used in the demo")
```

# Knowledge Base Layer 1: Text Chunks With Metadata

These chunks are short derived summaries with metadata. They are intended
for retrieval and traceability, not redistribution of long ADA passages.

```{r text-chunks}
ada_text_chunks <- tribble(
  ~chunk_id, ~source_type, ~recommendation_id, ~source_page,
  ~condition, ~drug_class, ~clinical_trigger, ~action_type,
  ~rule_strength, ~evidence_grade, ~summary_text, ~retrieval_keywords,
  "ADA2026_REC_9_7_ASCVD_GLP1_SGLT2", "recommendation", "9.7", 8,
  "ASCVD or high CVD risk", "GLP-1 RA; SGLT2 inhibitor",
  "established ASCVD or high CVD risk", "include",
  "strong_recommend", "A",
  "For type 2 diabetes with established or high ASCVD risk, include a GLP-1 RA and/or SGLT2 inhibitor with cardiovascular benefit irrespective of A1c.",
  "ASCVD CVD GLP-1 SGLT2 cardiovascular benefit",
  "ADA2026_REC_9_8_HF_SGLT2", "recommendation", "9.8", 8,
  "Heart failure", "SGLT2 inhibitor",
  "heart failure with reduced or preserved ejection fraction", "recommend",
  "strong_recommend", "A",
  "For type 2 diabetes with heart failure, recommend an SGLT2 inhibitor for glycemic management and prevention of HF hospitalization irrespective of A1c.",
  "heart failure HFpEF HFrEF SGLT2",
  "ADA2026_REC_9_10_CKD_SGLT2_GLP1", "recommendation", "9.10", 8,
  "CKD", "SGLT2 inhibitor; GLP-1 RA",
  "eGFR 20-60 or albuminuria", "use",
  "strong_recommend", "A",
  "For type 2 diabetes with CKD, use an SGLT2 inhibitor or GLP-1 RA with demonstrated benefit for glycemic management, CKD progression, and cardiovascular event reduction.",
  "CKD eGFR albuminuria SGLT2 GLP-1",
  "ADA2026_REC_9_11_ADV_CKD_GLP1", "recommendation", "9.11", 8,
  "Advanced CKD", "GLP-1 RA",
  "eGFR below 30", "prefer",
  "strong_recommend", "B/C",
  "For type 2 diabetes with advanced CKD, GLP-1 RA is preferred for glycemic management because of lower hypoglycemia risk and cardiovascular benefit.",
  "advanced CKD eGFR <30 GLP-1",
  "ADA2026_REC_9_18_DPP4_GLP1_AVOID", "recommendation", "9.18", 10,
  "Medication combination", "DPP-4 inhibitor; GLP-1 RA",
  "concurrent DPP-4 inhibitor and GLP-1 RA", "avoid",
  "hard_avoid_if_combined", "B",
  "Concurrent DPP-4 inhibitor use with GLP-1 RA or dual GIP/GLP-1 RA is not recommended because it adds no glucose-lowering benefit.",
  "DPP-4 GLP-1 combination avoid",
  "ADA2026_REC_9_20_INSULIN_SEVERE_HYPERGLYCEMIA", "recommendation", "9.20", 10,
  "Severe hyperglycemia", "insulin",
  "symptoms, A1c >10, or glucose >=300", "consider",
  "preference_sensitive", "E",
  "Insulin initiation should be considered when symptoms are present or A1c/glucose levels are very high.",
  "insulin A1c >10 glucose 300 symptoms",
  "ADA2026_REC_9_21_GLP1_BEFORE_INSULIN", "recommendation", "9.21", 10,
  "Injectable therapy", "GLP-1 RA",
  "no severe hyperglycemia or crisis", "prefer",
  "strong_recommend", "A",
  "Without severe hyperglycemia or crisis, GLP-1-based therapy is preferred to insulin for initial or add-on glucose-lowering therapy.",
  "GLP-1 before insulin injectable therapy",
  "ADA2026_REC_9_22_GLP1_WITH_INSULIN", "recommendation", "9.22", 10,
  "Insulin combination", "GLP-1 RA; insulin",
  "insulin is used", "combine_or_reassess",
  "strong_recommend", "A",
  "If insulin is used, GLP-1 RA combination is recommended for glycemic effectiveness and benefits on weight and hypoglycemia risk; insulin dosing should be reassessed.",
  "insulin GLP-1 combination dose reassess",
  "ADA2026_METFORMIN_EGFR_RATIONALE", "rationale", NA_character_, 15,
  "Kidney function", "metformin",
  "eGFR below 30 or eGFR 30-45", "avoid_or_caution",
  "hard_avoid", NA_character_,
  "Metformin is generally safe at eGFR >=30; risk rises when eGFR is below 30, and eGFR 30-45 warrants caution.",
  "metformin eGFR 30 lactic acidosis kidney"
)

write_csv(ada_text_chunks, file.path(kb_dir, "ada_text_chunks.csv"))
```

```{r show-text-chunks}
ada_text_chunks %>%
  select(
    chunk_id, recommendation_id, source_page, condition,
    drug_class, action_type, rule_strength, summary_text
  ) %>%
  show_table("Text chunks with metadata")
```

# Knowledge Base Layer 2: Structured Medication Table

This table is a drug-class property table derived from ADA Table 9.2. It is
not a detailed prescribing-dose table.

```{r medication-table}
ada_medication_table_structured <- tribble(
  ~drug_class, ~route, ~glycemic_efficacy, ~hypoglycemia_risk,
  ~weight_effect, ~mace_effect, ~heart_failure_effect,
  ~ckd_progression_effect, ~kidney_dosing_consideration,
  ~mash_effect, ~major_cautions, ~contraindicated_egfr_lt30,
  ~can_start_egfr_ge20_for_cardiorenal_benefit,
  ~reduced_glycemic_effect_egfr_lt45,
  ~renal_dose_adjustment_needed, ~ckd_hypoglycemia_caution,
  ~dose_label_default, ~source_table, ~source_page,
  "metformin", "oral", "high", "no",
  "neutral/modest loss", "potential benefit", "neutral",
  "neutral", "contraindicated with eGFR <30",
  "neutral", "GI intolerance; vitamin B12 deficiency",
  TRUE, FALSE, FALSE, "yes", FALSE,
  "same_or_reduced_dose", "Table 9.2", 11,
  "SGLT2 inhibitor", "oral", "intermediate to high", "no",
  "loss", "benefit for selected agents", "benefit for selected agents",
  "benefit for selected agents",
  "glucose-lowering effect reduced at eGFR <45; may start/continue for cardiorenal benefit if eGFR >20",
  "unknown", "DKA risk; genital infections; UTI; volume depletion",
  FALSE, TRUE, TRUE, "agent_specific", FALSE,
  "standard_starting_dose_if_eligible", "Table 9.2", 11,
  "GLP-1 RA", "subcutaneous/oral semaglutide", "high to very high", "no",
  "loss", "benefit for selected agents", "benefit for semaglutide SQ",
  "kidney endpoint benefit for selected agents",
  "no dose adjustment for dulaglutide/liraglutide/semaglutide; monitor kidney function with severe GI reactions",
  "benefit for semaglutide SQ", "GI effects; pancreatitis concern; gallbladder disease; retinopathy monitoring",
  FALSE, FALSE, FALSE, "no", FALSE,
  "standard_starting_dose", "Table 9.2", 12,
  "dual GIP/GLP-1 RA", "subcutaneous", "very high", "no",
  "loss very high", "under investigation", "benefit for tirzepatide",
  "potential benefit", "no dose adjustment; monitor kidney function with severe GI reactions",
  "potential benefit", "GI effects; pancreatitis concern; gallbladder disease",
  FALSE, FALSE, FALSE, "no", FALSE,
  "standard_starting_dose", "Table 9.2", 12,
  "DPP-4 inhibitor", "oral", "intermediate", "no",
  "neutral", "neutral", "neutral/potential saxagliptin risk",
  "neutral", "renal dose adjustment for sitagliptin/saxagliptin/alogliptin; not linagliptin",
  "unknown", "pancreatitis reports; joint pain; bullous pemphigoid",
  FALSE, FALSE, FALSE, "yes", FALSE,
  "renal_adjusted_dose_if_needed", "Table 9.2", 13,
  "pioglitazone/TZD", "oral", "high", "no",
  "gain", "potential benefit", "increased HF risk",
  "neutral", "generally not recommended in kidney impairment due to fluid retention",
  "potential benefit", "fluid retention/HF; fracture risk; bladder cancer caution",
  FALSE, FALSE, FALSE, "no", FALSE,
  "avoid_in_hf", "Table 9.2", 13,
  "sulfonylurea", "oral", "high", "yes",
  "gain", "neutral", "neutral",
  "neutral", "glyburide generally not recommended in CKD; glipizide/glimepiride conservative initiation",
  "unknown", "hypoglycemia risk; CV warning based on older sulfonylurea",
  FALSE, FALSE, FALSE, "agent_specific", TRUE,
  "conservative_start_or_avoid_high_risk", "Table 9.2", 13,
  "insulin", "subcutaneous", "high to very high", "yes",
  "gain", "neutral", "neutral",
  "neutral", "lower doses required as eGFR decreases; titrate clinically",
  "unknown", "injection reactions; hypoglycemia risk increases with kidney impairment",
  FALSE, FALSE, FALSE, "yes", TRUE,
  "individualized", "Table 9.2", 14
)

write_csv(
  ada_medication_table_structured,
  file.path(kb_dir, "ada_medication_table_structured.csv")
)
```

```{r show-medication-table}
ada_medication_table_structured %>%
  select(
    drug_class, glycemic_efficacy, hypoglycemia_risk, weight_effect,
    heart_failure_effect, ckd_progression_effect,
    kidney_dosing_consideration, major_cautions
  ) %>%
  show_table("Structured Medication Table from ADA Table 9.2")
```

# Knowledge Base Layer 3: Rule Registry

The rule registry is a human-readable audit table. It documents the
decision logic, but it is not the executable decision model. The executable
model is implemented in the next section.

```{r rule-registry}
ada_decision_rule_registry <- tribble(
  ~rule_id, ~source_figure_or_recommendation, ~source_page,
  ~trigger_logic, ~condition, ~drug_class, ~expected_action,
  ~expected_dose_label, ~rule_strength, ~exception_variables,
  ~supporting_chunk_ids, ~reason_template,
  "ADA2026_RULE_METFORMIN_EGFR_LT30", "Metformin rationale/Table 9.2", "11, 15",
  "egfr_ml_min_1_73m2 < 30", "advanced CKD", "metformin",
  "avoid_or_stop", "do_not_use", "hard_avoid", "none",
  "ADA2026_METFORMIN_EGFR_RATIONALE",
  "eGFR is below 30, so metformin should be avoided or stopped.",
  "ADA2026_RULE_METFORMIN_EGFR_GE30_BASELINE", "Metformin rationale/Table 9.2", "11, 15",
  "egfr_ml_min_1_73m2 >= 30 and baseline_metformin", "CKD caution", "metformin",
  "continue_allowed_or_renal_caution", "same_or_reduced_dose", "caution", "GI intolerance; low eGFR trend",
  "ADA2026_METFORMIN_EGFR_RATIONALE",
  "eGFR is at least 30 and metformin is already used; continuation is allowed with kidney caution.",
  "ADA2026_RULE_HF_SGLT2I", "Recommendation 9.8/Figure 9.4", "8",
  "heart_failure", "heart failure", "SGLT2 inhibitor",
  "recommend", "standard_starting_dose_if_eligible", "strong_recommend", "GU infection; intolerance; cost",
  "ADA2026_REC_9_8_HF_SGLT2",
  "Heart failure supports SGLT2 inhibitor use if eligible.",
  "ADA2026_RULE_CKD_SGLT2I_OR_GLP1RA", "Recommendation 9.10/Figure 9.4", "8",
  "ckd_or_albuminuria and sglt2i_eligible_by_egfr", "CKD/albuminuria", "SGLT2 inhibitor",
  "recommend_or_consider", "standard_starting_dose_if_eligible", "strong_recommend", "GU infection; intolerance; cost",
  "ADA2026_REC_9_10_CKD_SGLT2_GLP1",
  "CKD or albuminuria with eGFR above threshold supports SGLT2 inhibitor consideration.",
  "ADA2026_RULE_ADVANCED_CKD_GLP1RA", "Recommendation 9.11/Figure 9.4", "8",
  "advanced_ckd", "advanced CKD", "GLP-1 RA",
  "prefer_or_consider", "standard_starting_dose", "strong_recommend", "severe GI intolerance; pancreatitis risk",
  "ADA2026_REC_9_11_ADV_CKD_GLP1",
  "Advanced CKD supports GLP-1 RA preference for glycemic management.",
  "ADA2026_RULE_ASCVD_GLP1RA", "Recommendation 9.7/Figure 9.4", "8",
  "ascvd", "ASCVD", "GLP-1 RA",
  "consider", "standard_starting_dose", "strong_recommend", "severe GI intolerance; pancreatitis risk",
  "ADA2026_REC_9_7_ASCVD_GLP1_SGLT2",
  "ASCVD supports GLP-1 RA consideration for cardiovascular benefit.",
  "ADA2026_RULE_ASCVD_SGLT2I", "Recommendation 9.7/Figure 9.4", "8",
  "ascvd", "ASCVD", "SGLT2 inhibitor",
  "consider", "standard_starting_dose_if_eligible", "strong_recommend", "GU infection; intolerance; cost",
  "ADA2026_REC_9_7_ASCVD_GLP1_SGLT2",
  "ASCVD supports SGLT2 inhibitor consideration for cardiovascular benefit.",
  "ADA2026_RULE_OBESITY_MASH_GLP1RA", "Figure 9.4/Table 9.2", "9, 12",
  "obesity or mash_or_masld", "weight/MASLD/MASH", "GLP-1 RA",
  "consider", "standard_starting_dose", "preference_sensitive", "injection reluctance; severe GI intolerance; cost",
  "ADA2026_REC_9_7_ASCVD_GLP1_SGLT2",
  "Obesity or MASLD/MASH supports GLP-1 RA consideration.",
  "ADA2026_RULE_DPP4_WITH_GLP1_AVOID", "Recommendation 9.18", "10",
  "glp1_expected and baseline_dpp4i", "GLP-1 + DPP-4 combination", "DPP-4 inhibitor",
  "stop_or_avoid", "do_not_use_with_glp1", "hard_avoid_if_combined", "none",
  "ADA2026_REC_9_18_DPP4_GLP1_AVOID",
  "If GLP-1 RA is expected and DPP-4 inhibitor is present, stop or avoid DPP-4 inhibitor.",
  "ADA2026_RULE_A1C_GT10_INSULIN_CONSIDER", "Recommendation 9.20", "10",
  "severe_hyperglycemia", "severe hyperglycemia", "insulin",
  "consider", "individualized", "preference_sensitive", "GLP-1 option if no crisis",
  "ADA2026_REC_9_20_INSULIN_SEVERE_HYPERGLYCEMIA",
  "Very high A1c supports considering insulin.",
  "ADA2026_RULE_INSULIN_ADV_CKD_CAUTION", "Table 9.2", "14",
  "baseline_insulin and advanced_ckd", "advanced CKD on insulin", "insulin",
  "dose_caution_or_deintensify", "individualized_lower_dose_possible", "caution", "clinical response",
  "ADA2026_REC_9_22_GLP1_WITH_INSULIN",
  "Insulin doses may need reduction as eGFR decreases.",
  "ADA2026_RULE_PRANDIAL_INSULIN_SIMPLE_REGIMEN", "Person-centered treatment", "8",
  "baseline_prandial_insulin and simple_regimen_preference", "regimen simplification", "prandial insulin",
  "consider_deintensification", "individualized", "preference_sensitive", "glycemic control needs",
  "ADA2026_REC_9_5_PERSON_CENTERED",
  "Prandial insulin deintensification may align with simple-regimen preference if clinically appropriate."
)

write_csv(
  ada_decision_rule_registry,
  file.path(kb_dir, "ada_decision_rule_registry.csv")
)

ada_knowledge_base <- list(
  text_chunks = ada_text_chunks,
  medication_table = ada_medication_table_structured,
  decision_rule_registry = ada_decision_rule_registry
)

write_json(
  ada_knowledge_base,
  file.path(kb_dir, "ada_knowledge_base.json"),
  pretty = TRUE,
  auto_unbox = TRUE,
  na = "null"
)
```

```{r show-rule-registry}
ada_decision_rule_registry %>%
  select(
    rule_id, trigger_logic, drug_class, expected_action,
    expected_dose_label, rule_strength, supporting_chunk_ids
  ) %>%
  show_table("Decision rule registry for audit")
```

# Executable Decision Model

The following R functions are the actual deterministic decision model.

```{r decision-model-functions, echo=TRUE}
derive_patient_triggers <- function(cases) {
  cases %>%
    mutate(
      current_dm_drugs_clean = clean_text(current_dm_drugs),
      advanced_ckd = egfr_ml_min_1_73m2 < 30,
      ckd_or_albuminuria = egfr_ml_min_1_73m2 < 60 |
        uacr_mg_g >= 30 |
        clean_yes(chronic_kidney_disease),
      sglt2i_eligible_by_egfr = egfr_ml_min_1_73m2 > 20,
      heart_failure = clean_yes(heart_failure_present),
      ascvd = clean_yes(mi_history) |
        clean_yes(stroke_tia_history) |
        clean_yes(coronary_artery_disease),
      obesity = bmi_kg_m2 >= 30,
      mash_or_masld = clean_yes(masld_present) | clean_yes(mash_present),
      severe_hyperglycemia = hba1c_percent > 10,
      baseline_metformin = str_detect(current_dm_drugs_clean, "metformin"),
      baseline_dpp4i = str_detect(
        current_dm_drugs_clean,
        "sitagliptin|linagliptin|saxagliptin|alogliptin"
      ),
      baseline_insulin = str_detect(current_dm_drugs_clean, "insulin"),
      baseline_prandial_insulin = str_detect(
        current_dm_drugs_clean,
        "insulin aspart|insulin lispro|regular insulin"
      ),
      cost_barrier = clean_yes(financial_limitations),
      injection_reluctance = clean_yes(reluctant_injections),
      gu_infection_caution = clean_yes(gu_infection_history),
      simple_regimen_preference = clean_yes(preference_simple_regimen)
    )
}

build_action <- function(data, rule_id, drug_class) {
  rule <- ada_decision_rule_registry %>%
    filter(.data$rule_id == !!rule_id) %>%
    slice(1)

  data %>%
    transmute(
      case_id,
      rule_id = rule$rule_id,
      drug_class = drug_class,
      expected_action = rule$expected_action,
      expected_dose_label = rule$expected_dose_label,
      rule_strength = rule$rule_strength,
      supporting_chunk_ids = rule$supporting_chunk_ids,
      expected_reason = rule$reason_template
    )
}

apply_ada_decision_model <- function(triggered_cases) {
  actions <- list(
    triggered_cases %>%
      filter(advanced_ckd) %>%
      build_action("ADA2026_RULE_METFORMIN_EGFR_LT30", "metformin"),
    triggered_cases %>%
      filter(!advanced_ckd, baseline_metformin, egfr_ml_min_1_73m2 >= 30) %>%
      build_action("ADA2026_RULE_METFORMIN_EGFR_GE30_BASELINE", "metformin"),
    triggered_cases %>%
      filter(heart_failure) %>%
      build_action("ADA2026_RULE_HF_SGLT2I", "SGLT2 inhibitor"),
    triggered_cases %>%
      filter(ckd_or_albuminuria, sglt2i_eligible_by_egfr) %>%
      build_action("ADA2026_RULE_CKD_SGLT2I_OR_GLP1RA", "SGLT2 inhibitor"),
    triggered_cases %>%
      filter(advanced_ckd) %>%
      build_action("ADA2026_RULE_ADVANCED_CKD_GLP1RA", "GLP-1 RA"),
    triggered_cases %>%
      filter(ascvd) %>%
      build_action("ADA2026_RULE_ASCVD_GLP1RA", "GLP-1 RA"),
    triggered_cases %>%
      filter(ascvd) %>%
      build_action("ADA2026_RULE_ASCVD_SGLT2I", "SGLT2 inhibitor"),
    triggered_cases %>%
      filter(obesity | mash_or_masld) %>%
      build_action("ADA2026_RULE_OBESITY_MASH_GLP1RA", "GLP-1 RA"),
    triggered_cases %>%
      filter(severe_hyperglycemia) %>%
      build_action("ADA2026_RULE_A1C_GT10_INSULIN_CONSIDER", "insulin"),
    triggered_cases %>%
      filter(baseline_insulin, advanced_ckd) %>%
      build_action("ADA2026_RULE_INSULIN_ADV_CKD_CAUTION", "insulin"),
    triggered_cases %>%
      filter(baseline_prandial_insulin, simple_regimen_preference) %>%
      build_action("ADA2026_RULE_PRANDIAL_INSULIN_SIMPLE_REGIMEN", "prandial insulin")
  )

  base_actions <- bind_rows(actions)

  glp1_expected_cases <- base_actions %>%
    filter(drug_class == "GLP-1 RA") %>%
    distinct(case_id)

  dpp4_actions <- triggered_cases %>%
    semi_join(glp1_expected_cases, by = "case_id") %>%
    filter(baseline_dpp4i) %>%
    build_action("ADA2026_RULE_DPP4_WITH_GLP1_AVOID", "DPP-4 inhibitor")

  bind_rows(base_actions, dpp4_actions) %>%
    arrange(case_id, drug_class, rule_id)
}
```

```{r apply-decision-model}
demo_patient_triggers <- derive_patient_triggers(demo_cases)
demo_expected_guideline_actions <- apply_ada_decision_model(demo_patient_triggers)

write_csv(
  demo_patient_triggers,
  file.path(demo_dir, "demo_patient_triggers.csv")
)

write_csv(
  demo_expected_guideline_actions,
  file.path(demo_dir, "demo_expected_guideline_actions.csv")
)
```

```{r show-triggers}
demo_patient_triggers %>%
  select(
    case_id, advanced_ckd, ckd_or_albuminuria, sglt2i_eligible_by_egfr,
    heart_failure, ascvd, obesity, mash_or_masld, severe_hyperglycemia,
    baseline_metformin, baseline_dpp4i, baseline_insulin,
    baseline_prandial_insulin, cost_barrier, injection_reluctance,
    gu_infection_caution, simple_regimen_preference
  ) %>%
  show_table("Patient triggers derived from structured variables")
```

```{r show-expected-actions}
demo_expected_guideline_actions %>%
  show_table("Expected guideline actions from the executable R decision model")
```

# Parse Existing Model Medication Outputs

The demo parser uses the `medication` column as the primary source of
active model actions.

```{r parser-functions, echo=TRUE}
split_medication_lines <- function(text) {
  text %>%
    str_replace_all("<br\\s*/?>", "\n") %>%
    str_split("\n") %>%
    pluck(1) %>%
    str_replace_all("^\\s*[•\\-*]+\\s*", "") %>%
    str_squish() %>%
    discard(~ .x == "")
}

detect_action <- function(line) {
  line_clean <- clean_text(line)
  case_when(
    str_detect(line_clean, "\\bstop\\b|\\bdiscontinue\\b") ~ "stop",
    str_detect(line_clean, "\\bstart\\b|\\badd\\b|\\binitiate\\b") ~ "start",
    str_detect(line_clean, "\\bcontinue\\b") ~ "continue",
    str_detect(line_clean, "\\bchange\\b|\\breduce\\b|\\bdecrease\\b") ~ "change",
    str_detect(line_clean, "\\bincrease\\b|\\btitrate\\b") ~ "increase_or_titrate",
    TRUE ~ "mentioned"
  )
}

detect_drug_class <- function(line) {
  line_clean <- clean_text(line)
  case_when(
    str_detect(line_clean, "metformin") ~ "metformin",
    str_detect(line_clean, "empagliflozin|dapagliflozin|canagliflozin|ertugliflozin") ~ "SGLT2 inhibitor",
    str_detect(line_clean, "semaglutide|liraglutide|dulaglutide|tirzepatide|ideglira|insulin degludec.*liraglutide") ~ "GLP-1 RA",
    str_detect(line_clean, "sitagliptin|linagliptin|saxagliptin|alogliptin") ~ "DPP-4 inhibitor",
    str_detect(line_clean, "insulin aspart|insulin lispro|regular insulin") ~ "prandial insulin",
    str_detect(line_clean, "insulin glargine|insulin degludec|basal insulin") ~ "insulin",
    TRUE ~ NA_character_
  )
}

parse_demo_model_actions <- function(responses) {
  responses %>%
    mutate(
      response_id = row_number(),
      model_clean = format_model(model),
      medication_lines = map(medication, split_medication_lines)
    ) %>%
    select(response_id, case_id, model_clean, medication, medication_lines) %>%
    unnest_longer(medication_lines, values_to = "model_text_excerpt") %>%
    mutate(
      drug_class = map_chr(model_text_excerpt, detect_drug_class),
      model_action = map_chr(model_text_excerpt, detect_action),
      model_active = model_action %in% c("start", "continue", "change", "increase_or_titrate"),
      model_stop_avoid = model_action == "stop"
    ) %>%
    filter(!is.na(drug_class)) %>%
    group_by(response_id, case_id, model_clean, drug_class) %>%
    summarize(
      model_action = paste(unique(model_action), collapse = "; "),
      model_active = any(model_active),
      model_stop_avoid = any(model_stop_avoid),
      model_text_excerpt = paste(unique(model_text_excerpt), collapse = " | "),
      .groups = "drop"
    )
}
```

```{r apply-parser}
demo_parsed_model_actions <- parse_demo_model_actions(demo_responses)

write_csv(
  demo_parsed_model_actions,
  file.path(demo_dir, "demo_parsed_model_actions.csv")
)
```

```{r show-parsed-actions}
demo_parsed_model_actions %>%
  select(
    case_id, model_clean, drug_class, model_action,
    model_active, model_stop_avoid, model_text_excerpt
  ) %>%
  arrange(case_id, model_clean, drug_class) %>%
  show_table("Parsed model actions from the medication column")
```

# Score Guideline Alignment

```{r scoring-functions, echo=TRUE}
score_one_row <- function(expected_action, rule_strength, model_action,
                          model_active, model_stop_avoid) {
  has_model_row <- !is.na(model_action)

  case_when(
    rule_strength %in% c("hard_avoid", "hard_avoid_if_combined") &
      has_model_row & model_stop_avoid ~ "adherent",
    rule_strength %in% c("hard_avoid", "hard_avoid_if_combined") &
      has_model_row & model_active ~ "potential_safety_discordance",
    rule_strength %in% c("hard_avoid", "hard_avoid_if_combined") &
      !has_model_row ~ "not_applicable",
    rule_strength == "strong_recommend" &
      has_model_row & (model_active | model_stop_avoid) &
      str_detect(expected_action, "caution|deintensify") ~ "caution_addressed",
    rule_strength == "strong_recommend" &
      has_model_row & model_active ~ "adherent",
    rule_strength == "strong_recommend" &
      !has_model_row ~ "potential_omission",
    rule_strength == "preference_sensitive" &
      has_model_row & (model_active | model_stop_avoid) ~ "guideline_supported",
    rule_strength == "preference_sensitive" &
      !has_model_row ~ "preference_sensitive_omission",
    rule_strength == "caution" &
      has_model_row & (model_active | model_stop_avoid) ~ "caution_addressed",
    rule_strength == "caution" &
      !has_model_row ~ "dose_caution_needs_review",
    TRUE ~ "not_applicable"
  )
}

score_guideline_alignment <- function(expected_actions, parsed_actions, responses) {
  response_keys <- responses %>%
    mutate(model_clean = format_model(model)) %>%
    distinct(case_id, model_clean)

  response_keys %>%
    left_join(expected_actions, by = "case_id", relationship = "many-to-many") %>%
    left_join(
      parsed_actions,
      by = c("case_id", "model_clean", "drug_class")
    ) %>%
    mutate(
      model_action = coalesce(model_action, NA_character_),
      model_active = coalesce(model_active, FALSE),
      model_stop_avoid = coalesce(model_stop_avoid, FALSE),
      score_label = pmap_chr(
        list(expected_action, rule_strength, model_action,
             model_active, model_stop_avoid),
        score_one_row
      )
    ) %>%
    arrange(case_id, model_clean, drug_class, rule_id)
}
```

```{r apply-scoring}
demo_guideline_adherence_scores_long <- score_guideline_alignment(
  demo_expected_guideline_actions,
  demo_parsed_model_actions,
  demo_responses
)

demo_case_model_summary <- demo_guideline_adherence_scores_long %>%
  count(case_id, model_clean, score_label, name = "n") %>%
  pivot_wider(
    names_from = score_label,
    values_from = n,
    values_fill = 0
  ) %>%
  arrange(case_id, model_clean)

write_csv(
  demo_guideline_adherence_scores_long,
  file.path(demo_dir, "demo_guideline_adherence_scores_long.csv")
)

write_csv(
  demo_case_model_summary,
  file.path(demo_dir, "demo_case_model_summary.csv")
)
```

```{r show-scoring}
demo_guideline_adherence_scores_long %>%
  select(
    case_id, model_clean, drug_class, rule_id, expected_action,
    rule_strength, model_action, score_label, model_text_excerpt
  ) %>%
  show_table("Drug-class-level guideline alignment scoring demo")
```

```{r show-summary}
demo_case_model_summary %>%
  show_table("Demo case-model score summary")
```

# Demo Interpretation

`DM020` triggers CKD/albuminuria and HF rules supporting SGLT2 inhibitor
use. All three models start empagliflozin, so the SGLT2 inhibitor rows
score as adherent. This case also triggers obesity/MASLD/MASH logic
supporting GLP-1 RA consideration. Claude Opus starts semaglutide, while
Claude Sonnet and GPT-5.4 omit GLP-1 RA; these omissions are labeled
preference-sensitive rather than hard violations.

`DM021` triggers advanced CKD and ASCVD rules. The decision model prefers
or considers GLP-1 RA in advanced CKD. Claude Sonnet starts semaglutide
and Claude Opus starts an insulin degludec/liraglutide combination, while
GPT-5.4 starts empagliflozin but does not include GLP-1 RA. The SGLT2
inhibitor recommendation is guideline-supported with kidney-function
caution, while GLP-1 RA omission is flagged for review. All three models
stop prandial insulin, which aligns with simplification and insulin
caution considerations in advanced CKD.

# Output Checks

```{r output-checks}
output_checks <- tibble(
  check = c(
    "demo input rows",
    "unique demo cases",
    "text chunks",
    "structured medication table rows",
    "rule registry rows",
    "expected action rows",
    "parsed model action rows",
    "scoring rows",
    "score labels present"
  ),
  value = c(
    nrow(demo_responses),
    n_distinct(demo_responses$case_id),
    nrow(ada_text_chunks),
    nrow(ada_medication_table_structured),
    nrow(ada_decision_rule_registry),
    nrow(demo_expected_guideline_actions),
    nrow(demo_parsed_model_actions),
    nrow(demo_guideline_adherence_scores_long),
    paste(sort(unique(demo_guideline_adherence_scores_long$score_label)), collapse = "; ")
  )
)
```

```{r show-output-checks}
show_table(output_checks, "Output checks")
```
