---
title: "Intra-Model Validation for Repeated Model Responses"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 6
)

library(tidyverse)
library(kableExtra)
library(scales)
library(stringr)
library(readr)
library(purrr)
library(flextable)
library(officer)
```

# Purpose

This workflow evaluates intra-model consistency using the current repeated
response dataset. It does not call any API and does not generate new model
responses. Each model is evaluated separately across five repeated responses
per case.

The primary stability metric is the case-level variation ratio, calculated
from a structured treatment signature derived from parsed medication-action
indicators.

```{r paths}
project_root <- normalizePath(file.path(getwd(), "../.."), mustWork = TRUE)

raw_path <- file.path(project_root, "data/raw/Capstone_repeated_data.csv")
output_dir <- file.path(project_root, "data/inner model validation/outputs")
export_dir <- file.path(project_root, "data/inner model validation/exports")

dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(export_dir, recursive = TRUE, showWarnings = FALSE)

raw_path
output_dir
export_dir
```

# Input Preparation

```{r input-preparation}
clean_column_names <- function(x) {
  x %>%
    str_replace_all("[^A-Za-z0-9]+", "_") %>%
    str_replace_all("^_|_$", "") %>%
    str_to_lower()
}

raw <- read_csv(raw_path, show_col_types = FALSE) %>%
  rename_with(clean_column_names)

required_cols <- c("case_id", "model", "treatment_plan", "vignette_text")
missing_cols <- setdiff(required_cols, names(raw))

if (length(missing_cols) > 0) {
  stop(
    "Missing required columns: ",
    paste(missing_cols, collapse = ", ")
  )
}

standardize_model_name <- function(x) {
  x_lower <- str_to_lower(as.character(x))

  case_when(
    str_detect(x_lower, "gpt") & str_detect(x_lower, "5\\.4") ~ "GPT-5.4",
    str_detect(x_lower, "opus") ~ "Claude Opus 4.6",
    str_detect(x_lower, "sonnet") ~ "Claude Sonnet 4.6",
    TRUE ~ as.character(x)
  )
}

standardized <- raw %>%
  mutate(
    case_id = str_squish(as.character(case_id)),
    model_raw = as.character(model),
    model_clean = standardize_model_name(model_raw),
    treatment_plan = replace_na(as.character(treatment_plan), ""),
    vignette_text = as.character(vignette_text)
  ) %>%
  arrange(case_id, model_clean) %>%
  group_by(case_id, model_clean) %>%
  mutate(repeat_id = row_number()) %>%
  ungroup() %>%
  mutate(
    response_id = paste0("resp_", str_pad(row_number(), width = 5, pad = "0"))
  ) %>%
  relocate(
    response_id,
    case_id,
    model_raw,
    model_clean,
    repeat_id,
    vignette_text,
    treatment_plan
  )

repeat_check <- standardized %>%
  count(model_clean, case_id, name = "n_repeats") %>%
  summarise(
    model_clean = unique(model_clean),
    n_cases = n(),
    min_repeats = min(n_repeats),
    max_repeats = max(n_repeats),
    all_have_5_repeats = all(n_repeats == 5),
    .by = model_clean
  )

duplicate_check <- standardized %>%
  count(case_id, model_clean, repeat_id, name = "n") %>%
  filter(n > 1)

if (nrow(duplicate_check) > 0) {
  stop("Duplicate case_id + model_clean + repeat_id rows detected.")
}

if (!all(repeat_check$all_have_5_repeats)) {
  stop("Not every case_id + model_clean combination has exactly 5 repeats.")
}

write_csv(
  standardized,
  file.path(output_dir, "intra_model_standardized_responses.csv")
)
```

```{r dataset-overview}
overview <- tibble(
  metric = c(
    "response_rows",
    "unique_cases",
    "unique_models",
    "expected_repeats_per_case_model",
    "blank_treatment_plans"
  ),
  value = c(
    nrow(standardized),
    n_distinct(standardized$case_id),
    n_distinct(standardized$model_clean),
    5,
    sum(standardized$treatment_plan == "")
  )
)

overview %>%
  kable(caption = "Dataset overview") %>%
  kable_styling(full_width = FALSE)

repeat_check %>%
  arrange(model_clean) %>%
  kable(caption = "Repeat-count validation by model") %>%
  kable_styling(full_width = FALSE)
```

# Treatment Plan Parser

```{r parser-functions}
drug_patterns <- tibble::tribble(
  ~drug_class, ~drug_regex,
  "metformin",
  "\\bmetformin\\b",
  "sglt2i",
  "\\b(sglt2|sglt-2|empagliflozin|dapagliflozin|canagliflozin|ertugliflozin|jardiance|farxiga|invokana)\\b",
  "glp1",
  "\\b(glp-1|glp1|semaglutide|liraglutide|dulaglutide|exenatide|ozempic|wegovy|trulicity|victoza|rybelsus)\\b",
  "gip_glp1",
  "\\b(tirzepatide|mounjaro|zepbound|dual gip|gip/glp-1|gip-glp-1)\\b",
  "dpp4i",
  "\\b(dpp-4|dpp4|sitagliptin|linagliptin|saxagliptin|alogliptin|januvia|tradjenta|onglyza)\\b",
  "insulin_any",
  "\\b(insulin|glargine|degludec|detemir|lispro|aspart|regular insulin|nph|basal|bolus)\\b",
  "basal_insulin",
  "\\b(basal insulin|insulin glargine|glargine|degludec|detemir|nph)\\b",
  "prandial_insulin",
  "\\b(prandial insulin|mealtime insulin|bolus insulin|lispro|aspart|glulisine|regular insulin|basal-bolus)\\b",
  "sulfonylurea",
  "\\b(sulfonylurea|glipizide|glyburide|glimepiride)\\b",
  "tzd",
  "\\b(tzd|thiazolidinedione|pioglitazone|rosiglitazone)\\b"
)

action_patterns <- list(
  start_add = "\\b(start|starts|starting|started|initiate|initiates|initiating|initiated|add|adds|adding|added|begin|begins|prescribe|prescribed|new)\\b",
  continue = "\\b(continue|continues|continued|continuing|maintain|maintains|maintained|keep|resume|remain on|remains on|remained on|remains appropriate)\\b",
  increase_intensify = "\\b(increase|increases|increased|increasing|intensify|intensifies|intensified|escalate|escalates|escalated|titrate up|uptitrate|raise|adjust upward|change)\\b",
  reduce = "\\b(reduce|reduces|reduced|reducing|decrease|decreases|decreased|lower|lowered|dose reduction|down-titrate|downtitrate|adjust downward)\\b",
  stop_or_avoid = "\\b(stop|stops|stopped|stopping|discontinue|discontinues|discontinued|discontinuing|hold|held|avoid|avoids|avoided|avoiding|do not use|do not recommend|not recommended|contraindicated|inappropriate|do not start|remain off|remains off|should remain off|left off|keep off|withhold|withheld)\\b",
  conditional = "\\b(consider|considered|may|could|if tolerated|if affordable|if egfr|depending on|provided that|unless|option|reasonable option|alternative)\\b",
  caution_monitor = "\\b(caution|monitor|monitoring|watch|risk|adverse effect|side effect|renal function|kidney function|volume depletion|hypoglycemia|genital infection|gu infection|dose adjust|dose adjustment)\\b",
  not_preferred = "\\b(preferred over|rather than|instead of|not preferred|less preferred|would not choose|avoid adding|do not add|not necessary|no need for|not needed|not indicated|no indication|defer|deferred|lower priority|not the best fit|poor fit|poor initial choice|without requiring)\\b"
)

insulin_false_positive_regex <- "\\b(insulin sensitivity|insulin resistance|insulin secretion|insulin production|endogenous insulin)\\b"

insulin_treatment_context_regex <- "\\b(start|started|continue|continued|stop|stopped|discontinue|discontinued|reduce|reduced|increase|increased|dose|units|basal|bolus|prandial|mealtime|glargine|degludec|detemir|lispro|aspart|nph|regular insulin|need for insulin|requiring insulin|requires insulin|insulin therapy)\\b"

split_pipe_row <- function(line) {
  line %>%
    str_remove("^\\s*\\|") %>%
    str_remove("\\|\\s*$") %>%
    str_split("\\|", simplify = FALSE) %>%
    .[[1]] %>%
    str_squish()
}

is_separator_row <- function(line) {
  str_detect(
    line,
    "^\\s*\\|?\\s*:?-{3,}:?\\s*(\\|\\s*:?-{3,}:?\\s*)+\\|?\\s*$"
  )
}

empty_table <- function() {
  tibble(
    medication = character(),
    action = character(),
    plan = character(),
    dose = character(),
    frequency = character(),
    unit_text = character()
  )
}

parse_markdown_med_table <- function(text) {
  if (is.na(text) || text == "") {
    return(empty_table())
  }

  lines <- str_split(text, "\\n")[[1]] %>%
    str_squish()
  lines <- lines[lines != ""]
  table_lines <- lines[str_detect(lines, fixed("|"))]

  if (length(table_lines) < 2) {
    return(empty_table())
  }

  header_idx <- which(
    str_detect(str_to_lower(table_lines), "\\b(medication|drug)\\b") &
      str_detect(str_to_lower(table_lines), "\\b(plan|action|recommendation)\\b")
  )

  if (length(header_idx) == 0) {
    return(empty_table())
  }

  header_idx <- header_idx[1]
  headers <- split_pipe_row(table_lines[header_idx])
  headers_clean <- str_to_lower(headers)

  med_col <- which(str_detect(headers_clean, "\\b(medication|drug)\\b"))[1]
  action_col <- which(str_detect(headers_clean, "\\b(action)\\b"))[1]
  plan_col <- which(str_detect(headers_clean, "\\b(plan|recommendation)\\b"))[1]
  dose_col <- which(str_detect(headers_clean, "\\b(dose|dosage)\\b"))[1]
  freq_col <- which(str_detect(headers_clean, "\\b(frequency|freq)\\b"))[1]

  if (is.na(med_col) || header_idx >= length(table_lines)) {
    return(empty_table())
  }

  data_lines <- table_lines[(header_idx + 1):length(table_lines)]
  data_lines <- data_lines[!is_separator_row(data_lines)]
  data_lines <- data_lines[str_detect(data_lines, fixed("|"))]

  map_dfr(data_lines, function(line) {
    cells <- split_pipe_row(line)

    get_cell <- function(idx) {
      if (is.na(idx) || idx > length(cells)) {
        return("")
      }
      cells[idx]
    }

    if (length(cells) < med_col || get_cell(med_col) == "") {
      return(NULL)
    }

    plan_text <- if (!is.na(plan_col)) {
      get_cell(plan_col)
    } else {
      paste(cells[-med_col], collapse = " | ")
    }

    tibble(
      medication = get_cell(med_col),
      action = get_cell(action_col),
      plan = plan_text,
      dose = get_cell(dose_col),
      frequency = get_cell(freq_col),
      unit_text = paste(cells, collapse = " | ")
    )
  })
}

split_text_units <- function(text, remove_table_lines = TRUE) {
  if (is.na(text) || text == "") {
    return(character())
  }

  lines <- str_split(text, "\\n")[[1]] %>%
    str_squish()

  if (remove_table_lines) {
    lines <- lines[
      !str_detect(lines, fixed("|")) &
        !is_separator_row(lines)
    ]
  }

  units <- str_split(
    paste(lines, collapse = " "),
    "(?<=[.!?])\\s+|;\\s+"
  )[[1]]

  units %>%
    str_squish() %>%
    purrr::discard(~ .x == "")
}

remove_insulin_false_positive_units <- function(units, drug_class) {
  if (length(units) == 0 || drug_class != "insulin_any") {
    return(units)
  }

  false_positive <- str_detect(
    units,
    regex(insulin_false_positive_regex, ignore_case = TRUE)
  )
  true_treatment_context <- str_detect(
    units,
    regex(insulin_treatment_context_regex, ignore_case = TRUE)
  )

  units[!(false_positive & !true_treatment_context)]
}

parse_one_drug_class <- function(text, drug_class, drug_regex) {
  med_pat <- regex(drug_regex, ignore_case = TRUE)
  table_rows <- parse_markdown_med_table(text)
  table_has_rows <- nrow(table_rows) > 0

  table_units <- table_rows %>%
    filter(
      str_detect(medication, med_pat) |
        str_detect(action, med_pat) |
        str_detect(plan, med_pat) |
        str_detect(dose, med_pat) |
        str_detect(frequency, med_pat) |
        str_detect(unit_text, med_pat)
    ) %>%
    pull(unit_text)

  if (length(table_units) > 0) {
    units <- table_units
    source_used <- "table"
  } else {
    free_units <- split_text_units(text, remove_table_lines = TRUE)
    units <- free_units[str_detect(free_units, med_pat)]
    source_used <- if_else(
      table_has_rows,
      "free_text_rationale",
      "free_text_no_table"
    )
  }

  units <- remove_insulin_false_positive_units(units, drug_class)
  mentioned <- length(units) > 0

  if (!mentioned) {
    return(tibble(
      drug_class = drug_class,
      mentioned = 0L,
      active_recommendation = 0L,
      start_add = 0L,
      continue = 0L,
      increase_intensify = 0L,
      reduce = 0L,
      stop_or_avoid = 0L,
      conditional = 0L,
      caution_monitor = 0L,
      not_preferred = 0L,
      unclear = 0L,
      manual_review_flag = 0L,
      component_mixed_action = 0L,
      example_text = NA_character_,
      source_used = NA_character_,
      table_has_rows = as.integer(table_has_rows)
    ))
  }

  detect_action <- function(pattern) {
    any(str_detect(units, regex(pattern, ignore_case = TRUE)))
  }

  table_start <- any(str_detect(units, regex("\\bSTART\\b", ignore_case = TRUE)))
  table_continue <- any(str_detect(units, regex("\\bCONTINUE\\b", ignore_case = TRUE)))
  table_change <- any(str_detect(units, regex("\\b(CHANGE|SWITCH|MODIFY|ADJUST)\\b", ignore_case = TRUE)))
  table_stop <- any(str_detect(units, regex("\\bSTOP\\b", ignore_case = TRUE)))
  table_reduce <- any(str_detect(units, regex("\\b(REDUCE|DECREASE|LOWER)\\b", ignore_case = TRUE)))

  start_add <- detect_action(action_patterns$start_add) | table_start
  continue <- detect_action(action_patterns$continue) | table_continue
  increase_intensify <- detect_action(action_patterns$increase_intensify)
  reduce <- detect_action(action_patterns$reduce) | table_reduce
  stop_or_avoid <- detect_action(action_patterns$stop_or_avoid) | table_stop
  conditional <- detect_action(action_patterns$conditional)
  caution_monitor <- detect_action(action_patterns$caution_monitor)
  not_preferred <- detect_action(action_patterns$not_preferred)

  table_active_signal <- table_start |
    table_continue |
    table_change |
    increase_intensify |
    (start_add & !stop_or_avoid)

  free_text_active_signal <- (start_add | continue | increase_intensify) &
    !stop_or_avoid &
    !not_preferred

  active_recommendation <- case_when(
    source_used == "table" ~ table_active_signal,
    source_used == "free_text_no_table" ~ free_text_active_signal,
    source_used == "free_text_rationale" ~ FALSE,
    TRUE ~ FALSE
  )

  component_mixed_action <- as.integer(
    source_used == "table" &&
      active_recommendation &&
      stop_or_avoid
  )

  unclear <- mentioned &&
    !active_recommendation &&
    !stop_or_avoid &&
    !conditional &&
    !caution_monitor &&
    !not_preferred

  manual_review_flag <- unclear |
    ((active_recommendation & stop_or_avoid) & component_mixed_action == 0L) |
    ((stop_or_avoid & (start_add | continue | increase_intensify)) &
      component_mixed_action == 0L)

  tibble(
    drug_class = drug_class,
    mentioned = as.integer(mentioned),
    active_recommendation = as.integer(active_recommendation),
    start_add = as.integer(start_add),
    continue = as.integer(continue),
    increase_intensify = as.integer(increase_intensify),
    reduce = as.integer(reduce),
    stop_or_avoid = as.integer(stop_or_avoid),
    conditional = as.integer(conditional),
    caution_monitor = as.integer(caution_monitor),
    not_preferred = as.integer(not_preferred),
    unclear = as.integer(unclear),
    manual_review_flag = as.integer(manual_review_flag),
    component_mixed_action = as.integer(component_mixed_action),
    example_text = paste(head(units, 2), collapse = " || "),
    source_used = source_used,
    table_has_rows = as.integer(table_has_rows)
  )
}

parse_one_response <- function(text) {
  out <- map2_dfr(
    drug_patterns$drug_class,
    drug_patterns$drug_regex,
    ~ parse_one_drug_class(text, .x, .y)
  )

  get_max_indicator <- function(df, target_class, target_col) {
    vals <- df %>%
      filter(drug_class == target_class) %>%
      pull({{ target_col }})

    if (length(vals) == 0 || all(is.na(vals))) {
      return(0L)
    }

    as.integer(max(vals, na.rm = TRUE))
  }

  basal_active <- get_max_indicator(out, "basal_insulin", active_recommendation)
  prandial_active <- get_max_indicator(out, "prandial_insulin", active_recommendation)
  basal_stop <- get_max_indicator(out, "basal_insulin", stop_or_avoid)
  prandial_stop <- get_max_indicator(out, "prandial_insulin", stop_or_avoid)

  any_insulin_component_active <- basal_active == 1L | prandial_active == 1L
  any_insulin_component_stop <- basal_stop == 1L | prandial_stop == 1L
  mixed_insulin_components <- any_insulin_component_active &
    any_insulin_component_stop

  out %>%
    mutate(
      active_recommendation = if_else(
        drug_class == "insulin_any" & any_insulin_component_active,
        1L,
        active_recommendation
      ),
      mentioned = if_else(
        drug_class == "insulin_any" &
          (any_insulin_component_active | any_insulin_component_stop),
        1L,
        mentioned
      ),
      stop_or_avoid = if_else(
        drug_class == "insulin_any" & any_insulin_component_stop,
        1L,
        stop_or_avoid
      ),
      component_mixed_action = if_else(
        drug_class == "insulin_any" & mixed_insulin_components,
        1L,
        component_mixed_action
      ),
      manual_review_flag = if_else(
        drug_class == "insulin_any" & mixed_insulin_components,
        0L,
        manual_review_flag
      )
    )
}
```

# Parse Repeated Responses

```{r parse-responses}
parsed_long <- standardized %>%
  select(
    response_id,
    case_id,
    model_raw,
    model_clean,
    repeat_id,
    treatment_plan
  ) %>%
  mutate(parsed = map(treatment_plan, parse_one_response)) %>%
  unnest(parsed)

write_csv(
  parsed_long,
  file.path(output_dir, "intra_model_parsed_long.csv")
)

parsed_wide <- parsed_long %>%
  transmute(
    response_id,
    drug_class,
    mentioned,
    active = active_recommendation,
    start_add,
    continue,
    increase_intensify,
    reduce,
    stop_avoid = stop_or_avoid,
    conditional,
    caution_monitor,
    not_preferred,
    manual_review = manual_review_flag,
    component_mixed_action
  ) %>%
  pivot_wider(
    id_cols = response_id,
    names_from = drug_class,
    values_from = c(
      mentioned,
      active,
      start_add,
      continue,
      increase_intensify,
      reduce,
      stop_avoid,
      conditional,
      caution_monitor,
      not_preferred,
      manual_review,
      component_mixed_action
    ),
    names_glue = "{drug_class}_{.value}",
    values_fill = 0L
  ) %>%
  left_join(standardized, by = "response_id") %>%
  relocate(
    response_id,
    case_id,
    model_raw,
    model_clean,
    repeat_id,
    treatment_plan
  )

write_csv(
  parsed_wide,
  file.path(output_dir, "intra_model_parsed_wide.csv")
)

parser_qc <- tibble(
  metric = c(
    "parsed_long_rows",
    "parsed_wide_rows",
    "drug_classes",
    "expected_parsed_long_rows",
    "manual_review_flags"
  ),
  value = c(
    nrow(parsed_long),
    nrow(parsed_wide),
    n_distinct(parsed_long$drug_class),
    nrow(standardized) * nrow(drug_patterns),
    sum(parsed_long$manual_review_flag, na.rm = TRUE)
  )
)

parser_qc %>%
  kable(caption = "Parser QC") %>%
  kable_styling(full_width = FALSE)
```

# Structured Treatment Signatures

```{r treatment-signatures}
signature_cols <- c(
  "mentioned",
  "active_recommendation",
  "start_add",
  "continue",
  "increase_intensify",
  "reduce",
  "stop_or_avoid",
  "conditional",
  "not_preferred",
  "manual_review_flag"
)

treatment_signatures <- parsed_long %>%
  mutate(
    signature_piece = str_c(
      drug_class,
      mentioned,
      active_recommendation,
      start_add,
      continue,
      increase_intensify,
      reduce,
      stop_or_avoid,
      conditional,
      not_preferred,
      manual_review_flag,
      sep = ":"
    )
  ) %>%
  arrange(response_id, drug_class) %>%
  summarise(
    treatment_signature = paste(signature_piece, collapse = "|"),
    n_drug_classes_mentioned = sum(mentioned, na.rm = TRUE),
    n_active_classes = sum(active_recommendation, na.rm = TRUE),
    n_stop_avoid_classes = sum(stop_or_avoid, na.rm = TRUE),
    n_manual_review_classes = sum(manual_review_flag, na.rm = TRUE),
    .by = c(response_id, case_id, model_clean, repeat_id)
  )

signature_counts <- treatment_signatures %>%
  count(model_clean, case_id, treatment_signature, name = "signature_count") %>%
  arrange(model_clean, case_id, desc(signature_count))

case_stability <- signature_counts %>%
  summarise(
    n_repeats = sum(signature_count),
    n_unique_signatures = n(),
    modal_signature_count = max(signature_count),
    variation_ratio = 1 - modal_signature_count / n_repeats,
    exact_match = as.integer(n_unique_signatures == 1),
    .by = c(model_clean, case_id)
  ) %>%
  arrange(model_clean, desc(variation_ratio), case_id)

model_summary <- case_stability %>%
  summarise(
    n_cases = n(),
    mean_variation_ratio = mean(variation_ratio),
    median_variation_ratio = median(variation_ratio),
    exact_match_cases = sum(exact_match),
    exact_match_rate = mean(exact_match),
    cases_with_any_variation = sum(variation_ratio > 0),
    max_variation_ratio = max(variation_ratio),
    .by = model_clean
  ) %>%
  arrange(mean_variation_ratio)

variation_ratio_distribution <- case_stability %>%
  count(model_clean, variation_ratio, name = "n_cases") %>%
  group_by(model_clean) %>%
  mutate(percent_cases = n_cases / sum(n_cases)) %>%
  ungroup() %>%
  arrange(model_clean, variation_ratio)

top_unstable_cases <- case_stability %>%
  group_by(model_clean) %>%
  slice_max(
    order_by = variation_ratio,
    n = 10,
    with_ties = FALSE
  ) %>%
  ungroup() %>%
  left_join(
    standardized %>%
      select(case_id, vignette_text) %>%
      distinct(),
    by = "case_id"
  )

write_csv(
  treatment_signatures,
  file.path(output_dir, "intra_model_treatment_signatures.csv")
)
write_csv(
  signature_counts,
  file.path(output_dir, "intra_model_signature_counts.csv")
)
write_csv(
  case_stability,
  file.path(output_dir, "intra_model_case_stability.csv")
)
write_csv(
  model_summary,
  file.path(output_dir, "intra_model_model_summary.csv")
)
write_csv(
  variation_ratio_distribution,
  file.path(output_dir, "intra_model_variation_ratio_distribution.csv")
)
write_csv(
  top_unstable_cases,
  file.path(output_dir, "intra_model_top_unstable_cases.csv")
)
```

```{r stability-tables}
model_summary %>%
  mutate(
    mean_variation_ratio = round(mean_variation_ratio, 3),
    median_variation_ratio = round(median_variation_ratio, 3),
    exact_match_rate = percent(exact_match_rate, accuracy = 0.1)
  ) %>%
  kable(caption = "Model-level intra-model stability summary") %>%
  kable_styling(full_width = FALSE)

variation_ratio_distribution %>%
  mutate(percent_cases = percent(percent_cases, accuracy = 0.1)) %>%
  kable(caption = "Variation ratio distribution by model") %>%
  kable_styling(full_width = FALSE)
```

# Krippendorff's Alpha

Krippendorff's alpha is used here as a formal agreement statistic for repeated
model outputs. For each model, the five repeated responses are treated like five
raters assigning a structured treatment signature to each case. Because the
signature is a categorical plan identity, this workflow uses nominal alpha.

```{r krippendorff-alpha}
krippendorff_alpha_nominal <- function(x) {
  x <- as.matrix(x)

  unit_counts <- apply(x, 1, function(row_values) {
    row_values <- row_values[!is.na(row_values) & row_values != ""]

    if (length(row_values) < 2) {
      return(NA_real_)
    }

    tab <- table(row_values)
    n_unit <- sum(tab)

    sum(tab * (n_unit - tab)) / (n_unit * (n_unit - 1))
  })

  observed_disagreement <- mean(unit_counts, na.rm = TRUE)

  all_values <- as.vector(x)
  all_values <- all_values[!is.na(all_values) & all_values != ""]
  pooled_tab <- table(all_values)
  pooled_n <- sum(pooled_tab)

  expected_disagreement <- sum(pooled_tab * (pooled_n - pooled_tab)) /
    (pooled_n * (pooled_n - 1))

  if (is.na(expected_disagreement) || expected_disagreement == 0) {
    return(NA_real_)
  }

  1 - observed_disagreement / expected_disagreement
}

alpha_by_model <- treatment_signatures %>%
  select(model_clean, case_id, repeat_id, treatment_signature) %>%
  mutate(repeat_id = paste0("repeat_", repeat_id)) %>%
  pivot_wider(
    names_from = repeat_id,
    values_from = treatment_signature
  ) %>%
  group_by(model_clean) %>%
  group_modify(~ {
    signature_matrix <- .x %>%
      select(starts_with("repeat_")) %>%
      as.matrix()

    tibble(
      krippendorff_alpha = krippendorff_alpha_nominal(signature_matrix),
      n_cases = nrow(.x),
      n_repeats = ncol(signature_matrix)
    )
  }) %>%
  ungroup() %>%
  mutate(
    interpretation = case_when(
      is.na(krippendorff_alpha) ~ "Not estimable",
      krippendorff_alpha >= 0.80 ~ "Strong agreement",
      krippendorff_alpha >= 0.67 ~ "Tentative agreement",
      krippendorff_alpha > 0 ~ "Low agreement",
      TRUE ~ "No better than chance"
    )
  ) %>%
  arrange(desc(krippendorff_alpha))

model_summary <- model_summary %>%
  left_join(
    alpha_by_model %>%
      select(model_clean, krippendorff_alpha, alpha_interpretation = interpretation),
    by = "model_clean"
  )

write_csv(
  alpha_by_model,
  file.path(output_dir, "intra_model_krippendorff_alpha.csv")
)
write_csv(
  model_summary,
  file.path(output_dir, "intra_model_model_summary.csv")
)

alpha_by_model %>%
  mutate(
    krippendorff_alpha = round(krippendorff_alpha, 3)
  ) %>%
  kable(caption = "Krippendorff's alpha for structured treatment signatures") %>%
  kable_styling(full_width = FALSE)
```

Higher Krippendorff's alpha indicates that the model repeatedly assigned the
same structured treatment signature to the same case more often than expected
by chance. Values near 1 indicate strong intra-model consistency. Values near 0
indicate agreement no better than chance after accounting for the overall
distribution of treatment signatures.

# Alpha Sensitivity Analysis

The full treatment signature is intentionally strict because it includes
rationale-level mentions, stop/avoid language, conditional alternatives, and
manual-review flags. To separate core prescribing consistency from broader
response-style consistency, this section recalculates Krippendorff's alpha using
progressively stricter signature definitions.

```{r alpha-sensitivity}
make_response_signature <- function(data, signature_variables) {
  data %>%
    mutate(
      signature_piece = pmap_chr(
        select(., drug_class, all_of(signature_variables)),
        ~ paste(c(...), collapse = ":")
      )
    ) %>%
    arrange(response_id, drug_class) %>%
    summarise(
      treatment_signature = paste(signature_piece, collapse = "|"),
      .by = c(model_clean, case_id, repeat_id, response_id)
    )
}

calculate_alpha_for_signature <- function(signature_data) {
  signature_data %>%
    select(model_clean, case_id, repeat_id, treatment_signature) %>%
    mutate(repeat_id = paste0("repeat_", repeat_id)) %>%
    pivot_wider(
      names_from = repeat_id,
      values_from = treatment_signature
    ) %>%
    group_by(model_clean) %>%
    group_modify(~ {
      signature_matrix <- .x %>%
        select(starts_with("repeat_")) %>%
        as.matrix()

      tibble(alpha = krippendorff_alpha_nominal(signature_matrix))
    }) %>%
    ungroup()
}

alpha_sensitivity_specs <- list(
  "Active only alpha" = c("active_recommendation"),
  "Active + stop/avoid alpha" = c(
    "active_recommendation",
    "stop_or_avoid"
  ),
  "Full signature alpha" = signature_cols
)

alpha_sensitivity_long <- imap_dfr(
  alpha_sensitivity_specs,
  function(signature_variables, metric_definition) {
    make_response_signature(parsed_long, signature_variables) %>%
      calculate_alpha_for_signature() %>%
      mutate(metric_definition = metric_definition)
  }
) %>%
  mutate(
    metric_definition = factor(
      metric_definition,
      levels = names(alpha_sensitivity_specs)
    )
  ) %>%
  arrange(metric_definition, model_clean)

alpha_sensitivity_wide <- alpha_sensitivity_long %>%
  mutate(
    model_label = case_when(
      model_clean == "Claude Opus 4.6" ~ "Claude Opus",
      model_clean == "Claude Sonnet 4.6" ~ "Claude Sonnet",
      model_clean == "GPT-5.4" ~ "GPT-5.4",
      TRUE ~ model_clean
    ),
    alpha = round(alpha, 3)
  ) %>%
  select(metric_definition, model_label, alpha) %>%
  pivot_wider(
    names_from = model_label,
    values_from = alpha
  ) %>%
  arrange(metric_definition)

write_csv(
  alpha_sensitivity_long,
  file.path(output_dir, "intra_model_alpha_sensitivity_long.csv")
)
write_csv(
  alpha_sensitivity_wide,
  file.path(output_dir, "intra_model_alpha_sensitivity_table.csv")
)

alpha_sensitivity_wide %>%
  kable(caption = "Krippendorff's alpha sensitivity analysis") %>%
  kable_styling(full_width = FALSE)
```

The sensitivity results help interpret the low full-signature alpha values.
Active medication recommendations are substantially more stable than the full
signature, indicating that much of the strict-signature variation comes from
secondary language about stopping, avoiding, considering, or mentioning drugs
in the rationale.

# Medication Column Dose-Level Consistency

The `medication` column is a more compact representation of the final treatment
plan than the full `treatment_plan` text. This section parses medication names,
actions, and dose/frequency text from that column to estimate consistency in
the final prescription-style output.

```{r medication-column-consistency}
medication_name_patterns <- tibble::tribble(
  ~medication_name, ~medication_regex,
  "metformin", "\\bmetformin\\b",
  "sitagliptin", "\\bsitagliptin\\b",
  "linagliptin", "\\blinagliptin\\b",
  "empagliflozin", "\\bempagliflozin\\b",
  "dapagliflozin", "\\bdapagliflozin\\b",
  "canagliflozin", "\\bcanagliflozin\\b",
  "semaglutide", "\\bsemaglutide\\b",
  "liraglutide", "\\bliraglutide\\b",
  "dulaglutide", "\\bdulaglutide\\b",
  "tirzepatide", "\\btirzepatide\\b",
  "insulin glargine", "\\binsulin glargine\\b|\\bglargine\\b",
  "insulin aspart", "\\binsulin aspart\\b|\\baspart\\b",
  "insulin lispro", "\\binsulin lispro\\b|\\blispro\\b",
  "insulin degludec", "\\binsulin degludec\\b|\\bdegludec\\b",
  "nph insulin", "\\bnph\\b",
  "glipizide", "\\bglipizide\\b",
  "glimepiride", "\\bglimepiride\\b",
  "glyburide", "\\bglyburide\\b",
  "pioglitazone", "\\bpioglitazone\\b"
)

normalize_medication_text <- function(x) {
  x %>%
    replace_na("") %>%
    str_replace_all("<br\\s*/?>", "\n") %>%
    str_replace_all("&nbsp;", " ") %>%
    str_replace_all("[\u2022]", "\n") %>%
    str_replace_all("\\r", "\n")
}

split_medication_lines <- function(x) {
  normalize_medication_text(x) %>%
    str_split("\\n") %>%
    .[[1]] %>%
    str_squish() %>%
    purrr::discard(~ .x == "")
}

detect_medication_action <- function(line) {
  line_lower <- str_to_lower(line)

  case_when(
    str_detect(line_lower, "\\b(stop|discontinue|hold|avoid)\\b") ~ "stop",
    str_detect(line_lower, "\\b(continue|maintain|keep)\\b") ~ "continue",
    str_detect(line_lower, "\\b(start|add|initiate|begin|new)\\b") ~ "start",
    str_detect(line_lower, "\\b(change|switch|modify|adjust)\\b") ~ "change",
    str_detect(line_lower, "\\b(reduce|decrease|lower|down-titrate|downtitrate)\\b") ~ "reduce",
    str_detect(line_lower, "\\b(increase|raise|intensify|titrate up|uptitrate)\\b") ~ "increase",
    TRUE ~ "unknown"
  )
}

extract_medication_name <- function(line) {
  matched <- medication_name_patterns %>%
    filter(str_detect(line, regex(medication_regex, ignore_case = TRUE))) %>%
    pull(medication_name)

  if (length(matched) > 0) {
    return(matched[1])
  }

  line %>%
    str_to_lower() %>%
    str_remove("^\\s*(start|add|initiate|begin|new|continue|maintain|keep|stop|discontinue|hold|avoid|change|switch|modify|adjust|reduce|decrease|lower|increase|raise|intensify)\\s+") %>%
    str_remove("\\s+(to|at)?\\s*\\d.*$") %>%
    str_squish()
}

normalize_dose_text <- function(line, medication_name, action) {
  if (action == "stop") {
    return("")
  }

  dose <- line %>%
    str_to_lower() %>%
    str_remove("^\\s*(start|add|initiate|begin|new|continue|maintain|keep|change|switch|modify|adjust|reduce|decrease|lower|increase|raise|intensify)\\s+") %>%
    str_remove(regex(medication_name, ignore_case = TRUE)) %>%
    str_remove("^\\s*(to|at|-|:)\\s*") %>%
    str_replace_all("\\bby mouth\\b|\\bpo\\b", "") %>%
    str_replace_all("\\bsubcutaneous injection\\b|\\bsubcutaneously\\b|\\bsubcutaneous\\b", "sq") %>%
    str_replace_all("\\bonce daily\\b", "daily") %>%
    str_replace_all("\\btwice daily\\b", "bid") %>%
    str_replace_all("\\bthree times daily\\b", "tid") %>%
    str_replace_all("\\bonce weekly\\b", "weekly") %>%
    str_replace_all("\\s+", " ") %>%
    str_squish()

  dose
}

parse_medication_column_one_response <- function(medication_text) {
  lines <- split_medication_lines(medication_text)

  if (length(lines) == 0) {
    return(tibble(
      medication_name = character(),
      medication_action = character(),
      dose_frequency = character(),
      source_line = character()
    ))
  }

  map_dfr(lines, function(line) {
    action <- detect_medication_action(line)
    med_name <- extract_medication_name(line)
    dose_text <- normalize_dose_text(line, med_name, action)

    tibble(
      medication_name = med_name,
      medication_action = action,
      dose_frequency = dose_text,
      source_line = line
    )
  }) %>%
    filter(!is.na(medication_name), medication_name != "")
}

medication_parsed_long <- standardized %>%
  select(response_id, case_id, model_clean, repeat_id, medication) %>%
  mutate(parsed_medication = map(medication, parse_medication_column_one_response)) %>%
  unnest(parsed_medication) %>%
  mutate(
    medication_name = str_squish(str_to_lower(medication_name)),
    medication_action = str_squish(str_to_lower(medication_action)),
    dose_frequency = str_squish(str_to_lower(dose_frequency))
  ) %>%
  arrange(model_clean, case_id, repeat_id, medication_name)

make_medication_column_signature <- function(data, signature_level) {
  data %>%
    mutate(
      signature_piece = case_when(
        signature_level == "Medication-name alpha" ~ medication_name,
        signature_level == "Medication-action alpha" ~ str_c(
          medication_action,
          medication_name,
          sep = ":"
        ),
        signature_level == "Medication-action-dose alpha" ~ str_c(
          medication_action,
          medication_name,
          dose_frequency,
          sep = ":"
        ),
        TRUE ~ medication_name
      )
    ) %>%
    arrange(response_id, signature_piece) %>%
    summarise(
      medication_signature = paste(unique(signature_piece), collapse = "|"),
      n_medications = n_distinct(medication_name),
      .by = c(model_clean, case_id, repeat_id, response_id)
    )
}

medication_alpha_specs <- c(
  "Medication-name alpha",
  "Medication-action alpha",
  "Medication-action-dose alpha"
)

medication_alpha_long <- map_dfr(
  medication_alpha_specs,
  function(metric_definition) {
    make_medication_column_signature(
      medication_parsed_long,
      metric_definition
    ) %>%
      select(model_clean, case_id, repeat_id, medication_signature) %>%
      mutate(repeat_id = paste0("repeat_", repeat_id)) %>%
      pivot_wider(
        names_from = repeat_id,
        values_from = medication_signature
      ) %>%
      group_by(model_clean) %>%
      group_modify(~ {
        signature_matrix <- .x %>%
          select(starts_with("repeat_")) %>%
          as.matrix()

        tibble(alpha = krippendorff_alpha_nominal(signature_matrix))
      }) %>%
      ungroup() %>%
      mutate(metric_definition = metric_definition)
  }
) %>%
  mutate(
    metric_definition = factor(
      metric_definition,
      levels = medication_alpha_specs
    )
  ) %>%
  arrange(metric_definition, model_clean)

medication_alpha_wide <- medication_alpha_long %>%
  mutate(
    model_label = case_when(
      model_clean == "Claude Opus 4.6" ~ "Claude Opus",
      model_clean == "Claude Sonnet 4.6" ~ "Claude Sonnet",
      model_clean == "GPT-5.4" ~ "GPT-5.4",
      TRUE ~ model_clean
    ),
    alpha = round(alpha, 3)
  ) %>%
  select(metric_definition, model_label, alpha) %>%
  pivot_wider(
    names_from = model_label,
    values_from = alpha
  ) %>%
  arrange(metric_definition)

medication_dose_signatures <- make_medication_column_signature(
  medication_parsed_long,
  "Medication-action-dose alpha"
)

medication_dose_case_stability <- medication_dose_signatures %>%
  count(model_clean, case_id, medication_signature, name = "signature_count") %>%
  summarise(
    n_repeats = sum(signature_count),
    n_unique_medication_signatures = n(),
    modal_signature_count = max(signature_count),
    medication_dose_variation_ratio = 1 - modal_signature_count / n_repeats,
    .by = c(model_clean, case_id)
  ) %>%
  arrange(model_clean, desc(medication_dose_variation_ratio), case_id)

top_medication_dose_disagreement_cases <-
  medication_dose_case_stability %>%
  group_by(model_clean) %>%
  slice_max(
    order_by = medication_dose_variation_ratio,
    n = 10,
    with_ties = FALSE
  ) %>%
  ungroup()

write_csv(
  medication_parsed_long,
  file.path(output_dir, "intra_model_medication_column_parsed_long.csv")
)
write_csv(
  medication_alpha_long,
  file.path(output_dir, "intra_model_medication_column_alpha_long.csv")
)
write_csv(
  medication_alpha_wide,
  file.path(output_dir, "intra_model_medication_column_alpha_table.csv")
)
write_csv(
  medication_dose_case_stability,
  file.path(output_dir, "intra_model_medication_dose_case_stability.csv")
)
write_csv(
  top_medication_dose_disagreement_cases,
  file.path(output_dir, "intra_model_top_medication_dose_disagreement_cases.csv")
)

medication_alpha_wide %>%
  kable(caption = "Medication-column Krippendorff's alpha") %>%
  kable_styling(full_width = FALSE)

top_medication_dose_disagreement_cases %>%
  kable(caption = "Top medication action-dose disagreement cases") %>%
  kable_styling(full_width = FALSE)
```

Medication-column alpha is a dose-level sensitivity analysis focused on the
final prescription-style output. Medication-name alpha evaluates whether the
same drugs are listed; medication-action alpha adds start/stop/continue/change
decisions; medication-action-dose alpha additionally includes normalized dose
and frequency text.

# Secondary Agreement Metrics

```{r secondary-metrics}
medication_class_agreement <- parsed_long %>%
  summarise(
    repeat_count = n(),
    mentioned_agreement = as.integer(n_distinct(mentioned) == 1),
    active_agreement = as.integer(n_distinct(active_recommendation) == 1),
    stop_avoid_agreement = as.integer(n_distinct(stop_or_avoid) == 1),
    any_mentioned = max(mentioned, na.rm = TRUE),
    any_active = max(active_recommendation, na.rm = TRUE),
    any_stop_avoid = max(stop_or_avoid, na.rm = TRUE),
    .by = c(model_clean, case_id, drug_class)
  ) %>%
  summarise(
    n_case_drug_pairs = n(),
    mentioned_agreement_rate = mean(mentioned_agreement),
    active_agreement_rate = mean(active_agreement),
    stop_avoid_agreement_rate = mean(stop_avoid_agreement),
    n_pairs_ever_mentioned = sum(any_mentioned),
    n_pairs_ever_active = sum(any_active),
    n_pairs_ever_stop_avoid = sum(any_stop_avoid),
    .by = c(model_clean, drug_class)
  ) %>%
  arrange(model_clean, drug_class)

action_level_agreement <- parsed_long %>%
  pivot_longer(
    cols = all_of(signature_cols),
    names_to = "action_indicator",
    values_to = "value"
  ) %>%
  summarise(
    agreement = as.integer(n_distinct(value) == 1),
    ever_present = max(value, na.rm = TRUE),
    .by = c(model_clean, case_id, drug_class, action_indicator)
  ) %>%
  summarise(
    n_case_drug_action_pairs = n(),
    agreement_rate = mean(agreement),
    n_pairs_ever_present = sum(ever_present),
    .by = c(model_clean, action_indicator)
  ) %>%
  arrange(model_clean, action_indicator)

manual_review_burden <- parsed_long %>%
  summarise(
    manual_review_classes = sum(manual_review_flag, na.rm = TRUE),
    any_manual_review = as.integer(any(manual_review_flag == 1)),
    .by = c(model_clean, response_id, case_id, repeat_id)
  ) %>%
  summarise(
    n_responses = n(),
    responses_with_manual_review = sum(any_manual_review),
    response_manual_review_rate = mean(any_manual_review),
    total_manual_review_classes = sum(manual_review_classes),
    .by = model_clean
  ) %>%
  arrange(desc(response_manual_review_rate))

write_csv(
  medication_class_agreement,
  file.path(output_dir, "intra_model_medication_class_agreement.csv")
)
write_csv(
  action_level_agreement,
  file.path(output_dir, "intra_model_action_level_agreement.csv")
)
write_csv(
  manual_review_burden,
  file.path(output_dir, "intra_model_manual_review_burden.csv")
)

medication_class_agreement %>%
  mutate(
    mentioned_agreement_rate = percent(mentioned_agreement_rate, accuracy = 0.1),
    active_agreement_rate = percent(active_agreement_rate, accuracy = 0.1),
    stop_avoid_agreement_rate = percent(stop_avoid_agreement_rate, accuracy = 0.1)
  ) %>%
  kable(caption = "Medication-class agreement by model") %>%
  kable_styling(full_width = FALSE)

manual_review_burden %>%
  mutate(
    response_manual_review_rate = percent(
      response_manual_review_rate,
      accuracy = 0.1
    )
  ) %>%
  kable(caption = "Manual-review burden by model") %>%
  kable_styling(full_width = FALSE)
```

# Targeted Review of Unstable and Parser-Flagged Cases

This section focuses on the strictest instability pattern: cases with
`variation_ratio == 0.8`, where all five repeated responses produced different
full treatment signatures. Within those cases, we specifically inspect parser
flags from `free_text_rationale`, because these flags usually come from
rationale-level mentions rather than the medication table itself.

```{r targeted-review}
max_unstable_cases <- case_stability %>%
  filter(variation_ratio == 0.8)

max_unstable_case_counts <- max_unstable_cases %>%
  count(model_clean, name = "n_max_unstable_cases") %>%
  arrange(desc(n_max_unstable_cases))

parsed_with_unstable_flag <- parsed_long %>%
  left_join(
    max_unstable_cases %>%
      select(model_clean, case_id, variation_ratio),
    by = c("model_clean", "case_id")
  ) %>%
  mutate(
    is_max_unstable_case = !is.na(variation_ratio)
  )

targeted_review_rows <- parsed_with_unstable_flag %>%
  filter(
    is_max_unstable_case,
    manual_review_flag == 1,
    source_used == "free_text_rationale"
  ) %>%
  select(
    model_clean,
    case_id,
    repeat_id,
    response_id,
    drug_class,
    source_used,
    mentioned,
    active_recommendation,
    stop_or_avoid,
    conditional,
    not_preferred,
    unclear,
    manual_review_flag,
    example_text
  ) %>%
  arrange(model_clean, case_id, repeat_id, drug_class)

manual_review_rate_by_instability <- parsed_with_unstable_flag %>%
  summarise(
    any_manual_review = as.integer(any(manual_review_flag == 1)),
    .by = c(model_clean, case_id, repeat_id, is_max_unstable_case)
  ) %>%
  summarise(
    n_responses = n(),
    responses_with_manual_review = sum(any_manual_review),
    manual_review_response_rate = mean(any_manual_review),
    .by = c(model_clean, is_max_unstable_case)
  ) %>%
  arrange(model_clean, desc(is_max_unstable_case))

manual_review_by_model_drug <- parsed_long %>%
  summarise(
    manual_review_flags = sum(manual_review_flag, na.rm = TRUE),
    .by = c(model_clean, drug_class)
  ) %>%
  arrange(model_clean, desc(manual_review_flags))

manual_review_by_model_source <- parsed_long %>%
  filter(!is.na(source_used)) %>%
  summarise(
    manual_review_flags = sum(manual_review_flag, na.rm = TRUE),
    .by = c(model_clean, source_used)
  ) %>%
  arrange(model_clean, desc(manual_review_flags))

active_treatment_combinations <- parsed_long %>%
  semi_join(
    top_unstable_cases,
    by = c("model_clean", "case_id")
  ) %>%
  filter(active_recommendation == 1) %>%
  summarise(
    active_drug_classes = paste(sort(unique(drug_class)), collapse = " + "),
    .by = c(model_clean, case_id, repeat_id, response_id)
  ) %>%
  right_join(
    top_unstable_cases %>%
      select(model_clean, case_id, variation_ratio, n_unique_signatures) %>%
      tidyr::expand_grid(repeat_id = 1:5),
    by = c("model_clean", "case_id", "repeat_id")
  ) %>%
  mutate(
    active_drug_classes = replace_na(active_drug_classes, "No active class parsed")
  ) %>%
  arrange(model_clean, desc(variation_ratio), case_id, repeat_id)

targeted_review_examples <- targeted_review_rows %>%
  mutate(
    example_text_short = str_trunc(example_text, width = 240)
  ) %>%
  select(
    model_clean,
    case_id,
    repeat_id,
    drug_class,
    stop_or_avoid,
    conditional,
    not_preferred,
    unclear,
    example_text_short
  ) %>%
  group_by(model_clean) %>%
  slice_head(n = 12) %>%
  ungroup()

write_csv(
  max_unstable_case_counts,
  file.path(output_dir, "intra_model_max_unstable_case_counts.csv")
)
write_csv(
  manual_review_rate_by_instability,
  file.path(output_dir, "intra_model_manual_review_by_instability.csv")
)
write_csv(
  manual_review_by_model_drug,
  file.path(output_dir, "intra_model_manual_review_by_model_drug.csv")
)
write_csv(
  manual_review_by_model_source,
  file.path(output_dir, "intra_model_manual_review_by_model_source.csv")
)
write_csv(
  active_treatment_combinations,
  file.path(output_dir, "intra_model_top_unstable_active_combinations.csv")
)
write_csv(
  targeted_review_rows,
  file.path(output_dir, "intra_model_targeted_review_rows.csv")
)
write_csv(
  targeted_review_examples,
  file.path(output_dir, "intra_model_targeted_review_examples.csv")
)

max_unstable_case_counts %>%
  kable(caption = "Cases with variation ratio = 0.8 by model") %>%
  kable_styling(full_width = FALSE)

manual_review_rate_by_instability %>%
  mutate(
    manual_review_response_rate = percent(
      manual_review_response_rate,
      accuracy = 0.1
    )
  ) %>%
  kable(caption = "Manual-review burden by max-instability status") %>%
  kable_styling(full_width = FALSE)

manual_review_by_model_drug %>%
  filter(manual_review_flags > 0) %>%
  kable(caption = "Manual-review flags by model and drug class") %>%
  kable_styling(full_width = FALSE)

manual_review_by_model_source %>%
  filter(manual_review_flags > 0) %>%
  kable(caption = "Manual-review flags by model and source") %>%
  kable_styling(full_width = FALSE)

active_treatment_combinations %>%
  select(
    model_clean,
    case_id,
    repeat_id,
    variation_ratio,
    n_unique_signatures,
    active_drug_classes
  ) %>%
  head(45) %>%
  kable(caption = "Active-treatment combinations in top unstable cases") %>%
  kable_styling(full_width = FALSE)

targeted_review_examples %>%
  kable(caption = "Representative parser-flagged rationale examples") %>%
  kable_styling(full_width = FALSE)
```

The targeted review confirms that the medication tables are not the only source
of strict-signature instability. Many parser flags come from rationale text
where models discuss drugs that were considered, avoided, deferred, or not
selected. These rationale-level mentions are clinically informative, but they
can lower full-signature alpha even when the active medication plan is stable.

# Visualizations

```{r plots}
theme_report <- theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.minor = element_blank(),
    legend.position = "bottom"
  )

plot_variation_distribution <- variation_ratio_distribution %>%
  mutate(
    variation_ratio_label = factor(
      sprintf("%.2f", variation_ratio),
      levels = sprintf("%.2f", seq(0, 0.8, by = 0.2))
    )
  ) %>%
  ggplot(aes(x = variation_ratio_label, y = n_cases, fill = model_clean)) +
  geom_col(position = "dodge", width = 0.75) +
  geom_text(
    aes(label = n_cases),
    position = position_dodge(width = 0.75),
    vjust = -0.25,
    size = 3.2
  ) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Variation Ratio Distribution by Model",
    x = "Case-level variation ratio",
    y = "Number of cases",
    fill = "Model"
  ) +
  theme_report

plot_model_summary <- model_summary %>%
  ggplot(
    aes(
      x = reorder(model_clean, mean_variation_ratio),
      y = mean_variation_ratio,
      fill = model_clean
    )
  ) +
  geom_col(width = 0.65, show.legend = FALSE) +
  geom_text(
    aes(label = sprintf("%.2f", mean_variation_ratio)),
    vjust = -0.25,
    size = 3.5
  ) +
  scale_y_continuous(labels = number_format(accuracy = 0.01)) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Mean Intra-Model Variation Ratio",
    x = NULL,
    y = "Mean variation ratio"
  ) +
  theme_report

plot_exact_match <- model_summary %>%
  ggplot(
    aes(
      x = reorder(model_clean, exact_match_rate),
      y = exact_match_rate,
      fill = model_clean
    )
  ) +
  geom_col(width = 0.65, show.legend = FALSE) +
  geom_text(
    aes(label = percent(exact_match_rate, accuracy = 1)),
    vjust = -0.25,
    size = 3.5
  ) +
  scale_y_continuous(labels = percent_format()) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Exact-Match Rate by Model",
    x = NULL,
    y = "Cases with all 5 repeated plans matching"
  ) +
  theme_report

plot_unstable_cases <- top_unstable_cases %>%
  mutate(case_id = fct_reorder(case_id, variation_ratio)) %>%
  ggplot(aes(x = variation_ratio, y = case_id, fill = model_clean)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ model_clean, scales = "free_y") +
  scale_x_continuous(labels = number_format(accuracy = 0.1)) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Top Unstable Cases by Model",
    x = "Variation ratio",
    y = "Case ID"
  ) +
  theme_report

plot_med_agreement <- medication_class_agreement %>%
  ggplot(aes(x = drug_class, y = model_clean, fill = active_agreement_rate)) +
  geom_tile(color = "white") +
  geom_text(aes(label = percent(active_agreement_rate, accuracy = 1)), size = 3) +
  scale_fill_gradient(low = "#f1a340", high = "#1b7837", labels = percent) +
  labs(
    title = "Active-Recommendation Agreement by Medication Class",
    x = "Medication class",
    y = "Model",
    fill = "Agreement"
  ) +
  theme_report +
  theme(axis.text.x = element_text(angle = 35, hjust = 1))

plot_action_agreement <- action_level_agreement %>%
  ggplot(
    aes(
      x = action_indicator,
      y = agreement_rate,
      fill = model_clean
    )
  ) +
  geom_col(position = "dodge", width = 0.75) +
  scale_y_continuous(labels = percent_format()) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Action-Level Agreement by Model",
    x = "Parsed action indicator",
    y = "Agreement rate",
    fill = "Model"
  ) +
  theme_report +
  theme(axis.text.x = element_text(angle = 35, hjust = 1))

plot_alpha_sensitivity <- alpha_sensitivity_long %>%
  mutate(
    metric_definition = factor(
      metric_definition,
      levels = names(alpha_sensitivity_specs)
    ),
    model_label = case_when(
      model_clean == "Claude Opus 4.6" ~ "Claude Opus",
      model_clean == "Claude Sonnet 4.6" ~ "Claude Sonnet",
      model_clean == "GPT-5.4" ~ "GPT-5.4",
      TRUE ~ model_clean
    )
  ) %>%
  ggplot(
    aes(
      x = metric_definition,
      y = alpha,
      fill = model_label
    )
  ) +
  geom_col(position = "dodge", width = 0.75) +
  geom_text(
    aes(label = sprintf("%.3f", alpha)),
    position = position_dodge(width = 0.75),
    vjust = -0.25,
    size = 3.2
  ) +
  scale_y_continuous(
    limits = c(0, 1.05),
    labels = number_format(accuracy = 0.1)
  ) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Krippendorff's Alpha Sensitivity Analysis",
    x = "Metric definition",
    y = "Krippendorff's alpha",
    fill = "Model"
  ) +
  theme_report +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))

plot_medication_alpha <- medication_alpha_long %>%
  mutate(
    metric_definition = factor(
      metric_definition,
      levels = medication_alpha_specs
    ),
    model_label = case_when(
      model_clean == "Claude Opus 4.6" ~ "Claude Opus",
      model_clean == "Claude Sonnet 4.6" ~ "Claude Sonnet",
      model_clean == "GPT-5.4" ~ "GPT-5.4",
      TRUE ~ model_clean
    )
  ) %>%
  ggplot(
    aes(
      x = metric_definition,
      y = alpha,
      fill = model_label
    )
  ) +
  geom_col(position = "dodge", width = 0.75) +
  geom_text(
    aes(label = sprintf("%.3f", alpha)),
    position = position_dodge(width = 0.75),
    vjust = -0.25,
    size = 3.2
  ) +
  scale_y_continuous(
    limits = c(0, 1.05),
    labels = number_format(accuracy = 0.1)
  ) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Medication-Column Dose-Level Alpha",
    x = "Metric definition",
    y = "Krippendorff's alpha",
    fill = "Model"
  ) +
  theme_report +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))

ggsave(
  file.path(export_dir, "variation_ratio_distribution.png"),
  plot_variation_distribution,
  width = 10,
  height = 6,
  dpi = 300
)
ggsave(
  file.path(export_dir, "mean_variation_ratio_by_model.png"),
  plot_model_summary,
  width = 8,
  height = 5,
  dpi = 300
)
ggsave(
  file.path(export_dir, "exact_match_rate_by_model.png"),
  plot_exact_match,
  width = 8,
  height = 5,
  dpi = 300
)
ggsave(
  file.path(export_dir, "top_unstable_cases_by_model.png"),
  plot_unstable_cases,
  width = 10,
  height = 7,
  dpi = 300
)
ggsave(
  file.path(export_dir, "medication_class_agreement_heatmap.png"),
  plot_med_agreement,
  width = 10,
  height = 5,
  dpi = 300
)
ggsave(
  file.path(export_dir, "action_level_agreement_by_model.png"),
  plot_action_agreement,
  width = 11,
  height = 6,
  dpi = 300
)
ggsave(
  file.path(export_dir, "alpha_sensitivity_by_model.png"),
  plot_alpha_sensitivity,
  width = 10,
  height = 6,
  dpi = 300
)
ggsave(
  file.path(export_dir, "medication_column_alpha_by_model.png"),
  plot_medication_alpha,
  width = 10,
  height = 6,
  dpi = 300
)

plot_variation_distribution
plot_model_summary
plot_exact_match
plot_unstable_cases
plot_med_agreement
plot_action_agreement
plot_alpha_sensitivity
plot_medication_alpha
```

# Manual Review Sample

```{r manual-review-sample}
manual_review_sample <- top_unstable_cases %>%
  select(model_clean, case_id, variation_ratio, n_unique_signatures) %>%
  left_join(
    treatment_signatures %>%
      select(
        model_clean,
        case_id,
        repeat_id,
        response_id,
        treatment_signature,
        n_active_classes,
        n_stop_avoid_classes,
        n_manual_review_classes
      ),
    by = c("model_clean", "case_id")
  ) %>%
  left_join(
    standardized %>%
      select(response_id, treatment_plan),
    by = "response_id"
  ) %>%
  arrange(model_clean, desc(variation_ratio), case_id, repeat_id)

write_csv(
  manual_review_sample,
  file.path(output_dir, "intra_model_manual_review_sample.csv")
)

manual_review_sample %>%
  select(
    model_clean,
    case_id,
    repeat_id,
    variation_ratio,
    n_active_classes,
    n_stop_avoid_classes,
    n_manual_review_classes
  ) %>%
  head(30) %>%
  kable(caption = "Manual review sample from unstable cases") %>%
  kable_styling(full_width = FALSE)
```

# Manuscript Exports

```{r manuscript-exports}
make_export_table <- function(data) {
  flextable(data) %>%
    theme_booktabs() %>%
    autofit()
}

add_table_to_doc <- function(doc, title, data) {
  doc %>%
    body_add_par(title, style = "heading 2") %>%
    body_add_flextable(make_export_table(data)) %>%
    body_add_par("", style = "Normal")
}

model_summary_export <- model_summary %>%
  mutate(
    mean_variation_ratio = round(mean_variation_ratio, 3),
    median_variation_ratio = round(median_variation_ratio, 3),
    exact_match_rate = percent(exact_match_rate, accuracy = 0.1),
    krippendorff_alpha = round(krippendorff_alpha, 3)
  )

variation_ratio_distribution_export <- variation_ratio_distribution %>%
  mutate(percent_cases = percent(percent_cases, accuracy = 0.1))

alpha_by_model_export <- alpha_by_model %>%
  mutate(krippendorff_alpha = round(krippendorff_alpha, 3))

medication_class_agreement_export <- medication_class_agreement %>%
  mutate(
    mentioned_agreement_rate = percent(mentioned_agreement_rate, accuracy = 0.1),
    active_agreement_rate = percent(active_agreement_rate, accuracy = 0.1),
    stop_avoid_agreement_rate = percent(stop_avoid_agreement_rate, accuracy = 0.1)
  )

action_level_agreement_export <- action_level_agreement %>%
  mutate(agreement_rate = percent(agreement_rate, accuracy = 0.1))

manual_review_burden_export <- manual_review_burden %>%
  mutate(
    response_manual_review_rate = percent(
      response_manual_review_rate,
      accuracy = 0.1
    )
  )

manual_review_rate_by_instability_export <-
  manual_review_rate_by_instability %>%
  mutate(
    manual_review_response_rate = percent(
      manual_review_response_rate,
      accuracy = 0.1
    )
  )

medication_alpha_wide_export <- medication_alpha_wide

top_medication_dose_disagreement_export <-
  top_medication_dose_disagreement_cases %>%
  mutate(
    medication_dose_variation_ratio = round(
      medication_dose_variation_ratio,
      3
    )
  )

top_active_combinations_export <- active_treatment_combinations %>%
  select(
    model_clean,
    case_id,
    repeat_id,
    variation_ratio,
    n_unique_signatures,
    active_drug_classes
  ) %>%
  head(60)

targeted_review_examples_export <- targeted_review_examples

word_output_path <- file.path(
  export_dir,
  "intra_model_validation_tables.docx"
)

doc <- read_docx() %>%
  body_add_par("Intra-Model Validation Tables", style = "heading 1")

doc <- add_table_to_doc(doc, "Dataset overview", overview)
doc <- add_table_to_doc(doc, "Repeat-count validation", repeat_check)
doc <- add_table_to_doc(
  doc,
  "Model-level stability summary",
  model_summary_export
)
doc <- add_table_to_doc(
  doc,
  "Variation ratio distribution",
  variation_ratio_distribution_export
)
doc <- add_table_to_doc(
  doc,
  "Krippendorff's alpha",
  alpha_by_model_export
)
doc <- add_table_to_doc(
  doc,
  "Alpha sensitivity analysis",
  alpha_sensitivity_wide
)
doc <- add_table_to_doc(
  doc,
  "Medication-column alpha",
  medication_alpha_wide_export
)
doc <- add_table_to_doc(
  doc,
  "Top medication action-dose disagreement cases",
  top_medication_dose_disagreement_export
)
doc <- add_table_to_doc(
  doc,
  "Medication-class agreement",
  medication_class_agreement_export
)
doc <- add_table_to_doc(
  doc,
  "Action-level agreement",
  action_level_agreement_export
)
doc <- add_table_to_doc(
  doc,
  "Manual-review burden",
  manual_review_burden_export
)
doc <- add_table_to_doc(
  doc,
  "Max-unstable case counts",
  max_unstable_case_counts
)
doc <- add_table_to_doc(
  doc,
  "Manual-review burden by instability status",
  manual_review_rate_by_instability_export
)
doc <- add_table_to_doc(
  doc,
  "Manual-review flags by model and drug class",
  manual_review_by_model_drug %>% filter(manual_review_flags > 0)
)
doc <- add_table_to_doc(
  doc,
  "Manual-review flags by model and source",
  manual_review_by_model_source %>% filter(manual_review_flags > 0)
)
doc <- add_table_to_doc(
  doc,
  "Top unstable active-treatment combinations",
  top_active_combinations_export
)
doc <- add_table_to_doc(
  doc,
  "Representative parser-flagged rationale examples",
  targeted_review_examples_export
)

print(doc, target = word_output_path)

tibble(
  export_file = list.files(export_dir, full.names = FALSE)
) %>%
  arrange(export_file) %>%
  kable(caption = "Manuscript export files") %>%
  kable_styling(full_width = FALSE)
```

# Output Files

```{r output-files}
output_files <- tibble(
  file = list.files(output_dir, full.names = FALSE)
) %>%
  arrange(file)

output_files %>%
  kable(caption = "Files written by this workflow") %>%
  kable_styling(full_width = FALSE)
```
