---
title: "Follow-Up Analyses"
output: html_document
---

<style>
h2 { margin-top: 60px; }
</style>

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(dplyr)
library(readr)
library(knitr)
library(kableExtra)
library(broom)
library(tidyr)
library(stringr)

ind_path <- "C:/Users/rosie/Documents/capstone_project/data/processed/changchangpan/capstone_llm_data_explore_r_outputs/processed_llm_treatment_indicators.csv"
dis_path <- "C:/Users/rosie/Documents/capstone_project/data/processed/changchangpan/capstone_llm_data_explore_r_outputs/high_disagreement_cases.csv"
raw_path <- "C:/Users/rosie/Documents/capstone_project/data/raw/final_results_gpt_claude_models.csv"

ind <- read_csv(ind_path, show_col_types = FALSE)

# Patient-level frame: one row per case for patient covariates
case <- ind %>% distinct(case_id, .keep_all = TRUE)
```



## Disagreement vs. Complexity
Hypothesis that model disagreement rises with clinical
complexity; defined as number of comorbidities and lower eGFR.

```{r disagreement-corr}
dis <- read_csv(dis_path, show_col_types = FALSE)

comorbid_cols <- c(
  "heart_failure_present", "chronic_kidney_disease", "mi_history",
  "stroke_tia_history", "coronary_artery_disease", "atrial_fibrillation",
  "peripheral_neuropathy", "diabetic_retinopathy", "hypertension",
  "hyperlipidemia", "hypothyroidism", "depression", "copd", "asthma",
  "anemia", "osteoporosis", "rheumatoid_arthritis", "cancer_breast",
  "cancer_colon", "cancer_lung", "cancer_prostate", "cancer_endometrial"
)
comorbid_cols <- intersect(comorbid_cols, names(case))

case <- case %>%
  mutate(n_comorbid = rowSums(across(all_of(comorbid_cols), ~ . == "Yes"),
                              na.rm = TRUE)) %>%
  left_join(dis %>% select(case_id, disagreement_score), by = "case_id") %>%
  mutate(disagreement_score = replace_na(disagreement_score, 0))

c1 <- cor.test(case$n_comorbid, case$disagreement_score, method = "spearman")
c2 <- cor.test(case$egfr_ml_min_1_73m2, case$disagreement_score,
               method = "spearman")

tibble(
  Comparison = c("Disagreement vs. # Comorbidities",
                 "Disagreement vs. eGFR"),
  `Spearman rho` = round(c(c1$estimate, c2$estimate), 3),
  `p-value` = round(c(c1$p.value, c2$p.value), 4)
) %>%
  kable(caption = paste0("Spearman Correlations (n = ", nrow(case), " Cases)")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```


```{r disagreement-by-egfr}
case %>%
  mutate(egfr_band = cut(egfr_ml_min_1_73m2, c(0, 30, 45, 60, Inf),
                         labels = c("<30", "30-44", "45-59", ">=60"),
                         right = FALSE)) %>%
  group_by(`eGFR band` = egfr_band) %>%
  summarise(`Mean disagreement` = round(mean(disagreement_score), 2),
            N = n(), .groups = "drop") %>%
  kable(caption = "Mean Disagreement Score by eGFR Band") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

**Comments:** Both correlations run in the way expected, but neither is
significant at this sample size. Disagreement vs. comorbidity count is weak
(rho = 0.13, p = 0.20). Disagreement vs. eGFR is stronger and near-significant
(rho = -0.19, p = 0.07). Mean disagreement is highest in the CKD groups (2.08 at
eGFR <30, 1.90 at 30-44) and lower above 45 (1.32 at 45-59, 1.57 at >=60). This
is consistent with the idea that guidelines are clear for simple patients and
ambiguous for complex ones. Supports expanding the sample if we want to
confirm it statistically.



## Baseline Diabetes Medications
```{r baseline-dm-drugs}
drug_pat <- c(
  "Metformin" = "metformin",
  "SGLT2 inhibitor" = "gliflozin|empagliflozin|dapagliflozin|canagliflozin",
  "GLP-1 RA" = "semaglutide|dulaglutide|liraglutide|tirzepatide|exenatide|lixisenatide",
  "DPP-4 inhibitor" = "gliptin|sitagliptin|linagliptin|saxagliptin",
  "Sulfonylurea" = "glipizide|glyburide|glimepiride|gliclazide",
  "Basal insulin" = "glargine|detemir|degludec|NPH|isophane",
  "Prandial insulin" = "aspart|lispro|glulisine"
)

drugs_lc <- tolower(coalesce(case$current_dm_drugs, ""))
none_n <- sum(drugs_lc == "")

tibble(
  `Drug class` = names(drug_pat),
  `n cases` = sapply(drug_pat, function(p) sum(str_detect(drugs_lc, p)))
) %>%
  bind_rows(tibble(`Drug class` = "No diabetes medication (newly dx)",
                   `n cases` = none_n)) %>%
  mutate(`%` = round(100 * `n cases` / nrow(case), 1)) %>%
  arrange(desc(`n cases`)) %>%
  kable(caption = paste0("Baseline Diabetes Medications (n = ", nrow(case),
        " Cases; classes overlap, patients may be on >1)")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

**Comments:** Roughly half the population is on no diabetes medication at
baseline (newly diagnosed). Among those on therapy, GLP-1 RAs, metformin, and
SGLT2 inhibitors are the most common classes. Classes overlap because patients
may be on combination regimens, so percentages do not sum to 100.



## Race / Ethnicity

```{r race-dist}
case %>%
  mutate(race_ethnicity = coalesce(race_ethnicity, "Unknown")) %>%
  count(`Race / Ethnicity` = race_ethnicity) %>%
  mutate(`%` = round(100 * n / sum(n), 1)) %>%
  arrange(desc(n)) %>%
  kable(caption = paste0("Race / Ethnicity Distribution (n = ", nrow(case),
                         " Cases)")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```



## GPT-5.4 Metformin Review (eGFR < 30)
Pulling the GPT-5.4 plans for low-eGFR cases to see whether the apparent metformin
gap is real.

```{r metformin-review}
raw <- read_csv(raw_path, show_col_types = FALSE)

raw %>%
  filter(model == "azure/gpt-5.4", egfr_ml_min_1_73m2 < 30) %>%
  mutate(on_metformin = str_detect(tolower(coalesce(current_dm_drugs, "")),
                                   "metformin")) %>%
  arrange(case_id) %>%
  select(`Case` = case_id, `eGFR` = egfr_ml_min_1_73m2,
         `On metformin at baseline` = on_metformin,
         `Baseline DM drugs` = current_dm_drugs) %>%
  kable(caption = "GPT-5.4 Cases with eGFR < 30") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

**Comments:** The 15% figure is misleading. It comes from the
metformin_stop_or_caution flag, which only triggers when the text mentions
stopping or cautioning metformin, divided by all 13 eGFR <30 cases. Only 2 of the
13 eGFR <30 cases (DM072, DM077) were on metformin at baseline; these are the only
real metformin decisions, and GPT-5.4 discontinued the drug and cited the eGFR in
both. The remaining 11 patients were never on metformin, so there was nothing to
stop, and no model recommends starting metformin below eGFR 30. The 2/13 here is
the 15% from the original flag, so this is the same data, just counted correctly.
So GPT handled metformin correctly here. The low number is a counting issue, not a
safety problem. The full treatment_plan text is below.


```{r metformin-text, results='asis'}
txt <- raw %>%
  filter(model == "azure/gpt-5.4", egfr_ml_min_1_73m2 < 30) %>%
  arrange(case_id)

for (i in seq_len(nrow(txt))) {
  cat("\n\n**", txt$case_id[i], "** (eGFR ", txt$egfr_ml_min_1_73m2[i], ")\n\n",
      sep = "")
  cat("> ", gsub("\n", "\n> ", txt$treatment_plan[i]), "\n", sep = "")
}
```



## Univariate Logistic Regression (eGFR >= 60 reference)
Re-running the medication-class associations as univariate models with eGFR as
the sole predictor and eGFR >= 60 as the reference category, to address the
complete separation seen in the multivariable models.

```{r univariate-logit}
ind <- ind %>%
  mutate(egfr_cat = case_when(
    egfr_ml_min_1_73m2 < 30 ~ "<30",
    egfr_ml_min_1_73m2 < 45 ~ "30-44",
    egfr_ml_min_1_73m2 < 60 ~ "45-59",
    TRUE                    ~ ">=60"),
    egfr_cat = relevel(factor(egfr_cat), ref = ">=60"))

run_one <- function(outcome) {
  f <- as.formula(paste(outcome, "~ egfr_cat"))
  suppressWarnings(glm(f, data = ind, family = binomial)) %>%
    tidy(exponentiate = TRUE, conf.int = TRUE) %>%
    filter(term != "(Intercept)") %>%
    mutate(Outcome = outcome,
           Level = paste0(sub("egfr_cat", "", term), " vs >=60")) %>%
    select(Outcome, Level, OR = estimate, CI_low = conf.low,
           CI_high = conf.high, p = p.value)
}

outcomes <- c("metformin_active", "sglt2i_active",
              "glp1_active", "insulin_any_active")

bind_rows(lapply(outcomes, run_one)) %>%
  mutate(sep = is.infinite(CI_high) | CI_high > 1e6 | CI_low < 1e-6,
         across(c(OR, CI_low, CI_high), ~ round(., 2)),
         p = round(p, 4),
         note = ifelse(sep, "separation - report as proportion", "")) %>%
  mutate(across(c(OR, CI_low, CI_high), ~ ifelse(sep, NA, .))) %>%
  mutate(across(c(OR, CI_low, CI_high), ~ ifelse(is.na(.), "—", as.character(.)))) %>%
  select(-sep) %>%
  kable(caption = "Univariate Logistic Regression: Medication Class by eGFR (ref = >=60)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

```{r metformin-separation-proportion}
ind %>%
  filter(egfr_cat == "<30") %>%
  summarise(N = n(),
            `N metformin active` = sum(metformin_active, na.rm = TRUE),
            `%` = round(100 * mean(metformin_active, na.rm = TRUE), 1)) %>%
  kable(caption = "Metformin in eGFR < 30 (report as proportion, not OR)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

**Comments:** The univariate models with eGFR >= 60 as reference have sensible
CIs, unlike the multivariable models. Metformin is less likely as eGFR falls (OR =
0.24 at 45-59, 0.04 at 30-44). SGLT2 inhibitors peak in the moderate-CKD bands and
drop below 30. Insulin rises at low eGFR. The metformin eGFR <30 cell
remains a complete-separation case (0 of 39 responses recommend metformin) and is
reported as a simple proportion rather than an odds ratio.



## Guideline Flag (All Stop/Caution Flags)
Checking whether the denominator problem found in the metformin flag also affects
the other stop/caution flags. A flag should only count as a "miss" when the
patient was actually on the drug.

```{r flag-audit}
drug_kw <- c(
  metformin_stop_or_caution = "metformin",
  sglt2i_stop_or_caution = "gliflozin",
  glp1_stop_or_caution =
    "semaglutide|dulaglutide|liraglutide|tirzepatide|exenatide|lixisenatide",
  insulin_stop_or_caution =
    "glargine|detemir|degludec|aspart|lispro|glulisine|insulin|NPH"
  )

drugs_lc <- tolower(coalesce(ind$current_dm_drugs, ""))

audit <- lapply(names(drug_kw), function(flag) {
  on_drug <- str_detect(drugs_lc, drug_kw[[flag]])
  flagged <- ind[[flag]] == 1
  tibble(
    Flag = flag,
    `Rows flagged` = sum(flagged, na.rm = TRUE),
    `Flagged & on drug` = sum(flagged & on_drug, na.rm = TRUE),
    `Flagged but NOT on drug` = sum(flagged & !on_drug, na.rm = TRUE)
  )
}) %>% bind_rows()

audit %>%
  kable(caption = "Stop/Caution Flag Audit: How Many Flagged Rows are Patients not on the Drug") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

**Comments:** The denominator problem is not specific to metformin. For all four
stop/caution flags, the majority of flagged rows are patients who were not on the
drug at baseline, so the flag is counting "did not mention stopping a drug the
patient was not taking" as a failure. Any guideline-adherence rate built on these
flags is inflated until it is restricted to patients actually on the drug.



## Heart Failure as a Complexity 
Heart failure as a source of competing indications.
Looking at disagreement by HF status, and SGLT2i recommendation rates by HF
status (HF is a guideline push toward SGLT2i).

```{r hf-disagreement}
case %>%
  group_by(`Heart failure` = heart_failure_present) %>%
  summarise(`Mean disagreement` = round(mean(disagreement_score), 2),
            N = n(), .groups = "drop") %>%
  kable(caption = "Mean Disagreement by Heart Failure Status") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

# Mann-Whitney
hf_test <- suppressWarnings(
  wilcox.test(disagreement_score ~ heart_failure_present, data = case)
  )
hf_p <- round(hf_test$p.value, 2)
# hf_p
```


```{r hf-sglt2i}
ind %>%
  group_by(Model = model_short, `Heart failure` = heart_failure_present) %>%
  summarise(`SGLT2i recommended (%)` = round(100 * mean(sglt2i_active, na.rm = TRUE), 0),
            N = n(), .groups = "drop") %>%
  kable(caption = "SGLT2i Recommendation Rate by Model and Heart Failure Status") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

**Comments:** HF cases do trend higher on disagreement (2.0 vs 1.6), but there
are only 12 of them, so it is not significant (Mann-Whitney p = 0.57). All three
models push SGLT2i much harder in HF cases (GPT and Opus to 100%, Sonnet to 75%),
which is what guidelines call for, and the split between Sonnet and the other two
is part of where the disagreement comes from.
