---
title: "Claude Intra-Model Response Generation"
author: "Group 30"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 2
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)

required_packages <- c(
  "httr2", "jsonlite", "readr", "dplyr", "purrr",
  "tibble", "stringr", "knitr"
)
missing_packages <- required_packages[
  !vapply(required_packages, requireNamespace, logical(1), quietly = TRUE)
]

if (length(missing_packages) > 0) {
  stop("Missing required packages: ", paste(missing_packages, collapse = ", "))
}

library(httr2)
library(jsonlite)
library(readr)
library(dplyr)
library(purrr)
library(tibble)
library(stringr)
library(knitr)

find_project_root <- function(start_dir = getwd()) {
  candidates <- normalizePath(
    file.path(start_dir, c(".", "..", "../..", "../../..")),
    mustWork = FALSE
  )

  matched <- candidates[
    file.exists(file.path(candidates, "data/raw/final_results_gpt_claude_models.csv"))
  ]

  if (length(matched) == 0) {
    stop("Could not find project root containing data/raw/final_results_gpt_claude_models.csv.")
  }

  matched[[1]]
}

project_root <- find_project_root()
knitr::opts_knit$set(root.dir = project_root)
dir.create(file.path(project_root, "data/inner model validation"), showWarnings = FALSE, recursive = TRUE)
```

# Purpose

This workflow is generation-only. It reads the prebuilt prompt request dataset for intra-model validation, sends each `prompt_text` row to Claude through the Anthropic Messages API, and writes one `treatment_plan` response per request row.

It does not rebuild prompts, reshape the raw vignette dataset, or perform scoring.

# Configuration

```{r configuration}
run_generation <- TRUE

prompt_request_input_path <- "data/inner model validation/intra_model_validation_prompt_requests.csv"
response_output_path <- "data/inner model validation/claude_intra_model_responses.csv"

# API keys are read from the local environment only.
# Do not paste secrets into this file.
api_key <- Sys.getenv("ANTHROPIC_API_KEY")
base_url <- "https://api.anthropic.com/v1/messages"
anthropic_version <- "2023-06-01"
model_id <- "claude-sonnet-4-6"

temperature <- 0
max_output_tokens <- 2500
request_pause_seconds <- 1
overwrite_existing <- TRUE
max_requests <- 100
retry_attempts <- 2
```

# Input Preparation

```{r input-preparation}
prompt_requests_raw <- read_csv(prompt_request_input_path, show_col_types = FALSE)

blank_names <- names(prompt_requests_raw) == ""
if (any(blank_names)) {
  names(prompt_requests_raw)[blank_names] <- paste0("blank_column_", seq_len(sum(blank_names)))
}

required_input_columns <- c(
  "case_id",
  "original_case_id",
  "case_variant_id",
  "vignette_text",
  "prompt_text",
  "model_type",
  "prompt_type",
  "repeat_id"
)

missing_input_columns <- setdiff(required_input_columns, names(prompt_requests_raw))
if (length(missing_input_columns) > 0) {
  stop("Missing required input columns: ", paste(missing_input_columns, collapse = ", "))
}

prompt_requests <- prompt_requests_raw %>%
  mutate(
    case_id = as.character(case_id),
    original_case_id = as.character(original_case_id),
    case_variant_id = as.integer(case_variant_id),
    vignette_text = as.character(vignette_text),
    prompt_text = as.character(prompt_text),
    model_type = as.character(model_type),
    prompt_type = as.character(prompt_type),
    repeat_id = suppressWarnings(as.integer(repeat_id))
  )

if (any(is.na(prompt_requests$repeat_id))) {
  stop("repeat_id contains missing or non-integer values.")
}

if (any(!nzchar(str_squish(prompt_requests$prompt_text)))) {
  stop("prompt_text contains blank rows. Clean the prompt request dataset before generation.")
}

duplicate_request_keys <- prompt_requests %>%
  count(case_id, model_type, repeat_id, name = "n") %>%
  filter(n > 1)

if (nrow(duplicate_request_keys) > 0) {
  stop("Duplicate case_id + model_type + repeat_id rows found in the prompt request dataset.")
}

prompt_requests <- prompt_requests %>%
  mutate(model_type = model_id) %>%
  arrange(case_id, repeat_id)

if (!is.null(max_requests)) {
  prompt_requests <- prompt_requests %>%
    slice_head(n = max_requests)
}

input_validation <- tibble(
  metric = c(
    "Input rows",
    "Distinct case IDs",
    "Distinct original case IDs",
    "Duplicate request keys",
    "Configured model ID",
    "Configured max_requests"
  ),
  value = c(
    nrow(prompt_requests),
    n_distinct(prompt_requests$case_id),
    n_distinct(prompt_requests$original_case_id),
    nrow(duplicate_request_keys),
    model_id,
    ifelse(is.null(max_requests), "NULL", as.character(max_requests))
  )
)

input_validation %>% kable()
```

```{r input-preview}
prompt_requests %>%
  mutate(prompt_text = str_trunc(prompt_text, 200)) %>%
  select(case_id, original_case_id, case_variant_id, model_type, prompt_type, repeat_id, prompt_text) %>%
  slice_head(n = min(10, nrow(prompt_requests))) %>%
  kable(caption = "Prompt request input preview")
```

# Anthropic API Helpers

```{r api-helpers}
validate_generation_configuration <- function(api_key, base_url, model_id, retry_attempts) {
  if (!is.character(api_key) || !nzchar(api_key)) {
    stop("ANTHROPIC_API_KEY is not set. Add it to your local environment before running generation.")
  }
  if (!is.character(base_url) || !nzchar(base_url)) {
    stop("base_url is blank.")
  }
  if (!str_detect(str_to_lower(base_url), "anthropic\\.com")) {
    stop("base_url must be an Anthropic endpoint for this workflow.")
  }
  if (!is.character(model_id) || !nzchar(model_id)) {
    stop("model_id is blank.")
  }
  if (!str_detect(model_id, "^claude-")) {
    stop("This workflow is restricted to Claude model IDs (must start with 'claude-').")
  }
  if (!is.numeric(retry_attempts) || length(retry_attempts) != 1 || retry_attempts < 0) {
    stop("retry_attempts must be a non-negative number.")
  }
}

extract_anthropic_response_text <- function(body) {
  if (!is.null(body$content) && length(body$content) > 0) {
    text_parts <- unlist(
      map(body$content, function(item) {
        if (!is.null(item$text)) item$text else ""
      })
    )

    text_parts <- text_parts[nzchar(text_parts)]
    if (length(text_parts) > 0) {
      return(paste(text_parts, collapse = "\n"))
    }
  }

  ""
}

call_anthropic_response <- function(prompt_text, model_id, api_key, base_url, anthropic_version, temperature, max_output_tokens) {
  body <- list(
    model = model_id,
    max_tokens = max_output_tokens,
    temperature = temperature,
    messages = list(
      list(role = "user", content = prompt_text)
    )
  )

  response <- request(base_url) %>%
    req_headers(
      `x-api-key` = api_key,
      `anthropic-version` = anthropic_version,
      `Content-Type` = "application/json"
    ) %>%
    req_body_json(body, auto_unbox = TRUE) %>%
    req_perform()

  parsed <- resp_body_json(response, simplifyVector = FALSE)
  extract_anthropic_response_text(parsed)
}

call_anthropic_response_with_retry <- function(prompt_text, model_id, api_key, base_url, anthropic_version, temperature, max_output_tokens, retry_attempts) {
  max_attempts <- retry_attempts + 1L
  last_error <- NULL

  for (attempt in seq_len(max_attempts)) {
    result <- tryCatch(
      call_anthropic_response(
        prompt_text = prompt_text,
        model_id = model_id,
        api_key = api_key,
        base_url = base_url,
        anthropic_version = anthropic_version,
        temperature = temperature,
        max_output_tokens = max_output_tokens
      ),
      error = function(e) {
        last_error <<- e
        NULL
      }
    )

    if (!is.null(result)) {
      return(result)
    }

    if (attempt < max_attempts) {
      Sys.sleep(2 ^ (attempt - 1))
    }
  }

  stop(conditionMessage(last_error))
}
```

# Claude Response Generation

```{r generation}
claude_responses <- tibble()

if (run_generation) {
  validate_generation_configuration(
    api_key = api_key,
    base_url = base_url,
    model_id = model_id,
    retry_attempts = retry_attempts
  )

  if (file.exists(response_output_path) && !overwrite_existing) {
    stop("Response output file already exists and overwrite_existing is FALSE: ", response_output_path)
  }

  claude_responses <- pmap_dfr(
    prompt_requests,
    function(case_id, original_case_id, case_variant_id, vignette_text, prompt_text, model_type, prompt_type, repeat_id) {
      result <- tryCatch(
        {
          response_text <- call_anthropic_response_with_retry(
            prompt_text = prompt_text,
            model_id = model_id,
            api_key = api_key,
            base_url = base_url,
            anthropic_version = anthropic_version,
            temperature = temperature,
            max_output_tokens = max_output_tokens,
            retry_attempts = retry_attempts
          )

          tibble(
            case_id = case_id,
            original_case_id = original_case_id,
            case_variant_id = case_variant_id,
            model_type = model_id,
            prompt_type = prompt_type,
            repeat_id = repeat_id,
            temperature = temperature,
            prompt_text = prompt_text,
            treatment_plan = response_text,
            error_message = ""
          )
        },
        error = function(e) {
          tibble(
            case_id = case_id,
            original_case_id = original_case_id,
            case_variant_id = case_variant_id,
            model_type = model_id,
            prompt_type = prompt_type,
            repeat_id = repeat_id,
            temperature = temperature,
            prompt_text = prompt_text,
            treatment_plan = "",
            error_message = conditionMessage(e)
          )
        }
      )

      Sys.sleep(request_pause_seconds)
      result
    }
  )
}
```

# Validation and Export

```{r validation-export}
processed_rows <- if (run_generation) nrow(claude_responses) else 0L
non_empty_responses <- if (run_generation) sum(nzchar(claude_responses$treatment_plan)) else 0L
failed_rows <- if (run_generation) sum(nzchar(claude_responses$error_message)) else 0L

duplicate_response_keys <- if (run_generation) {
  claude_responses %>%
    count(case_id, model_type, repeat_id, name = "n") %>%
    filter(n > 1)
} else {
  tibble(case_id = character(), model_type = character(), repeat_id = integer(), n = integer())
}

validation_summary <- tibble(
  metric = c(
    "Input rows",
    "Processed rows",
    "Non-empty responses",
    "Failed rows",
    "Duplicate response keys"
  ),
  value = c(
    nrow(prompt_requests),
    processed_rows,
    non_empty_responses,
    failed_rows,
    nrow(duplicate_response_keys)
  )
)

validation_summary %>% kable()

if (run_generation) {
  if (processed_rows != nrow(prompt_requests)) {
    stop("Processed row count does not match input request row count.")
  }
  if (nrow(duplicate_response_keys) > 0) {
    stop("Duplicate case_id + model_type + repeat_id rows detected in generated responses.")
  }
  if (non_empty_responses == 0) {
    stop("All generated treatment_plan values are blank. Check the API key, endpoint, and model access.")
  }

  write_csv(claude_responses, response_output_path)
}
```

```{r response-preview}
if (run_generation) {
  claude_responses %>%
    mutate(
      prompt_text = str_trunc(prompt_text, 120),
      treatment_plan = str_trunc(treatment_plan, 200)
    ) %>%
    select(case_id, repeat_id, model_type, prompt_type, treatment_plan, error_message) %>%
    slice_head(n = min(10, nrow(claude_responses))) %>%
    kable(caption = "Generated response preview")
}
```

# Run Notes

Render once with `run_generation <- FALSE` to verify the input file and validation tables. For a first live test, set `run_generation <- TRUE`, make sure `ANTHROPIC_API_KEY` is available in your local environment, and set `max_requests <- 10` before running all 100 prompt requests.
