---
title: "Intra-Model Consistency API Generation"
author: "Group 30"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 2
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)

required_packages <- c(
  "httr2", "jsonlite", "readr", "dplyr", "purrr",
  "tibble", "stringr", "knitr", "tidyr"
)
missing_packages <- required_packages[
  !vapply(required_packages, requireNamespace, logical(1), quietly = TRUE)
]

if (length(missing_packages) > 0) {
  stop("Missing required packages: ", paste(missing_packages, collapse = ", "))
}

library(httr2)
library(jsonlite)
library(readr)
library(dplyr)
library(purrr)
library(tibble)
library(stringr)
library(knitr)
library(tidyr)

find_project_root <- function(start_dir = getwd()) {
  candidates <- normalizePath(
    file.path(start_dir, c(".", "..", "../..", "../../..")),
    mustWork = FALSE
  )

  matched <- candidates[
    file.exists(file.path(candidates, "data/raw/final_results_gpt_claude_models.csv"))
  ]

  if (length(matched) == 0) {
    stop("Could not find project root containing data/raw/final_results_gpt_claude_models.csv.")
  }

  matched[[1]]
}

project_root <- find_project_root()
knitr::opts_knit$set(root.dir = project_root)
dir.create(file.path(project_root, "data/inner model validation"), showWarnings = FALSE, recursive = TRUE)
dir.create(file.path(project_root, "data/processed"), showWarnings = FALSE, recursive = TRUE)
```

# Purpose

This workflow creates a new dataset for intra-model validation. It reads the original model-output dataset, extracts one row per clinical vignette, combines each vignette with the original diabetes medication prompt, expands each case to repeated runs, and optionally calls a local or remote model endpoint to generate new `treatment_plan` responses.

The main output needed before any model call is the prompt request dataset:

`data/inner model validation/intra_model_validation_prompt_requests.csv`

# Configuration

```{r configuration}
# Keep FALSE to only create the validation prompt dataset.
# Set TRUE only when model endpoint settings below are ready.
run_generation <- FALSE

source_data_path <- "data/raw/final_results_gpt_claude_models.csv"
prompt_request_output_path <- "data/inner model validation/intra_model_validation_prompt_requests.csv"
response_output_path <- NULL

# Use one model per run. This value is written into the prompt request dataset.
model_id <- "local-model"

n_repeats <- 5
prompt_type <- "baseline"
temperature <- 0
max_output_tokens <- 2500
request_pause_seconds <- 1
overwrite_existing <- TRUE

# Optional: limit cases for a small test dataset. Use NULL for all cases.
max_cases <- NULL

# Local model endpoints usually expose an OpenAI-compatible API, for example:
# base_url <- "http://localhost:11434/v1/chat/completions"
# base_url <- "http://localhost:1234/v1/chat/completions"
# api_key can often be blank for local models.
base_url <- ""
api_key <- ""
```

# Original Prompt

```{r original-prompt}
original_prompt <- paste(
  "You are a primary care provider helping this patient manage their diabetes.",
  "They present for a visit in your office with the following history.",
  "Please recommend changes or additions to their diabetes medications.",
  "Please output one row per medication.",
  "If the medication dose or frequency is being changed, describe the change and the final dose and frequency.",
  "If the medication is being continued without change, say so and give the continued dose and frequency.",
  "If the medication is being stopped note this.",
  "If a new medication is being added, give the name of the medication and the dose and frequency.",
  "If titration is planned in the next few months, please simply state that the medication will be titrated up.",
  "For rapid acting insulin regimens, please summarize the dosing schedule concisely but it is fine to take 2-3 lines of text if needed.",
  "Do NOT give contingency plans or choices between several options.",
  "We need the patient to walk out with a single plan to implement.",
  "ONLY provide recommendations for antihyperglycemic drugs - no vitamins, blood pressure medications, lipid lowering medications, for example.",
  "Per medication, this output should all go in a single cell of a table.",
  "The output should include a table with one column and one row per medication that can be easily copied and pasted into a CSV file.",
  sep = " "
)

build_prompt_text <- function(vignette_text) {
  paste(
    original_prompt,
    "",
    "Patient Vignette:",
    vignette_text,
    sep = "\n"
  )
}
```

# Build Validation Prompt Dataset

```{r build-prompt-requests}
raw_data <- read_csv(source_data_path, show_col_types = FALSE)

blank_names <- names(raw_data) == ""
if (any(blank_names)) {
  names(raw_data)[blank_names] <- paste0("blank_column_", seq_len(sum(blank_names)))
}

required_source_columns <- c("case_id", "vignette_text")
missing_source_columns <- setdiff(required_source_columns, names(raw_data))

if (length(missing_source_columns) > 0) {
  stop("Missing required source columns: ", paste(missing_source_columns, collapse = ", "))
}

case_vignettes <- raw_data %>%
  distinct(case_id, vignette_text) %>%
  arrange(case_id, vignette_text) %>%
  group_by(case_id) %>%
  mutate(
    case_variant_id = row_number(),
    n_case_variants = n(),
    validation_case_id = if_else(
      n_case_variants == 1,
      case_id,
      paste0(case_id, "_variant", case_variant_id)
    )
  ) %>%
  ungroup()

if (!is.null(max_cases)) {
  case_vignettes <- case_vignettes %>%
    slice_head(n = max_cases)
}

prompt_requests <- case_vignettes %>%
  transmute(
    original_case_id = case_id,
    case_id = validation_case_id,
    case_variant_id = case_variant_id,
    vignette_text = vignette_text,
    prompt_text = map_chr(vignette_text, build_prompt_text),
    model_type = model_id,
    prompt_type = prompt_type
  ) %>%
  tidyr::uncount(weights = n_repeats, .id = "repeat_id") %>%
  mutate(repeat_id = as.integer(repeat_id)) %>%
  arrange(case_id, model_type, repeat_id)

duplicate_request_keys <- prompt_requests %>%
  count(case_id, model_type, repeat_id, name = "n") %>%
  filter(n > 1)

if (nrow(duplicate_request_keys) > 0) {
  stop("Duplicate case_id + model_type + repeat_id rows were created.")
}

write_csv(prompt_requests, prompt_request_output_path)

validation_overview <- tibble(
  metric = c(
    "Source rows",
    "Distinct original case IDs",
    "Distinct case/vignette pairs",
    "Case IDs with multiple vignette texts",
    "Prompt request rows",
    "Repeats per case/model",
    "Prompt request output"
  ),
  value = c(
    nrow(raw_data),
    n_distinct(raw_data$case_id),
    nrow(case_vignettes),
    sum(case_vignettes$n_case_variants > 1),
    nrow(prompt_requests),
    n_repeats,
    prompt_request_output_path
  )
)

validation_overview %>% kable()
```

```{r prompt-preview}
prompt_requests %>%
  mutate(prompt_text = str_trunc(prompt_text, 250)) %>%
  select(case_id, original_case_id, model_type, repeat_id, prompt_text) %>%
  slice_head(n = min(10, nrow(prompt_requests))) %>%
  kable(caption = "Prompt request dataset preview")
```

# API Helpers

```{r api-helpers}
detect_provider_from_url <- function(base_url) {
  url_lower <- str_to_lower(base_url)

  if (!nzchar(str_squish(url_lower))) {
    return(NA_character_)
  }

  if (str_detect(url_lower, "openai\\.com") || str_detect(url_lower, "/chat/completions")) {
    return("openai_compatible_chat")
  }

  if (str_detect(url_lower, "anthropic\\.com")) {
    return("anthropic")
  }

  stop("Unsupported base_url. Use an OpenAI-compatible chat completions endpoint or Anthropic endpoint.")
}

validate_generation_configuration <- function(base_url, model_id, n_repeats) {
  if (!is.character(base_url) || !nzchar(base_url)) {
    stop("base_url is blank. Set a local or remote model endpoint before running generation.")
  }
  if (!is.character(model_id) || !nzchar(model_id)) {
    stop("model_id is blank.")
  }
  if (!is.numeric(n_repeats) || length(n_repeats) != 1 || n_repeats < 1) {
    stop("n_repeats must be a positive integer.")
  }
}

extract_openai_chat_text <- function(body) {
  if (!is.null(body$choices) && length(body$choices) > 0) {
    out <- body$choices[[1]]$message$content
    if (!is.null(out)) {
      return(out)
    }
  }
  ""
}

extract_anthropic_text <- function(body) {
  if (!is.null(body$content) && length(body$content) > 0) {
    text_parts <- unlist(
      map(body$content, function(item) if (!is.null(item$text)) item$text else NULL)
    )
    if (length(text_parts) > 0) {
      return(paste(text_parts, collapse = "\n"))
    }
  }
  ""
}

call_openai_compatible_chat <- function(prompt_text, model_id, api_key, base_url, temperature, max_output_tokens) {
  body <- list(
    model = model_id,
    messages = list(
      list(role = "user", content = prompt_text)
    ),
    temperature = temperature,
    max_tokens = max_output_tokens
  )

  req <- request(base_url) %>%
    req_headers(`Content-Type` = "application/json") %>%
    req_body_json(body, auto_unbox = TRUE)

  if (nzchar(api_key)) {
    req <- req %>% req_headers(Authorization = paste("Bearer", api_key))
  }

  response <- req %>% req_perform()
  parsed <- resp_body_json(response, simplifyVector = FALSE)
  extract_openai_chat_text(parsed)
}

call_anthropic_response <- function(prompt_text, model_id, api_key, base_url, temperature, max_output_tokens) {
  if (!nzchar(api_key)) {
    stop("Anthropic API calls require api_key.")
  }

  body <- list(
    model = model_id,
    max_tokens = max_output_tokens,
    temperature = temperature,
    messages = list(
      list(role = "user", content = prompt_text)
    )
  )

  response <- request(base_url) %>%
    req_headers(
      `x-api-key` = api_key,
      `anthropic-version` = "2023-06-01",
      `content-type` = "application/json"
    ) %>%
    req_body_json(body, auto_unbox = TRUE) %>%
    req_perform()

  parsed <- resp_body_json(response, simplifyVector = FALSE)
  extract_anthropic_text(parsed)
}

generate_treatment_plan <- function(prompt_text, provider, model_id, api_key, base_url, temperature, max_output_tokens) {
  if (provider == "openai_compatible_chat") {
    return(call_openai_compatible_chat(
      prompt_text = prompt_text,
      model_id = model_id,
      api_key = api_key,
      base_url = base_url,
      temperature = temperature,
      max_output_tokens = max_output_tokens
    ))
  }

  if (provider == "anthropic") {
    return(call_anthropic_response(
      prompt_text = prompt_text,
      model_id = model_id,
      api_key = api_key,
      base_url = base_url,
      temperature = temperature,
      max_output_tokens = max_output_tokens
    ))
  }

  stop("Unsupported provider: ", provider)
}

safe_model_id <- function(x) {
  x %>%
    str_replace_all("[^A-Za-z0-9]+", "_") %>%
    str_replace_all("^_+|_+$", "")
}
```

# Optional Response Generation

```{r generation}
repeated_outputs_tbl <- tibble()
repeated_outputs_list <- list()

if (run_generation) {
  validate_generation_configuration(
    base_url = base_url,
    model_id = model_id,
    n_repeats = n_repeats
  )

  provider <- detect_provider_from_url(base_url)

  repeated_outputs_tbl <- pmap_dfr(
    prompt_requests,
    function(case_id, original_case_id, case_variant_id, vignette_text, prompt_text, model_type, prompt_type, repeat_id) {
      result <- tryCatch(
        {
          response_text <- generate_treatment_plan(
            prompt_text = prompt_text,
            provider = provider,
            model_id = model_type,
            api_key = api_key,
            base_url = base_url,
            temperature = temperature,
            max_output_tokens = max_output_tokens
          )

          tibble(
            case_id = case_id,
            original_case_id = original_case_id,
            model_type = model_type,
            provider = provider,
            prompt_type = prompt_type,
            repeat_id = repeat_id,
            temperature = temperature,
            treatment_plan = response_text,
            error_message = ""
          )
        },
        error = function(e) {
          tibble(
            case_id = case_id,
            original_case_id = original_case_id,
            model_type = model_type,
            provider = provider,
            prompt_type = prompt_type,
            repeat_id = repeat_id,
            temperature = temperature,
            treatment_plan = "",
            error_message = conditionMessage(e)
          )
        }
      )

      Sys.sleep(request_pause_seconds)
      result
    }
  )

  repeated_outputs_list <- split(repeated_outputs_tbl, repeated_outputs_tbl$case_id)
}
```

# Response Validation and Export

```{r response-validation-export}
if (run_generation) {
  expected_rows <- nrow(prompt_requests)
  actual_rows <- nrow(repeated_outputs_tbl)

  duplicate_response_keys <- repeated_outputs_tbl %>%
    count(case_id, model_type, repeat_id, name = "n") %>%
    filter(n > 1)

  response_validation <- tibble(
    check = c(
      "Expected response rows",
      "Actual response rows",
      "Duplicate response keys",
      "Non-empty treatment_plan rows",
      "Rows with error_message"
    ),
    value = c(
      expected_rows,
      actual_rows,
      nrow(duplicate_response_keys),
      sum(nzchar(repeated_outputs_tbl$treatment_plan)),
      sum(nzchar(repeated_outputs_tbl$error_message))
    )
  )

  response_validation %>% kable()

  if (actual_rows != expected_rows) {
    stop("Generated response row count does not match expected rows.")
  }
  if (nrow(duplicate_response_keys) > 0) {
    stop("Duplicate case_id + model_type + repeat_id rows detected in responses.")
  }
  if (sum(nzchar(repeated_outputs_tbl$treatment_plan)) == 0) {
    stop("All generated treatment_plan values are blank. Check base_url, model_id, and local model status.")
  }

  resolved_response_output_path <- response_output_path
  if (is.null(resolved_response_output_path)) {
    resolved_response_output_path <- file.path(
      "data/inner model validation",
      paste0("intra_model_validation_responses_", safe_model_id(model_id), ".csv")
    )
  }

  if (file.exists(resolved_response_output_path) && !overwrite_existing) {
    stop("Response output file already exists and overwrite_existing is FALSE: ", resolved_response_output_path)
  }

  write_csv(repeated_outputs_tbl, resolved_response_output_path)

  tibble(
    output = c("Prompt request dataset", "Generated response dataset"),
    path = c(prompt_request_output_path, resolved_response_output_path)
  ) %>%
    kable()
}
```

# Run Notes

The prompt request dataset is generated every time this R Markdown is rendered. To run a local model, set `run_generation <- TRUE`, set `model_id`, and set `base_url` to an OpenAI-compatible local endpoint such as `http://localhost:11434/v1/chat/completions` or `http://localhost:1234/v1/chat/completions`.
