---
title: "Claude Intra-Model Stability Analysis"
author: "Group 30"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 2
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)

required_packages <- c(
  "readr", "dplyr", "tidyr", "stringr", "purrr",
  "tibble", "knitr", "ggplot2"
)
missing_packages <- required_packages[
  !vapply(required_packages, requireNamespace, logical(1), quietly = TRUE)
]

if (length(missing_packages) > 0) {
  stop("Missing required packages: ", paste(missing_packages, collapse = ", "))
}

library(readr)
library(dplyr)
library(tidyr)
library(stringr)
library(purrr)
library(tibble)
library(knitr)
library(ggplot2)

find_project_root <- function(start_dir = getwd()) {
  candidates <- normalizePath(
    file.path(start_dir, c(".", "..", "../..", "../../..")),
    mustWork = FALSE
  )

  matched <- candidates[
    file.exists(file.path(candidates, "data/raw/final_results_gpt_claude_models.csv"))
  ]

  if (length(matched) == 0) {
    stop("Could not find project root containing data/raw/final_results_gpt_claude_models.csv.")
  }

  matched[[1]]
}

project_root <- find_project_root()
knitr::opts_knit$set(root.dir = project_root)
dir.create(file.path(project_root, "data/inner model validation"), showWarnings = FALSE, recursive = TRUE)
```

# Purpose

This workflow applies the structured `treatment_plan` parsing strategy to repeated Claude model outputs and computes intra-model stability metrics. The primary stability metric is variation ratio using a structured treatment signature derived from parsed medication, action, dose, and frequency fields.

# Configuration

```{r configuration}
model_label <- "claude-sonnet-4-6"
input_path <- "data/inner model validation/claude_intra_model_responses.csv"
output_dir <- "data/inner model validation"
expected_repeats_per_case <- 5L
manual_review_n <- 10L
```

# Load Data

```{r load-data}
raw_data <- read_csv(input_path, show_col_types = FALSE)

blank_names <- names(raw_data) == ""
if (any(blank_names)) {
  names(raw_data)[blank_names] <- paste0("blank_column_", seq_len(sum(blank_names)))
}

required_columns <- c(
  "case_id", "original_case_id", "case_variant_id", "model_type",
  "prompt_type", "repeat_id", "treatment_plan"
)
missing_columns <- setdiff(required_columns, names(raw_data))

if (length(missing_columns) > 0) {
  stop("Missing required columns: ", paste(missing_columns, collapse = ", "))
}

raw_data <- raw_data %>%
  mutate(
    row_id = row_number(),
    model = model_type,
    repeat_id = as.integer(repeat_id),
    case_variant_id = as.integer(case_variant_id)
  ) %>%
  select(
    row_id, case_id, original_case_id, case_variant_id, model,
    prompt_type, repeat_id, treatment_plan, everything()
  )

duplicate_keys <- raw_data %>%
  count(case_id, repeat_id, name = "n") %>%
  filter(n > 1)

if (nrow(duplicate_keys) > 0) {
  stop("Duplicate case_id + repeat_id rows detected in the Claude response file.")
}

repeat_counts <- raw_data %>%
  count(case_id, name = "n_repeats")

invalid_repeat_counts <- repeat_counts %>%
  filter(n_repeats != expected_repeats_per_case)

if (nrow(invalid_repeat_counts) > 0) {
  stop("Some case_id values do not have exactly ", expected_repeats_per_case, " repeats.")
}

raw_data %>%
  summarise(
    rows = n(),
    cases = n_distinct(case_id),
    original_cases = n_distinct(original_case_id),
    missing_treatment_plan = sum(is.na(treatment_plan) | treatment_plan == "")
  ) %>%
  kable(caption = paste("Data overview —", model_label))
```

# Fixed Medication Dictionary

```{r medication-dictionary}
medication_dictionary <- tribble(
  ~med_id, ~med_label, ~regex,
  "metformin", "Metformin", "\\bmetformin\\b",
  "sglt2_inhibitor", "SGLT2 inhibitor", "\\bsglt2\\b|sodium-glucose|empagliflozin|jardiance|dapagliflozin|farxiga|canagliflozin|invokana|ertugliflozin|steglatro",
  "empagliflozin", "Empagliflozin", "\\bempagliflozin\\b|\\bjardiance\\b",
  "dapagliflozin", "Dapagliflozin", "\\bdapagliflozin\\b|\\bfarxiga\\b",
  "canagliflozin", "Canagliflozin", "\\bcanagliflozin\\b|\\binvokana\\b",
  "ertugliflozin", "Ertugliflozin", "\\bertugliflozin\\b|\\bsteglatro\\b",
  "glp1_receptor_agonist", "GLP-1 receptor agonist", "\\bglp-?1\\b|glp1|semaglutide|ozempic|rybelsus|dulaglutide|trulicity|liraglutide|victoza|exenatide|byetta|lixisenatide|adlyxin|tirzepatide|mounjaro",
  "semaglutide", "Semaglutide", "\\bsemaglutide\\b|\\bozempic\\b|\\brybelsus\\b",
  "dulaglutide", "Dulaglutide", "\\bdulaglutide\\b|\\btrulicity\\b",
  "liraglutide", "Liraglutide", "\\bliraglutide\\b|\\bvictoza\\b",
  "exenatide", "Exenatide", "\\bexenatide\\b|\\bbyetta\\b|\\bbydureon\\b",
  "lixisenatide", "Lixisenatide", "\\blixisenatide\\b|\\badlyxin\\b",
  "tirzepatide", "Tirzepatide", "\\btirzepatide\\b|\\bmounjaro\\b",
  "dpp4_inhibitor", "DPP-4 inhibitor", "\\bdpp-?4\\b|sitagliptin|januvia|linagliptin|trajenta|saxagliptin|onglyza|alogliptin|nesina",
  "sitagliptin", "Sitagliptin", "\\bsitagliptin\\b|\\bjanuvia\\b",
  "linagliptin", "Linagliptin", "\\blinagliptin\\b|\\btrajenta\\b",
  "saxagliptin", "Saxagliptin", "\\bsaxagliptin\\b|\\bonglyza\\b",
  "alogliptin", "Alogliptin", "\\balogliptin\\b|\\bnesina\\b",
  "insulin", "Insulin", "\\binsulin\\b|glargine|lantus|basaglar|aspart|novolog|lispro|humalog|degludec|tresiba|detemir|levemir|nph|regular insulin",
  "basal_insulin", "Basal insulin", "basal insulin|glargine|lantus|basaglar|degludec|tresiba|detemir|levemir|nph",
  "insulin_glargine", "Insulin glargine", "\\bglargine\\b|\\blantus\\b|\\bbasaglar\\b",
  "insulin_degludec", "Insulin degludec", "\\bdegludec\\b|\\btresiba\\b",
  "insulin_detemir", "Insulin detemir", "\\bdetemir\\b|\\blevemir\\b",
  "rapid_acting_insulin", "Rapid-acting insulin", "rapid[- ]acting insulin|mealtime insulin|bolus insulin|prandial insulin|aspart|novolog|lispro|humalog|glulisine|apidra",
  "insulin_aspart", "Insulin aspart", "\\baspart\\b|\\bnovolog\\b",
  "insulin_lispro", "Insulin lispro", "\\blispro\\b|\\bhumalog\\b",
  "insulin_glulisine", "Insulin glulisine", "\\bglulisine\\b|\\bapidra\\b",
  "sulfonylurea", "Sulfonylurea", "\\bsulfonylurea\\b|glipizide|glyburide|glimepiride",
  "glipizide", "Glipizide", "\\bglipizide\\b",
  "glyburide", "Glyburide", "\\bglyburide\\b",
  "glimepiride", "Glimepiride", "\\bglimepiride\\b",
  "thiazolidinedione", "Thiazolidinedione", "\\bthiazolidinedione\\b|\\btzd\\b|pioglitazone|rosiglitazone",
  "pioglitazone", "Pioglitazone", "\\bpioglitazone\\b",
  "rosiglitazone", "Rosiglitazone", "\\brosiglitazone\\b",
  "alpha_glucosidase_inhibitor", "Alpha-glucosidase inhibitor", "alpha[- ]glucosidase|acarbose|miglitol",
  "acarbose", "Acarbose", "\\bacarbose\\b",
  "miglitol", "Miglitol", "\\bmiglitol\\b",
  "meglitinide", "Meglitinide", "\\bmeglitinide\\b|repaglinide|nateglinide",
  "repaglinide", "Repaglinide", "\\brepaglinide\\b",
  "nateglinide", "Nateglinide", "\\bnateglinide\\b",
  "pramlintide", "Pramlintide", "\\bpramlintide\\b|\\bsymlin\\b"
)
```

# Parsing Functions

```{r parsing-functions}
normalize_plan_text <- function(x) {
  x %>%
    coalesce("") %>%
    str_replace_all("\r\n", "\n") %>%
    str_replace_all("\r", "\n") %>%
    str_replace_all("\u2014|\u2013", "-") %>%
    str_squish()
}

split_plan_segments <- function(text) {
  text <- coalesce(text, "")
  raw_segments <- str_split(text, "\n|\\|")[[1]]

  segments <- raw_segments %>%
    str_replace_all("\u2014|\u2013", "-") %>%
    str_remove_all("^[-: ]+$") %>%
    str_squish()

  segments <- segments[
    segments != "" &
      !str_detect(str_to_lower(segments), "^[-]+$|^medication|^recommendation|^plan$")
  ]

  if (length(segments) == 0) {
    segments <- str_squish(text)
  }

  segments
}

has_pattern <- function(text, pattern) {
  str_detect(text, regex(pattern, ignore_case = TRUE))
}

extract_matching_text <- function(text, pattern) {
  segments <- split_plan_segments(text)
  matched <- segments[has_pattern(segments, pattern)]

  if (length(matched) == 0) {
    return("")
  }

  paste(unique(matched), collapse = " || ")
}

detect_action <- function(text, action_pattern) {
  text != "" & has_pattern(text, action_pattern)
}

extract_dose_value <- function(text) {
  match <- str_match(
    text,
    regex("(\\d+(?:\\.\\d+)?)\\s*(mg|mcg|units?|unit)\\b", ignore_case = TRUE)
  )

  ifelse(is.na(match[, 2]), 0, as.numeric(match[, 2]))
}

extract_dose_unit <- function(text) {
  match <- str_match(
    text,
    regex("(\\d+(?:\\.\\d+)?)\\s*(mg|mcg|units?|unit)\\b", ignore_case = TRUE)
  )

  ifelse(is.na(match[, 3]), "", str_to_lower(match[, 3]))
}

extract_frequency <- function(text) {
  frequency_pattern <- paste(
    c(
      "twice daily", "once daily", "once weekly", "weekly", "daily",
      "nightly", "every night", "at bedtime", "bedtime",
      "before each meal", "before meals", "with meals",
      "in the morning", "morning", "evening",
      "bid", "tid", "qhs", "qam"
    ),
    collapse = "|"
  )

  out <- str_extract(text, regex(frequency_pattern, ignore_case = TRUE))
  ifelse(is.na(out), "", str_to_lower(out))
}

action_patterns <- list(
  start = "\\b(start|add|initiate|begin|new)\\b",
  continue = "\\b(continue|maintain|keep)\\b",
  stop = "\\b(stop|discontinue|hold|avoid|do not use|not restart)\\b",
  reduce = "\\b(reduce|decrease|lower|dose reduction)\\b",
  increase = "\\b(increase|uptitrate|up-titrate|titrate up|raise)\\b",
  titrate = "\\b(titrate|uptitrate|up-titrate|adjust by)\\b"
)

parse_one_medication <- function(row_id, case_id, model, treatment_plan, med_id, med_label, regex) {
  plan_text <- normalize_plan_text(treatment_plan)
  raw_text <- extract_matching_text(plan_text, regex)

  mentioned <- as.integer(raw_text != "")
  stop <- as.integer(detect_action(raw_text, action_patterns$stop))
  start <- as.integer(detect_action(raw_text, action_patterns$start))
  continue <- as.integer(detect_action(raw_text, action_patterns$continue))
  reduce <- as.integer(detect_action(raw_text, action_patterns$reduce))
  increase <- as.integer(detect_action(raw_text, action_patterns$increase))
  titrate <- as.integer(detect_action(raw_text, action_patterns$titrate))
  dose_value <- if (mentioned == 1) extract_dose_value(raw_text) else 0
  dose_unit <- if (mentioned == 1) extract_dose_unit(raw_text) else ""
  frequency <- if (mentioned == 1) extract_frequency(raw_text) else ""

  action_present <- start == 1 | continue == 1 | reduce == 1 | increase == 1 | titrate == 1
  recommended <- as.integer(mentioned == 1 & stop == 0 & (action_present | dose_value > 0))
  conflict_flag <- as.integer(stop == 1 & (start == 1 | continue == 1 | increase == 1 | titrate == 1))

  tibble(
    row_id = row_id,
    case_id = case_id,
    model = model,
    med_id = med_id,
    med_label = med_label,
    mentioned = mentioned,
    recommended = recommended,
    start = start,
    continue = continue,
    stop = stop,
    reduce = reduce,
    increase = increase,
    titrate = titrate,
    dose_value = dose_value,
    dose_unit = dose_unit,
    frequency = frequency,
    raw_text = ifelse(mentioned == 1, raw_text, ""),
    conflict_flag = conflict_flag
  )
}
```

# Parse Repeated Responses

```{r parse-treatment-plans}
parsed_long <- pmap_dfr(
  raw_data %>% select(row_id, case_id, model, treatment_plan),
  function(row_id, case_id, model, treatment_plan) {
    pmap_dfr(
      medication_dictionary,
      function(med_id, med_label, regex) {
        parse_one_medication(
          row_id = row_id,
          case_id = case_id,
          model = model,
          treatment_plan = treatment_plan,
          med_id = med_id,
          med_label = med_label,
          regex = regex
        )
      }
    )
  }
)

parsed_wide_medications <- parsed_long %>%
  select(
    row_id, med_id, mentioned, recommended, start, continue, stop, reduce,
    increase, titrate, dose_value, dose_unit, frequency, raw_text, conflict_flag
  ) %>%
  pivot_wider(
    id_cols = row_id,
    names_from = med_id,
    values_from = c(
      mentioned, recommended, start, continue, stop, reduce, increase,
      titrate, dose_value, dose_unit, frequency, raw_text, conflict_flag
    ),
    names_glue = "{med_id}_{.value}",
    values_fill = list(
      mentioned = 0,
      recommended = 0,
      start = 0,
      continue = 0,
      stop = 0,
      reduce = 0,
      increase = 0,
      titrate = 0,
      dose_value = 0,
      dose_unit = "",
      frequency = "",
      raw_text = "",
      conflict_flag = 0
    )
  )

parsed_wide <- raw_data %>%
  left_join(parsed_wide_medications, by = "row_id")
```

# Unmatched Medication Audit

```{r unmatched-audit}
dictionary_regex_union <- paste(medication_dictionary$regex, collapse = "|")

candidate_stopwords <- c(
  "start", "add", "continue", "stop", "reduce", "increase", "dose", "from",
  "to", "then", "target", "goal", "maximum", "max", "minimum", "the",
  "and", "with", "before", "after", "daily", "weekly", "morning", "nightly",
  "bedtime", "titrate", "titration", "planned", "coming", "few", "months",
  "glucose", "fasting", "uacr", "egfr", "ckd", "a1c"
)

non_drug_candidate_pattern <- paste(
  c(
    "dose from", "increase dose", "decrease dose", "and then",
    "target", "goal", "glucose", "fasting", "postprandial",
    "above", "below", "toward", "maximum", "minimum",
    "few months", "coming months", "planned titration",
    "morning glucose", "afternoon glucose", "bedtime glucose"
  ),
  collapse = "|"
)

extract_dose_adjacent_candidates <- function(row_id, case_id, model, treatment_plan) {
  text <- normalize_plan_text(treatment_plan)
  candidate_pattern <- regex(
    "([a-z][a-z0-9/-]{2,}(?:\\s+[a-z][a-z0-9/-]{2,}){0,3})\\s+(?:xr\\s+|er\\s+)?\\d+(?:\\.\\d+)?\\s*(?:mg|mcg|units?|unit)\\b",
    ignore_case = TRUE
  )

  matches <- str_match_all(text, candidate_pattern)[[1]]

  if (nrow(matches) == 0) {
    return(tibble())
  }

  tibble(
    row_id = row_id,
    case_id = case_id,
    model = model,
    candidate_phrase = str_squish(str_to_lower(matches[, 2])),
    matched_full_text = str_squish(matches[, 1])
  )
}

unmatched_candidates <- pmap_dfr(
  raw_data %>% select(row_id, case_id, model, treatment_plan),
  extract_dose_adjacent_candidates
) %>%
  mutate(
    candidate_phrase = str_remove_all(candidate_phrase, "^\\b(start|add|continue|stop|reduce|increase|then|to|from)\\b\\s+"),
    candidate_phrase = str_squish(candidate_phrase),
    dictionary_matched = has_pattern(candidate_phrase, dictionary_regex_union),
    stopword_only = candidate_phrase %in% candidate_stopwords,
    non_drug_candidate = has_pattern(candidate_phrase, non_drug_candidate_pattern)
  ) %>%
  filter(!dictionary_matched, !stopword_only, !non_drug_candidate, candidate_phrase != "") %>%
  count(candidate_phrase, matched_full_text, sort = TRUE, name = "n") %>%
  arrange(desc(n), candidate_phrase)
```

# Build Treatment Signatures

```{r build-signatures}
signature_components <- parsed_long %>%
  filter(mentioned == 1 | recommended == 1) %>%
  mutate(
    rounded_dose_value = round(dose_value, 2),
    signature_component = paste(
      med_id,
      recommended,
      start,
      continue,
      stop,
      increase,
      reduce,
      titrate,
      rounded_dose_value,
      dose_unit,
      frequency,
      sep = "::"
    )
  ) %>%
  arrange(row_id, med_id)

treatment_signatures <- raw_data %>%
  select(row_id, case_id, original_case_id, case_variant_id, model, prompt_type, repeat_id, treatment_plan) %>%
  left_join(
    signature_components %>%
      group_by(row_id) %>%
      summarise(
        treatment_signature = paste(signature_component, collapse = " || "),
        .groups = "drop"
      ),
    by = "row_id"
  ) %>%
  mutate(
    treatment_signature = if_else(
      is.na(treatment_signature) | treatment_signature == "",
      "NO_MEDICATIONS_DETECTED",
      treatment_signature
    )
  )
```

# Intra-Model Stability Metrics

```{r stability-metrics}
case_signature_counts <- treatment_signatures %>%
  count(case_id, treatment_signature, name = "signature_count") %>%
  arrange(case_id, desc(signature_count), treatment_signature)

case_stability <- case_signature_counts %>%
  group_by(case_id) %>%
  summarise(
    n_repeats = sum(signature_count),
    modal_signature_count = max(signature_count),
    n_unique_signatures = n(),
    modal_treatment_signature = treatment_signature[which.max(signature_count)][1],
    variation_ratio = 1 - (modal_signature_count / n_repeats),
    .groups = "drop"
  ) %>%
  left_join(
    raw_data %>%
      distinct(case_id, original_case_id, case_variant_id),
    by = "case_id"
  ) %>%
  arrange(desc(variation_ratio), case_id)

variation_ratio_distribution <- case_stability %>%
  count(variation_ratio, name = "n_cases") %>%
  arrange(variation_ratio)

stability_summary <- tibble(
  metric = c(
    "Cases analyzed",
    "Expected repeats per case",
    "Mean variation ratio",
    "Median variation ratio",
    "Exact-match rate"
  ),
  value = c(
    nrow(case_stability),
    expected_repeats_per_case,
    round(mean(case_stability$variation_ratio), 3),
    round(median(case_stability$variation_ratio), 3),
    round(mean(case_stability$variation_ratio == 0), 3)
  )
)

exact_match_cases <- sum(case_stability$variation_ratio == 0)
high_instability_cases <- sum(case_stability$variation_ratio >= 0.4)
top_unstable_case <- case_stability %>%
  slice_head(n = 1)

key_medications <- c(
  "metformin", "sglt2_inhibitor", "glp1_receptor_agonist",
  "insulin", "basal_insulin", "insulin_glargine", "sitagliptin"
)

medication_class_agreement <- parsed_long %>%
  filter(med_id %in% key_medications) %>%
  group_by(case_id, med_id) %>%
  summarise(
    dominant_recommended_count = max(sum(recommended == 0), sum(recommended == 1)),
    repeat_level_agreement = dominant_recommended_count / n(),
    .groups = "drop"
  ) %>%
  group_by(med_id) %>%
  summarise(
    mean_repeat_level_agreement = mean(repeat_level_agreement),
    median_repeat_level_agreement = median(repeat_level_agreement),
    .groups = "drop"
  ) %>%
  arrange(desc(mean_repeat_level_agreement), med_id)

action_level_agreement <- parsed_long %>%
  pivot_longer(
    cols = c(start, continue, stop, increase, reduce, titrate),
    names_to = "action_type",
    values_to = "action_value"
  ) %>%
  group_by(case_id, med_id, action_type) %>%
  summarise(
    dominant_action_count = max(sum(action_value == 0), sum(action_value == 1)),
    repeat_level_agreement = dominant_action_count / n(),
    .groups = "drop"
  ) %>%
  group_by(action_type) %>%
  summarise(
    mean_repeat_level_agreement = mean(repeat_level_agreement),
    median_repeat_level_agreement = median(repeat_level_agreement),
    .groups = "drop"
  ) %>%
  arrange(desc(mean_repeat_level_agreement), action_type)

dose_stability <- parsed_long %>%
  filter(med_id %in% c("insulin", "basal_insulin", "insulin_glargine", "metformin", "empagliflozin", "sitagliptin")) %>%
  group_by(case_id, med_id) %>%
  filter(any(mentioned == 1 | recommended == 1)) %>%
  summarise(
    dominant_dose_count = max(table(paste(round(dose_value, 2), dose_unit, frequency, sep = "::"))),
    n_repeats = n(),
    exact_dose_match_rate = dominant_dose_count / n_repeats,
    .groups = "drop"
  ) %>%
  group_by(med_id) %>%
  summarise(
    mean_exact_dose_match_rate = mean(exact_dose_match_rate),
    median_exact_dose_match_rate = median(exact_dose_match_rate),
    .groups = "drop"
  ) %>%
  arrange(desc(mean_exact_dose_match_rate), med_id)
```

# Validation and Manual Review

```{r validation}
expected_rows <- nrow(raw_data)
actual_rows <- nrow(parsed_wide)
expected_long_rows <- nrow(raw_data) * nrow(medication_dictionary)
actual_long_rows <- nrow(parsed_long)

if (actual_rows != expected_rows) {
  stop("Wide parsed output row count does not match input row count.")
}

if (actual_long_rows != expected_long_rows) {
  stop("Long parsed output row count does not equal rows x medication dictionary size.")
}

stopped_but_recommended <- parsed_long %>%
  filter(stop == 1, recommended == 1)

if (nrow(stopped_but_recommended) > 0) {
  stop("Some stopped medications were incorrectly marked as recommended.")
}

manual_review_sample <- treatment_signatures %>%
  left_join(
    parsed_long %>%
      filter(mentioned == 1 | recommended == 1) %>%
      group_by(row_id) %>%
      summarise(
        parsed_summary = paste(
          paste0(
            med_id, " [rec=", recommended,
            ", start=", start,
            ", continue=", continue,
            ", stop=", stop,
            ", increase=", increase,
            ", reduce=", reduce,
            ", titrate=", titrate,
            ", dose=", round(dose_value, 2), " ", dose_unit,
            ", freq=", frequency, "]"
          ),
          collapse = " || "
        ),
        .groups = "drop"
      ),
    by = "row_id"
  ) %>%
  mutate(parsed_summary = coalesce(parsed_summary, "NO_MEDICATIONS_DETECTED")) %>%
  slice_head(n = min(manual_review_n, nrow(treatment_signatures)))

validation_summary <- tibble(
  check = c(
    "Input response rows",
    "Parsed wide rows",
    "Medication dictionary size",
    "Parsed long rows",
    "Expected long rows",
    "Cases with 5 repeats",
    "Duplicate case_id + repeat_id rows",
    "Stopped medications marked recommended",
    "Unmatched candidate rows"
  ),
  value = c(
    expected_rows,
    actual_rows,
    nrow(medication_dictionary),
    actual_long_rows,
    expected_long_rows,
    nrow(repeat_counts),
    nrow(duplicate_keys),
    nrow(stopped_but_recommended),
    nrow(unmatched_candidates)
  )
)

validation_summary %>% kable()
```

# Key Outputs

```{r key-outputs}
stability_summary %>% kable(caption = "Overall stability summary")

variation_ratio_distribution %>% kable(caption = "Variation ratio distribution")

case_stability %>%
  select(case_id, original_case_id, case_variant_id, n_repeats, n_unique_signatures, variation_ratio) %>%
  slice_head(n = min(10, nrow(case_stability))) %>%
  kable(caption = "Most unstable cases")

medication_class_agreement %>%
  kable(caption = "Medication-class repeat-level agreement")

action_level_agreement %>%
  kable(caption = "Action-level repeat agreement")
```

# Visual Summary and Interpretation

```{r visual-summary-data}
variation_ratio_plot_data <- variation_ratio_distribution %>%
  mutate(
    variation_ratio_label = factor(
      sprintf("%.1f", variation_ratio),
      levels = sprintf("%.1f", sort(unique(variation_ratio)))
    )
  )

most_unstable_plot_data <- case_stability %>%
  slice_head(n = min(10, nrow(case_stability))) %>%
  mutate(case_id = factor(case_id, levels = rev(case_id)))

medication_class_plot_data <- medication_class_agreement %>%
  mutate(med_id = factor(med_id, levels = rev(med_id)))

action_level_plot_data <- action_level_agreement %>%
  mutate(action_type = factor(action_type, levels = rev(action_type)))
```

```{r interpretation, results='asis'}
cat(
  paste0(
    "Across **", nrow(case_stability), "** ", model_label, " cases with **",
    expected_repeats_per_case,
    "** repeated generations each, the mean variation ratio was **",
    sprintf("%.2f", mean(case_stability$variation_ratio)),
    "** and the median was **",
    sprintf("%.2f", median(case_stability$variation_ratio)),
    "**. "
  ),
  paste0(
    "**", exact_match_cases, "** cases (",
    sprintf("%.0f", 100 * mean(case_stability$variation_ratio == 0)),
    "%) had perfectly identical parsed treatment plans across all repeats, while **",
    high_instability_cases,
    "** cases had variation ratio **>= 0.40**, indicating noticeable within-model instability. "
  ),
  paste0(
    "The most unstable case in the current sample was **",
    top_unstable_case$case_id,
    "** with variation ratio **",
    sprintf("%.2f", top_unstable_case$variation_ratio),
    "** and **",
    top_unstable_case$n_unique_signatures,
    "** unique parsed treatment signatures across 5 repeats."
  ),
  sep = ""
)
```

```{r variation-ratio-plot, fig.width=7, fig.height=4.5}
ggplot(variation_ratio_plot_data, aes(x = variation_ratio_label, y = n_cases)) +
  geom_col(fill = "#2F6DB3", width = 0.7) +
  geom_text(aes(label = n_cases), vjust = -0.3, size = 4) +
  labs(
    title = paste("Distribution of Case-Level Variation Ratio 2014", model_label),
    x = "Variation Ratio",
    y = "Number of Cases"
  ) +
  ylim(0, max(variation_ratio_plot_data$n_cases) + 1) +
  theme_minimal(base_size = 12)
```

```{r unstable-cases-plot, fig.width=8, fig.height=5}
ggplot(most_unstable_plot_data, aes(x = case_id, y = variation_ratio)) +
  geom_col(fill = "#C44E52", width = 0.7) +
  coord_flip() +
  labs(
    title = paste("Most Unstable Cases 2014", model_label),
    x = "Case ID",
    y = "Variation Ratio"
  ) +
  theme_minimal(base_size = 12)
```

```{r medication-agreement-plot, fig.width=8, fig.height=5}
ggplot(medication_class_plot_data, aes(x = med_id, y = mean_repeat_level_agreement)) +
  geom_col(fill = "#55A868", width = 0.7) +
  coord_flip() +
  labs(
    title = paste("Medication-Class Repeat-Level Agreement 2014", model_label),
    x = "Medication Class",
    y = "Mean Agreement"
  ) +
  ylim(0, 1) +
  theme_minimal(base_size = 12)
```

```{r action-agreement-plot, fig.width=7, fig.height=4.5}
ggplot(action_level_plot_data, aes(x = action_type, y = mean_repeat_level_agreement)) +
  geom_col(fill = "#8172B2", width = 0.7) +
  coord_flip() +
  labs(
    title = paste("Action-Level Repeat Agreement 2014", model_label),
    x = "Action Type",
    y = "Mean Agreement"
  ) +
  ylim(0, 1) +
  theme_minimal(base_size = 12)
```

# Export Processed Data

```{r export}
safe_label <- gsub("[^A-Za-z0-9]+", "_", model_label)

write_csv(parsed_wide, file.path(output_dir, paste0(safe_label, "_intra_model_parsed_wide.csv")))
write_csv(parsed_long, file.path(output_dir, paste0(safe_label, "_intra_model_parsed_long.csv")))
write_csv(unmatched_candidates, file.path(output_dir, paste0(safe_label, "_intra_model_unmatched_candidates.csv")))
write_csv(treatment_signatures, file.path(output_dir, paste0(safe_label, "_intra_model_treatment_signatures.csv")))
write_csv(case_signature_counts, file.path(output_dir, paste0(safe_label, "_intra_model_signature_counts.csv")))
write_csv(case_stability, file.path(output_dir, paste0(safe_label, "_intra_model_case_stability.csv")))
write_csv(variation_ratio_distribution, file.path(output_dir, paste0(safe_label, "_intra_model_variation_ratio_distribution.csv")))
write_csv(stability_summary, file.path(output_dir, paste0(safe_label, "_intra_model_stability_summary.csv")))
write_csv(medication_class_agreement, file.path(output_dir, paste0(safe_label, "_intra_model_medication_class_agreement.csv")))
write_csv(action_level_agreement, file.path(output_dir, paste0(safe_label, "_intra_model_action_level_agreement.csv")))
write_csv(dose_stability, file.path(output_dir, paste0(safe_label, "_intra_model_dose_stability.csv")))
write_csv(manual_review_sample, file.path(output_dir, paste0(safe_label, "_intra_model_manual_review_sample.csv")))
saveRDS(parsed_wide, file.path(output_dir, paste0(safe_label, "_intra_model_parsed_wide.rds")))

tibble(
  output = c(
    "Parsed wide data",
    "Parsed long data",
    "Unmatched candidate audit",
    "Treatment signatures",
    "Case stability metrics",
    "Medication-class agreement",
    "Action-level agreement",
    "Dose stability",
    "Manual review sample",
    "Parsed wide RDS"
  ),
  path = c(
    file.path(output_dir, paste0(safe_label, "_intra_model_parsed_wide.csv")),
    file.path(output_dir, paste0(safe_label, "_intra_model_parsed_long.csv")),
    file.path(output_dir, paste0(safe_label, "_intra_model_unmatched_candidates.csv")),
    file.path(output_dir, paste0(safe_label, "_intra_model_treatment_signatures.csv")),
    file.path(output_dir, paste0(safe_label, "_intra_model_case_stability.csv")),
    file.path(output_dir, paste0(safe_label, "_intra_model_medication_class_agreement.csv")),
    file.path(output_dir, paste0(safe_label, "_intra_model_action_level_agreement.csv")),
    file.path(output_dir, paste0(safe_label, "_intra_model_dose_stability.csv")),
    file.path(output_dir, paste0(safe_label, "_intra_model_manual_review_sample.csv")),
    file.path(output_dir, paste0(safe_label, "_intra_model_parsed_wide.rds"))
  )
) %>%
  kable()
```
