---
title: "Descriptive Statistics"
date: "2026-06-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set-Up
```{r packages}
# Packages
library(dplyr)
library(tidyr)
library(readr)
library(knitr)
library(kableExtra)
library(ggplot2)
library(stringr)
library(forcats)
```

```{r Path}
# Path
p <- "../data/processed/changchangpan/capstone_llm_data_explore_r_outputs"
```


# Vignette Characteristics
100 synthetic clinical vignettes evaluated by three LLMs: Claude Opus 4.6,
Claude Sonnet 4.6, and GPT-5.4 (300 total responses).

## Continuous Variables
```{r continuous-vignettes}
read_csv(file.path(p, "vignette_numeric_summary.csv"), show_col_types = FALSE) %>%
  mutate(across(where(is.numeric), ~ round(., 1))) %>%
  rename(Variable = variable, N = n, Mean = mean, SD = sd,
         Median = median, Q1 = q1, Q3 = q3, Min = min, Max = max) %>%
  kable(caption = "Continuous Characteristics (n = 100 vignettes)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## Binary Characteristics
```{r binary-vignettes}
read_csv(file.path(p, "vignette_binary_summary.csv"), show_col_types = FALSE) %>%
  filter(value == "Yes") %>%
  select(Variable = variable, `n (Yes)` = n, `% (Yes)` = pct) %>%
  kable(caption = "Binary Characteristics — % present (n = 100 vignettes)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```


# Clinical Subgroup Distributions
```{r subgroup-dist}
read_csv(file.path(p, "clinical_subgroup_counts.csv"), show_col_types = FALSE) %>%
  rename(Group = group, Category = category, N = n, `%` = pct) %>%
  kable(caption = "Clinical Subgroup Distributions (n = 100 vignettes)") %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  collapse_rows(columns = 1, valign = "top")
```


# Missingness
```{r missingness}
read_csv(file.path(p, "missingness_key_variables.csv"), show_col_types = FALSE)%>%
  rename(Variable = variable, `Missing (n)` = missing_n, `Missing (%)` =
           missing_pct) %>%
  arrange(desc(`Missing (%)`)) %>%
  kable(caption = "Missingness in Key Variables (300 rows)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

**Note:** `current_dm_drugs` is missing in 49% of rows because newly diagnosed
patients had no prior diabetes medications.


# Data Quality
```{r data-quality}
read_csv(file.path(p, "data_quality_basic_summary.csv"), show_col_types = FALSE)%>%
  kable(caption = "Data Quality Summary") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```


