---
title: "Descriptive Statistics"
date: "2026-05-31"
output: html_document
---

# Set-Up
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Libraries
library(dplyr)
library(readr)
library(knitr)
library(kableExtra)
library(ggplot2)
library(tidyr)
library(forcats)

# Path
p <- "../data/processed/changchangpan/capstone_llm_data_explore_r_outputs"
```


# Vignette Characteristics
100 synthetic clinical vignettes evaluated by three LLMs: Claude Opus 4.6,
Claude Sonnet 4.6, and GPT-5.4 (300 total responses).

## Continuous Variables
```{r continuous-vignettes}
read_csv(file.path(p, "vignette_numeric_summary.csv"), show_col_types = FALSE) %>%
  mutate(across(where(is.numeric), ~ round(., 1))) %>%
  rename(Variable = variable, N = n, Mean = mean, SD = sd,
         Median = median, Q1 = q1, Q3 = q3, Min = min, Max = max) %>%
  kable(caption = "Continuous Characteristics (n = 100 vignettes)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## Binary Characteristics
```{r binary-vignettes}
read_csv(file.path(p, "vignette_binary_summary.csv"), show_col_types = FALSE) %>%
  filter(value == "Yes") %>%
  select(Variable = variable, `n (Yes)` = n, `% (Yes)` = pct) %>%
  kable(caption = "Binary Characteristics - % present (n = 100 vignettes)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

**Key Takeaway:** The vignette population is complex at a clinical level and
skewed toward high-risk patients - 54% have CKD, 70% have hypertension, and 67%
have hyperlipidemia. Mean HbA1c is 8.3% and mean eGFR is 59 mL/min/1.73m²,
placing many patients in the CKD stage 3 range where medication choices
(especially metformin and SGLT2 inhibitors) require consideration.


# Clinical Subgroup Distributions
```{r subgroup-dist}
read_csv(file.path(p, "clinical_subgroup_counts.csv"), show_col_types = FALSE) %>%
  rename(Group = group, Category = category, N = n, `%` = pct) %>%
  kable(caption = "Clinical Subgroup Distributions (n = 100 vignettes)") %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  collapse_rows(columns = 1, valign = "top")
```

**Key Takeaway:** The majority of vignettes (58%) fall in the HbA1c 8-8.9%
range, reflecting moderately uncontrolled diabetes. On the renal side, 52% of
vignettes have eGFR below 60 (CKD stages 3-5) and 70% have at least
microalbuminuria (UACR >= 30 mg/g). Both are strong guideline indications for
SGLT2 inhibitor or GLP-1 RA use.


# Missingness
```{r missingness}
read_csv(file.path(p, "missingness_key_variables.csv"), show_col_types = FALSE)%>%
  rename(Variable = variable, `Missing (n)` = missing_n, `Missing (%)` =
           missing_pct) %>%
  arrange(desc(`Missing (%)`)) %>%
  kable(caption = "Missingness in Key Variables (300 rows)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

**Key Takeaway:** Missingness is minimal for all clinical variables. The only
substantial missing data is `current_dm_drugs` (49%), which is expected. 
Newly diagnosed patients had no prior diabetes medications. All key covariates
used in analysis (HbA1c, eGFR, UACR, comorbidities) are complete.


# Data Quality

## Summary
```{r data-quality}
read_csv(file.path(p, "data_quality_basic_summary.csv"), show_col_types = FALSE)%>%
  kable(caption = "Data Quality Summary") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

## Case Completeness
```{r case-completeness}
read_csv(file.path(p, "case_completeness_by_case_id.csv"), show_col_types = FALSE)%>%
  rename(`Case ID` = case_id, `N models observed` = n_models_observed,
         `Expected N models` = expected_n_models,
         `Complete` = complete_case_id) %>%
  filter(Complete == FALSE) %>%
  kable(caption = "Incomplete Cases (fewer than 3 model responses)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

## Duplicated Case-Model Combinations
```{r duplicates}
read_csv(file.path(p, "duplicated_case_model_combinations.csv"),
         show_col_types = FALSE) %>%
  rename(`Case ID` = case_id, Model = model_short, `N rows` = n_rows) %>%
  kable(caption = "Duplicated Case-Model Combinations (expected = 1 row each)")%>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

**Key Takeaway:** Data quality is generally good. The main known issue is
DM026, which has 3 distinct vignette contents mapped to one case ID and
should be treated with caution. Cases DM085-DM088 are entirely absent
from the dataset.

