---
title: "Follow-up Analyses"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(dplyr)
library(readr)
library(knitr)
library(kableExtra)
library(broom)
library(tidyr)
library(stringr)

library(here)
ind_path <- here("data/processed/changchangpan/capstone_llm_data_explore_r_outputs/processed_llm_treatment_indicators.csv")
dis_path <- here("data/processed/changchangpan/capstone_llm_data_explore_r_outputs/high_disagreement_cases.csv")
raw_path <- here("data/raw/final_results_gpt_claude_models.csv")

ind <- read_csv(ind_path, show_col_types = FALSE)

# Patient-level frame: one row per case for patient covariates
case <- ind %>% distinct(case_id, .keep_all = TRUE)
```

# 1. Disagreement vs. Complexity

Testing James's hypothesis that model disagreement rises with clinical
complexity — operationalized as number of comorbidities and lower eGFR.

```{r disagreement-corr}
dis <- read_csv(dis_path, show_col_types = FALSE)

comorbid_cols <- c(
  "heart_failure_present", "chronic_kidney_disease", "mi_history",
  "stroke_tia_history", "coronary_artery_disease", "atrial_fibrillation",
  "peripheral_neuropathy", "diabetic_retinopathy", "hypertension",
  "hyperlipidemia", "hypothyroidism", "depression", "copd", "asthma",
  "anemia", "osteoporosis", "rheumatoid_arthritis", "cancer_breast",
  "cancer_colon", "cancer_lung", "cancer_prostate", "cancer_endometrial"
)
comorbid_cols <- intersect(comorbid_cols, names(case))

case <- case %>%
  mutate(n_comorbid = rowSums(across(all_of(comorbid_cols), ~ . == "Yes"),
                              na.rm = TRUE)) %>%
  left_join(dis %>% select(case_id, disagreement_score), by = "case_id") %>%
  mutate(disagreement_score = replace_na(disagreement_score, 0))

c1 <- cor.test(case$n_comorbid, case$disagreement_score, method = "spearman")
c2 <- cor.test(case$egfr_ml_min_1_73m2, case$disagreement_score,
               method = "spearman")

tibble(
  Comparison = c("Disagreement vs. # comorbidities",
                 "Disagreement vs. eGFR"),
  `Spearman rho` = round(c(c1$estimate, c2$estimate), 3),
  `p-value` = round(c(c1$p.value, c2$p.value), 4)
) %>%
  kable(caption = paste0("Spearman correlations (n = ", nrow(case), " cases)")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

```{r disagreement-by-egfr}
case %>%
  mutate(egfr_band = cut(egfr_ml_min_1_73m2, c(0, 30, 45, 60, Inf),
                         labels = c("<30", "30-44", "45-59", ">=60"),
                         right = FALSE)) %>%
  group_by(`eGFR band` = egfr_band) %>%
  summarise(`Mean disagreement` = round(mean(disagreement_score), 2),
            N = n(), .groups = "drop") %>%
  kable(caption = "Mean disagreement score by eGFR band") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

**Key Takeaway:** Both correlations run in the direction James predicted —
disagreement increases with comorbidity count (rho ~ +0.18) and decreases with
eGFR (rho ~ -0.19) — but neither reaches significance at this sample size
(both p ~ 0.07-0.08). The eGFR-band table is the clearer signal: mean
disagreement is highest in the CKD groups (~2.1 at eGFR <30, ~1.9 at 30-44)
and falls to ~1.5 once eGFR exceeds 45. This is consistent with the
"guidelines are clear for simple patients, ambiguous for complex ones"
interpretation, and supports the case for expanding the sample if we want to
confirm it statistically.

# 2. Baseline Diabetes Medications

```{r baseline-dm-drugs}
drug_pat <- c(
  "Metformin"        = "metformin",
  "SGLT2 inhibitor"  = "gliflozin|empagliflozin|dapagliflozin|canagliflozin",
  "GLP-1 RA"         = "semaglutide|dulaglutide|liraglutide|tirzepatide|exenatide|lixisenatide",
  "DPP-4 inhibitor"  = "gliptin|sitagliptin|linagliptin|saxagliptin",
  "Sulfonylurea"     = "glipizide|glyburide|glimepiride|gliclazide",
  "Basal insulin"    = "glargine|detemir|degludec|NPH|isophane",
  "Prandial insulin" = "aspart|lispro|glulisine"
)

drugs_lc <- tolower(coalesce(case$current_dm_drugs, ""))
none_n <- sum(drugs_lc == "")

tibble(
  `Drug class` = names(drug_pat),
  `n cases` = sapply(drug_pat, function(p) sum(str_detect(drugs_lc, p)))
) %>%
  bind_rows(tibble(`Drug class` = "No diabetes medication (newly dx)",
                   `n cases` = none_n)) %>%
  mutate(`%` = round(100 * `n cases` / nrow(case), 1)) %>%
  arrange(desc(`n cases`)) %>%
  kable(caption = paste0("Baseline diabetes medications (n = ", nrow(case),
        " cases; classes overlap, patients may be on >1)")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

**Key Takeaway:** Roughly half the population is on no diabetes medication at
baseline (newly diagnosed). Among those on therapy, GLP-1 RAs, metformin, and
SGLT2 inhibitors are the most common classes. Classes overlap because patients
may be on combination regimens, so percentages do not sum to 100.

# 3. Race / Ethnicity

```{r race-dist}
case %>%
  mutate(race_ethnicity = coalesce(race_ethnicity, "Unknown")) %>%
  count(`Race / Ethnicity` = race_ethnicity) %>%
  mutate(`%` = round(100 * n / sum(n), 1)) %>%
  arrange(desc(n)) %>%
  kable(caption = paste0("Race / ethnicity distribution (n = ", nrow(case),
                         " cases)")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

**Key Takeaway:** Race/ethnicity is balanced by design across the four groups.

# 4. GPT-5.4 Metformin Review (eGFR < 30)

Pulling the verbatim GPT-5.4 plans for low-eGFR cases to check the apparent
metformin guideline-adherence gap.

```{r metformin-review}
raw <- read_csv(raw_path, show_col_types = FALSE)

raw %>%
  filter(model == "azure/gpt-5.4", egfr_ml_min_1_73m2 < 30) %>%
  mutate(on_metformin = str_detect(tolower(coalesce(current_dm_drugs, "")),
                                   "metformin")) %>%
  arrange(case_id) %>%
  select(`Case` = case_id, `eGFR` = egfr_ml_min_1_73m2,
         `On metformin at baseline` = on_metformin,
         `Baseline DM drugs` = current_dm_drugs) %>%
  kable(caption = "GPT-5.4 cases with eGFR < 30") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

**Key Takeaway:** The "15% flagged metformin" figure is an artifact of the
regex-derived `metformin_stop_or_caution` flag, whose denominator includes all
eGFR <30 cases. Only the small number of patients actually on metformin at
baseline represent real metformin decisions; in those, GPT-5.4 discontinued it
and cited the eGFR. No model recommends starting metformin below eGFR 30. The
full verbatim `treatment_plan` text for these cases should be reviewed manually
(printed below) before reporting.

```{r metformin-text, results='asis'}
txt <- raw %>%
  filter(model == "azure/gpt-5.4", egfr_ml_min_1_73m2 < 30) %>%
  arrange(case_id)

for (i in seq_len(nrow(txt))) {
  cat("\n\n**", txt$case_id[i], "** (eGFR ", txt$egfr_ml_min_1_73m2[i], ")\n\n",
      sep = "")
  cat("> ", gsub("\n", "\n> ", txt$treatment_plan[i]), "\n", sep = "")
}
```

# 5. Univariate Logistic Regression (eGFR >= 60 reference)

Re-running the medication-class associations as univariate models with eGFR as
the sole predictor and eGFR >= 60 as the reference category, to address the
complete separation seen in the multivariable models.

```{r univariate-logit}
ind <- ind %>%
  mutate(egfr_cat = case_when(
    egfr_ml_min_1_73m2 < 30 ~ "<30",
    egfr_ml_min_1_73m2 < 45 ~ "30-44",
    egfr_ml_min_1_73m2 < 60 ~ "45-59",
    TRUE                    ~ ">=60"),
    egfr_cat = relevel(factor(egfr_cat), ref = ">=60"))

run_one <- function(outcome) {
  f <- as.formula(paste(outcome, "~ egfr_cat"))
  suppressWarnings(glm(f, data = ind, family = binomial)) %>%
    tidy(exponentiate = TRUE, conf.int = TRUE) %>%
    filter(term != "(Intercept)") %>%
    mutate(Outcome = outcome,
           Level = paste0(sub("egfr_cat", "", term), " vs >=60")) %>%
    select(Outcome, Level, OR = estimate, CI_low = conf.low,
           CI_high = conf.high, p = p.value)
}

outcomes <- c("metformin_active", "sglt2i_active",
              "glp1_active", "insulin_any_active")

bind_rows(lapply(outcomes, run_one)) %>%
  mutate(across(c(OR, CI_low, CI_high), ~ round(., 2)),
         p = round(p, 4),
         note = ifelse(is.infinite(CI_high) | CI_high > 1e6,
                       "separation - report as proportion", "")) %>%
  kable(caption = "Univariate logistic regression: medication class by eGFR (ref = >=60)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

```{r metformin-separation-proportion}
ind %>%
  filter(egfr_cat == "<30") %>%
  summarise(N = n(),
            `N metformin active` = sum(metformin_active, na.rm = TRUE),
            `%` = round(100 * mean(metformin_active, na.rm = TRUE), 1)) %>%
  kable(caption = "Metformin in eGFR < 30 (report as proportion, not OR)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

**Key Takeaway:** The univariate models with eGFR >= 60 as reference are
well-behaved (sensible CIs), unlike the multivariable models. Metformin is
sharply less likely as eGFR falls (OR ~0.24 at 45-59, ~0.04 at 30-44); SGLT2
inhibitors peak in the moderate-CKD bands and drop below 30; insulin rises
steeply at low eGFR. The metformin eGFR <30 cell remains a complete-separation
case (0 of 39 responses recommend metformin) and is reported as a simple
proportion rather than an odds ratio.
```
